### 2023-05-23 paper.Rmd
Abstract

---
title: "Implications of afforestation policies for global markets"
author: "Eric Davis, Maros Ivanic, Brent Sohngen"
output: bookdown::pdf_document2
toc: FALSE
abstract: "The projected growth of the population and incomes as implied by various shared socioeconomic pathways is expected to create  pressures on converting forests into farmland as food prices increase. Depending on the specific path taken, SSP1 through SSP5, and the continuation of the current trends in agricultural productivity growth, the market impacts are expected to vary but generally resulting in increases in food prices. Policies aimed at reversing the deforestation trend by afforesting crop land would result in severe negative market impacts such as  higher food prices, especially in low agriculutrla productivity/high population growht scenarios."
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 2023-05-29 paper.Rmd
Update paper.Rmd

@@
 toc: FALSE
+bibliography: references.bib
 abstract: "The projected growth of the population and incomes as implied by various shared socioeconomic pathways is expected to create  pressures on converting forests into farmland as food prices increase.
@@
 ```
+
+\doublespacing
+
+# Introduction

### 2023-05-29 paper.Rmd
Update paper.Rmd

@@
 # Introduction
+Forests play an important role in maintaining and growing the land-based carbon sink, and with increasingly negative climate impacts, the pressure to enact afforestation policies is increasing.
+The potential gain from afforestation projects, however, is uncertain, with estimates ranging from 42 gigatonnes [@veldman19] to 205 gigatonnes [@bastin19].
+At the same time, forests are continually threatened by issues such as population growth, rising incomes, and rising food prices that put pressure on policymakers to allow greater conversion on forestland into farmland.
+The manner is which these various concerns are addressed is certain to have significant impacts on the global community.
+
+This project contributes to the literature by melding together two widely accepted global models, the Global Trade Analysis Project (GTAP) computable general equilibrium model and the Global Timber Model (GTM), to better assess the outcomes of several likely scenarios.

### 2023-05-30 paper.Rmd
Update paper.Rmd

@@
 The potential gain from afforestation projects, however, is uncertain, with estimates ranging from 42 gigatonnes [@veldman19] to 205 gigatonnes [@bastin19].
-At the same time, forests are continually threatened by issues such as population growth, rising incomes, and rising food prices that put pressure on policymakers to allow greater conversion on forestland into farmland.
-The manner is which these various concerns are addressed is certain to have significant impacts on the global community.
+At the same time, forests are continually threatened by issues such as rising incomes and food prices and population growth that put pressure on policymakers to allow greater conversion on forestland into farmland.
+The manner is which these various oftentimes competing concerns are addressed will no doublt have significant impacts on the global community.
+
+There is growing attention being placed globally to reforestation efforts, which is highligted by the UN's declaration of the years from 2021 through 2030 as the decade of ecosystem resotration [@sewell2020].
 
 This project contributes to the literature by melding together two widely accepted global models, the Global Trade Analysis Project (GTAP) computable general equilibrium model and the Global Timber Model (GTM), to better assess the outcomes of several likely scenarios.
+
+Scenarios
+1) Full attainment of UN strategic plan for forests 2030 (UNFF):  This says that all deforestation will halt and there will be a 3% (120 million hectare) increase in forest area worldwide by 2030 [@sewell2020].
+2)

### 2023-05-30 paper.Rmd
Update paper.Rmd

@@
 # Introduction
-Forests play an important role in maintaining and growing the land-based carbon sink, and with increasingly negative climate impacts, the pressure to enact afforestation policies is increasing.
-The potential gain from afforestation projects, however, is uncertain, with estimates ranging from 42 gigatonnes [@veldman19] to 205 gigatonnes [@bastin19].
+Forests play an important role in maintaining and growing the land-based carbon sink, currently sequestering about 30% of GHG emmissions [@roe2019].
+With increasingly negative climate impacts, the pressure to enhance the terrestrial sink is growing, and one of the tools to achieve this is through afforestation.
+In fact, many of the Nationally Determined Contributions (NDCs) that countries have submitted that detail their plans to keep warming under the 2C threshold rely to a significant degree on afforestation projects [@roe2019].
+The gains that can be realized, however, are uncertain, with estimates ranging from 42 gigatonnes [@veldman19] to 205 gigatonnes [@bastin19].
 At the same time, forests are continually threatened by issues such as rising incomes and food prices and population growth that put pressure on policymakers to allow greater conversion on forestland into farmland.
 The manner is which these various oftentimes competing concerns are addressed will no doublt have significant impacts on the global community.
+Previous research has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017].
 
@@
 1) Full attainment of UN strategic plan for forests 2030 (UNFF):  This says that all deforestation will halt and there will be a 3% (120 million hectare) increase in forest area worldwide by 2030 [@sewell2020].
-2)
+2) Full attainment of AFR100:  This says that there will be a 100 million hectare increase in forest area in Sub-Saharan Africa by 2030 [@sewell2020].
+3) Full attainment of goals of Bonn Challenge and New York Declaration on Forests:  This says that all deforestation will halt and there will be a 200 million hectare increase in forest area worldwide by 2020 and a 350 million hectare increase in forest area worldwide by 2030 [@sewell2020].

### 2023-05-30 paper.Rmd
Update paper.Rmd

@@
 With increasingly negative climate impacts, the pressure to enhance the terrestrial sink is growing, and one of the tools to achieve this is through afforestation.
-In fact, many of the Nationally Determined Contributions (NDCs) that countries have submitted that detail their plans to keep warming under the 2C threshold rely to a significant degree on afforestation projects [@roe2019].
+In fact, many of the Nationally Determined Contributions (NDCs)  countries have submitted that detail their plans to keep warming under the 2C threshold rely to a significant degree on afforestation projects [@roe2019].
 The gains that can be realized, however, are uncertain, with estimates ranging from 42 gigatonnes [@veldman19] to 205 gigatonnes [@bastin19].
-At the same time, forests are continually threatened by issues such as rising incomes and food prices and population growth that put pressure on policymakers to allow greater conversion on forestland into farmland.
-The manner is which these various oftentimes competing concerns are addressed will no doublt have significant impacts on the global community.
-Previous research has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017].
+At the same time, forests are continually threatened by issues such as rising incomes and food prices and population growth that put pressure on policymakers to allow greater conversion of forestland into farmland.
+Previous research though has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017].Thus, the manner in which these various and oftentimes competing concerns are addressed will have significant impacts on the global community.
 
-There is growing attention being placed globally to reforestation efforts, which is highligted by the UN's declaration of the years from 2021 through 2030 as the decade of ecosystem resotration [@sewell2020].
+There is growing attention being placed globally to reforestation efforts, which is highligted by the UN's declaration of the years from 2021 through 2030 as the UN Decade of Ecosystem Resotration [@sewell2020].
 
@@
 1) Full attainment of UN strategic plan for forests 2030 (UNFF):  This says that all deforestation will halt and there will be a 3% (120 million hectare) increase in forest area worldwide by 2030 [@sewell2020].
-2) Full attainment of AFR100:  This says that there will be a 100 million hectare increase in forest area in Sub-Saharan Africa by 2030 [@sewell2020].
-3) Full attainment of goals of Bonn Challenge and New York Declaration on Forests:  This says that all deforestation will halt and there will be a 200 million hectare increase in forest area worldwide by 2020 and a 350 million hectare increase in forest area worldwide by 2030 [@sewell2020].
+2) Full attainment of AFR100:  This says that there will be a 100 million hectare increase in forest area just in Sub-Saharan Africa by 2030 [@sewell2020].
+3) Full attainment of goals of Bonn Challenge and New York Declaration on Forests:  This says that all deforestation will halt and there will be a 350 million hectare increase in forest area worldwide by 2030 [@sewell2020].

### 2023-05-30 paper.Rmd
Update paper.Rmd

@@
 At the same time, forests are continually threatened by issues such as rising incomes and food prices and population growth that put pressure on policymakers to allow greater conversion of forestland into farmland.
-Previous research though has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017].Thus, the manner in which these various and oftentimes competing concerns are addressed will have significant impacts on the global community.
+Previous research though has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017].
+
+In addition, afforestation appears to not provide the same benefits in all regions of the world.
+In particular, boreal forests appear to provide the smallest net benefits, as the carbon sequestrations benefits of conifer forests are mitigated by the reduction in albedo in winter [@mykleby2017] [@li2015].
+That contrasts with the results for tropical forests where forests there provide a cooling impact due to changes in both albedo and evapotranspiration [@li2015].
+Thus, the manner in which these various and oftentimes competing concerns are addressed will have significant impacts on the global community.
 

### 2023-05-31 paper.Rmd
Update paper.Rmd

@@
 In fact, many of the Nationally Determined Contributions (NDCs)  countries have submitted that detail their plans to keep warming under the 2C threshold rely to a significant degree on afforestation projects [@roe2019].
-The gains that can be realized, however, are uncertain, with estimates ranging from 42 gigatonnes [@veldman19] to 205 gigatonnes [@bastin19].
+This is because of forests proven ability to sequester carbon in the stems of trees [@davis2022] as well as in their soils [@guo2021].
+The gains that can be realized, however, are uncertain, with estimates ranging from 42 gigatonnes [@veldman2019] to 205 gigatonnes [@bastin2019].
 At the same time, forests are continually threatened by issues such as rising incomes and food prices and population growth that put pressure on policymakers to allow greater conversion of forestland into farmland.

### 2023-05-31 paper.Rmd
Update paper.Rmd

--- 2023-05-30 paper.Rmd
@@
 In fact, many of the Nationally Determined Contributions (NDCs)  countries have submitted that detail their plans to keep warming under the 2C threshold rely to a significant degree on afforestation projects [@roe2019].
-The gains that can be realized, however, are uncertain, with estimates ranging from 42 gigatonnes [@veldman19] to 205 gigatonnes [@bastin19].
+This is because of forests proven ability to sequester carbon not only the stems of trees [@davis2022] but their soils as well [@guo2021].
 At the same time, forests are continually threatened by issues such as rising incomes and food prices and population growth that put pressure on policymakers to allow greater conversion of forestland into farmland.
 Previous research though has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017].
+Thus, the manner in which these various and oftentimes competing concerns are addressed will have significant impacts on the global community.
 
-In addition, afforestation appears to not provide the same benefits in all regions of the world.
+The global gains that can be realized from afforestion, however, are uncertain, with estimates ranging from 42 gigatonnes [@veldman2019] to 205 gigatonnes [@bastin2019], as afforestation appears to not provide the same benefits in all regions of the world.
 In particular, boreal forests appear to provide the smallest net benefits, as the carbon sequestrations benefits of conifer forests are mitigated by the reduction in albedo in winter [@mykleby2017] [@li2015].
 That contrasts with the results for tropical forests where forests there provide a cooling impact due to changes in both albedo and evapotranspiration [@li2015].
-Thus, the manner in which these various and oftentimes competing concerns are addressed will have significant impacts on the global community.
 

### 2023-05-31 paper.Rmd
Update paper.Rmd

@@
 In particular, boreal forests appear to provide the smallest net benefits, as the carbon sequestrations benefits of conifer forests are mitigated by the reduction in albedo in winter [@mykleby2017] [@li2015].
-That contrasts with the results for tropical forests where forests there provide a cooling impact due to changes in both albedo and evapotranspiration [@li2015].
+Moreover, coniferous forests have been shown to increase soil organic carbon less than broadleaf forests [@guo2021].
+With tropical forests, a strong positive result is derived from afforestation, as the carbon sequestration benefits is complemented by a cooling impact due to changes in both albedo and evapotranspiration, and with temperate forests, the impact of afforestation is shown to be positive, but less strongly so, due to a winter warming effect [@li2015].
 

### 2023-05-31 paper.Rmd
Update paper.Rmd

--- 2023-05-31 paper.Rmd
@@
 In particular, boreal forests appear to provide the smallest net benefits, as the carbon sequestrations benefits of conifer forests are mitigated by the reduction in albedo in winter [@mykleby2017] [@li2015].
-That contrasts with the results for tropical forests where forests there provide a cooling impact due to changes in both albedo and evapotranspiration [@li2015].
+Moreover, coniferous forests have been shown to increase soil organic carbon less than broadleaf forests [@guo2021].
+With tropical forests, a strong positive result is derived from afforestation, as the carbon sequestration benefit is complemented by a cooling impact due to changes in both albedo and evapotranspiration, and with temperate forests, the impact of afforestation is also shown to be positive, albeit less strongly so due to a winter warming effect [@li2015].
 

### 2023-05-31 paper.Rmd
Update paper.Rmd

@@
 
+These are not the only concerns that need to be considered by policymakers planning afforestation activities, as for example, there are also risks of impermance due to weak governance.
+The African Forest Landscape Restoration Initiative (AFR100) aims to afforest 100 million hectares by 2030.
+However, many of the commitments come from nations that are unstable or border unstable nations.
+This creates concern about how successful any investment may be.
+
 There is growing attention being placed globally to reforestation efforts, which is highligted by the UN's declaration of the years from 2021 through 2030 as the UN Decade of Ecosystem Resotration [@sewell2020].

### 2023-05-31 paper.Rmd
Update paper.Rmd

@@
 These are not the only concerns that need to be considered by policymakers planning afforestation activities, as for example, there are also risks of impermance due to weak governance.
-The African Forest Landscape Restoration Initiative (AFR100) aims to afforest 100 million hectares by 2030.
+The African Forest Landscape Restoration Initiative (AFR100) aims to afforest 100 million hectares by 2030 [@auda2023].
 However, many of the commitments come from nations that are unstable or border unstable nations.
-This creates concern about how successful any investment may be.
+This creates concern about how long-lasting any investment may be [@doelman2018].
 

### 2023-05-31 paper.Rmd
Update paper.Rmd

--- 2023-05-31 paper.Rmd
@@
 
+South America and Sub-Saharan Africa are two regions that show much potential for afforestation.
+It is estimated that these two regions hold at least 50% of the potential global gain while regions such as Northern Africa and the Middle East show little promise as growth rates are quite low [@doelman2020].
+Thus, it is promising that the African Forest Landscape Restoration Initiative (AFR100) aims to afforest 100 million hectares by 2030 with roughly 70 percent of the committments coming from Sub-Saharan nations [@auda2023].
+However, while many of these nations may be well-suited in big-physical terms for afforestation, they also may be politically unstable or border regions of instability.
+This creates concern about how long-lasting any investment may be.
+Other nations that have a big potential for gains from afforestation face different challenges.
+India has a great potential to increase its carbon sequestration through afforestation, but due to its masssive population and food security worries, its NDC does not entertain the idea of agricultural land being repurposed and instead primarily considers wastelands that may not be well suited for afforestation [@amjath2019].
+These are not the only concerns that need to be considered by policymakers planning afforestation activities, as for example, oelman2020].
+
 There is growing attention being placed globally to reforestation efforts, which is highligted by the UN's declaration of the years from 2021 through 2030 as the UN Decade of Ecosystem Resotration [@sewell2020].

### 2023-05-31 paper.Rmd
Update paper.Rmd

@@
 # Introduction
-Forests play an important role in maintaining and growing the land-based carbon sink, currently sequestering about 30% of GHG emmissions [@roe2019].
+In 1995, there was 4.2 billion hectares of forestland globally,which covered roughly 32.8% of the land [@Humpenöder2014].
+These forests play an important role in maintaining and growing the land-based carbon sink, currently sequestering about 30% of GHG emmissions [@roe2019].
 With increasingly negative climate impacts, the pressure to enhance the terrestrial sink is growing, and one of the tools to achieve this is through afforestation.
@@
 Thus, the manner in which these various and oftentimes competing concerns are addressed will have significant impacts on the global community.
+There is growing attention being placed globally to afforestation efforts, which is highligted by the UN's declaration of the years from 2021 through 2030 as the UN Decade of Ecosystem Resotration [@sewell2020].
 
@@
 Thus, it is promising that the African Forest Landscape Restoration Initiative (AFR100) aims to afforest 100 million hectares by 2030 with roughly 70 percent of the committments coming from Sub-Saharan nations [@auda2023].
-However, while many of these nations may be well-suited in big-physical terms for afforestation, they also may be politically unstable or border regions of instability.
-This creates concern about how long-lasting any investment may be.
+While many of these nations may be well-suited in big-physical terms for afforestation, there is concern about how long-lasting any investment may be due to proximate zones of political instability.
 Other nations that have a big potential for gains from afforestation face different challenges.
-India has a great potential to increase its carbon sequestration through afforestation, but due to its masssive population and food security worries, its NDC does not entertain the idea of agricultural land being repurposed and instead primarily considers wastelands that may not be well suited for afforestation [@amjath2019].
+India has a great potential to increase its carbon sequestration through afforestation, but due to its masssive population and food security worries, its NDC does not entertain the idea of agricultural land being repurposed and instead primarily considers targeting wastelands that may not be well suited for afforestation [@amjath2019].
+
 These are not the only concerns that need to be considered by policymakers planning afforestation activities, as for example, oelman2020].
-
-There is growing attention being placed globally to reforestation efforts, which is highligted by the UN's declaration of the years from 2021 through 2030 as the UN Decade of Ecosystem Resotration [@sewell2020].
 

### 2023-05-31 paper.Rmd
Update paper.Rmd

@@
 # Introduction
-In 1995, there was 4.2 billion hectares of forestland globally,which covered roughly 32.8% of the land [@Humpenöder2014].
+In 2023, there are 4.1 billion hectares of forestland globally,which covered roughly 31.4% of the land [@fao2023].
 These forests play an important role in maintaining and growing the land-based carbon sink, currently sequestering about 30% of GHG emmissions [@roe2019].
@@
 It is estimated that these two regions hold at least 50% of the potential global gain while regions such as Northern Africa and the Middle East show little promise as growth rates are quite low [@doelman2020].
-Thus, it is promising that the African Forest Landscape Restoration Initiative (AFR100) aims to afforest 100 million hectares by 2030 with roughly 70 percent of the committments coming from Sub-Saharan nations [@auda2023].
-While many of these nations may be well-suited in big-physical terms for afforestation, there is concern about how long-lasting any investment may be due to proximate zones of political instability.
+Thus, it is promising that the African Forest Landscape Restoration Initiative (AFR100) aims to afforest 100 million hectares by 2030 with roughly 70% of the committments coming from Sub-Saharan nations [@auda2023].
+While many of these nations may be well-suited in biophysical terms for afforestation, there is concern about how long-lasting any investment may be due to proximate zones of political instability.
 Other nations that have a big potential for gains from afforestation face different challenges.
@@
 These are not the only concerns that need to be considered by policymakers planning afforestation activities, as for example, oelman2020].
+
+[@Humpenöder2014].
 

### 2023-05-31 paper.Rmd
Update paper.Rmd

@@
 # Introduction
-In 2023, there are 4.1 billion hectares of forestland globally,which covered roughly 31.4% of the land [@fao2023].
+In 2020, there were 4.1 billion hectares of forestland globally that cover roughly 31.4% of the land [@fao2023].
 These forests play an important role in maintaining and growing the land-based carbon sink, currently sequestering about 30% of GHG emmissions [@roe2019].
-With increasingly negative climate impacts, the pressure to enhance the terrestrial sink is growing, and one of the tools to achieve this is through afforestation.
-In fact, many of the Nationally Determined Contributions (NDCs) countries have submitted that detail their plans to keep warming under the 2C threshold rely to a significant degree on afforestation projects [@roe2019].
-This is because of forests proven ability to sequester carbon not only the stems of trees [@davis2022] but their soils as well [@guo2021].
+With increasingly negative climate impacts, the pressure to enhance the terrestrial sink is growing, and one of the tools to achieve this is afforestation.
+There is growing attention being placed globally to afforestation efforts, which is highligted by the UN's declaration of the years from 2021 through 2030 as the UN Decade of Ecosystem Resotration [@sewell2020], and many of the Nationally Determined Contributions (NDCs) countries have submitted, which detail their plans to keep warming under the 2C threshold, rely to a significant degree on afforestation projects [@roe2019].
+This is because of forests proven ability to sequester carbon not only in the stems of trees [@davis2022] but in their soils as well [@guo2021].
 At the same time, forests are continually threatened by issues such as rising incomes and food prices and population growth that put pressure on policymakers to allow greater conversion of forestland into farmland.
@@
 Thus, the manner in which these various and oftentimes competing concerns are addressed will have significant impacts on the global community.
-There is growing attention being placed globally to afforestation efforts, which is highligted by the UN's declaration of the years from 2021 through 2030 as the UN Decade of Ecosystem Resotration [@sewell2020].
 

### 2023-05-31 paper.Rmd
Update paper.Rmd

@@
 
-The global gains that can be realized from afforestion, however, are uncertain, with estimates ranging from 42 gigatonnes [@veldman2019] to 205 gigatonnes [@bastin2019], as afforestation appears to not provide the same benefits in all regions of the world.
+The global gains that can be realized from afforestion, however, are uncertain, with broad estimates of cumulative sequestration ranging from 42 gigatonnes to 700 gigatonnes by the end of the century [@veldman2019]  [@bastin201]  [@Humpenöder2014].
+In addition, the amount of land conversion that would be required also varies greatly, with estimates ranging from 0.3 billion hectares to 2.8 billion hectares [@veldman2019]  [@bastin201]  [@Humpenöder2014].
+In many cases, these results suggest massive losses in pasture and cropland.
+
+Moreover, afforestation appears to not provide the same benefits in all regions of the world.
 In particular, boreal forests appear to provide the smallest net benefits, as the carbon sequestrations benefits of conifer forests are mitigated by the reduction in albedo in winter [@mykleby2017] [@li2015].
@@
 
-[@Humpenöder2014].
+.
 

### 2023-05-31 paper.Rmd
Update paper.Rmd

@@
 This is because of forests proven ability to sequester carbon not only in the stems of trees [@davis2022] but in their soils as well [@guo2021].
+
+The global gains that can be realized from afforestion, however, are uncertain, with broad estimates of cumulative sequestration ranging from 42 gigatonnes to 700 gigatonnes by the end of the century [@veldman2019]  [@bastin201]  [@Humpenöder2014]  [@sathaye2006].
+In addition, the amount of land conversion that would be required also varies greatly, with estimates ranging from 0.3 billion hectares to 2.8 billion hectares [@veldman2019]  [@bastin201]  [@Humpenöder2014].
+In many cases, these results suggest massive losses in pasture and cropland.
 At the same time, forests are continually threatened by issues such as rising incomes and food prices and population growth that put pressure on policymakers to allow greater conversion of forestland into farmland.
 Previous research though has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017].
-Thus, the manner in which these various and oftentimes competing concerns are addressed will have significant impacts on the global community.
-
-The global gains that can be realized from afforestion, however, are uncertain, with broad estimates of cumulative sequestration ranging from 42 gigatonnes to 700 gigatonnes by the end of the century [@veldman2019]  [@bastin201]  [@Humpenöder2014].
-In addition, the amount of land conversion that would be required also varies greatly, with estimates ranging from 0.3 billion hectares to 2.8 billion hectares [@veldman2019]  [@bastin201]  [@Humpenöder2014].
-In many cases, these results suggest massive losses in pasture and cropland.
+Still, the manner in which these various and oftentimes competing concerns are addressed will have significant impacts on the global community.
 

### 2023-05-31 paper.Rmd
Update paper.Rmd

@@
 
-The global gains that can be realized from afforestion, however, are uncertain, with broad estimates of cumulative sequestration ranging from 42 gigatonnes to 700 gigatonnes by the end of the century [@veldman2019]  [@bastin201]  [@Humpenöder2014]  [@sathaye2006].
-In addition, the amount of land conversion that would be required also varies greatly, with estimates ranging from 0.3 billion hectares to 2.8 billion hectares [@veldman2019]  [@bastin201]  [@Humpenöder2014].
+The global gains that can be realized from afforestion, however, are uncertain, with broad estimates of cumulative sequestration ranging from 42 gigatonnes to 700 gigatonnes by the end of the century [@veldman2019]  [@bastin201]  [@Humpenöder2014]  [@sathaye2006]  [@doelman2020].
+In addition, the amount of land conversion that would be required also varies greatly, with estimates ranging from 0.3 billion hectares to 2.8 billion hectares [@veldman2019]  [@bastin201]  [@Humpenöder2014]  [@doelman2020].
 In many cases, these results suggest massive losses in pasture and cropland.
-At the same time, forests are continually threatened by issues such as rising incomes and food prices and population growth that put pressure on policymakers to allow greater conversion of forestland into farmland.
+This puts into question the feasibility of attaining these results, as forests are continually threatened by issues such as rising incomes and food prices and population growth that put pressure on policymakers to allow greater conversion of forestland into farmland.
 Previous research though has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017].
@@
 South America and Sub-Saharan Africa are two regions that show much potential for afforestation.
-It is estimated that these two regions hold at least 50% of the potential global gain while regions such as Northern Africa and the Middle East show little promise as growth rates are quite low [@doelman2020].
+It is estimated that these two regions hold at least 50% of the potential global gain while regions such as Northern Africa and the Middle East show little promise as forest growth rates are quite low [@doelman2020].
 Thus, it is promising that the African Forest Landscape Restoration Initiative (AFR100) aims to afforest 100 million hectares by 2030 with roughly 70% of the committments coming from Sub-Saharan nations [@auda2023].
 While many of these nations may be well-suited in biophysical terms for afforestation, there is concern about how long-lasting any investment may be due to proximate zones of political instability.
-Other nations that have a big potential for gains from afforestation face different challenges.
-India has a great potential to increase its carbon sequestration through afforestation, but due to its masssive population and food security worries, its NDC does not entertain the idea of agricultural land being repurposed and instead primarily considers targeting wastelands that may not be well suited for afforestation [@amjath2019].
-
-These are not the only concerns that need to be considered by policymakers planning afforestation activities, as for example, oelman2020].
-
-.
+There are other nations that have a big potential for gains from afforestation, but they also face challenges.
+India, for example, has a great potential to increase its carbon sequestration through afforestation, but due to its masssive population and food security worries, its NDC does not entertain the idea of agricultural land being repurposed and instead primarily considers targeting wastelands that may not be well suited for afforestation [@amjath2019].
 

### 2023-06-01 paper.Rmd
Update paper.Rmd

@@
 In particular, boreal forests appear to provide the smallest net benefits, as the carbon sequestrations benefits of conifer forests are mitigated by the reduction in albedo in winter [@mykleby2017] [@li2015].
-Moreover, coniferous forests have been shown to increase soil organic carbon less than broadleaf forests [@guo2021].
+Coniferous forests have also been shown to increase soil organic carbon less than broadleaf forests [@guo2021].
 With tropical forests, a strong positive result is derived from afforestation, as the carbon sequestration benefit is complemented by a cooling impact due to changes in both albedo and evapotranspiration, and with temperate forests, the impact of afforestation is also shown to be positive, albeit less strongly so due to a winter warming effect [@li2015].
@@
 It is estimated that these two regions hold at least 50% of the potential global gain while regions such as Northern Africa and the Middle East show little promise as forest growth rates are quite low [@doelman2020].
-Thus, it is promising that the African Forest Landscape Restoration Initiative (AFR100) aims to afforest 100 million hectares by 2030 with roughly 70% of the committments coming from Sub-Saharan nations [@auda2023].
+Thus, it is promising that the African Forest Landscape Restoration Initiative (AFR100), which aims to afforest 100 million hectares by 2030, has roughly 70% of its committments coming from Sub-Saharan nations [@auda2023].
 While many of these nations may be well-suited in biophysical terms for afforestation, there is concern about how long-lasting any investment may be due to proximate zones of political instability.

### 2023-06-01 paper.Rmd
Update paper.Rmd

@@
 
-This project contributes to the literature by melding together two widely accepted global models, the Global Trade Analysis Project (GTAP) computable general equilibrium model and the Global Timber Model (GTM), to better assess the outcomes of several likely scenarios.
+This project contributes to the literature by melding together two widely accepted global models, the Global Trade Analysis Project (GTAP) computable general equilibrium model [@aguilar2019] and the Global Timber Model (GTM), to better assess the outcomes of several likely scenarios.
+The GTAP model facilitates a clear analysis of the impact of various shocks on things like agricultural prices and trade volumes, but the standard model does not well account for land-use changes, and the GTAP Land Use and Land Cover (LULC) database has limitations due to its infrequent updates.
+To get around these limitations, the GTM is quite helpful, as it is a dynamic optimization model that can examine the impact of afforestation or avoided deforestation on the carbon sink.
+By joining the two models and passing outputs back and forth between the models, any limitations of the models can be minimized and more robust results, therefore, attained.
 

### 2023-06-01 paper.Rmd
Update paper.Rmd

@@
 
-This project contributes to the literature by melding together two widely accepted global models, the Global Trade Analysis Project (GTAP) computable general equilibrium model [@aguilar2019] and the Global Timber Model (GTM), to better assess the outcomes of several likely scenarios.
+This project contributes to the literature by melding together two widely accepted global models, the Global Trade Analysis Project (GTAP) computable general equilibrium model [@aguilar2022] and the Global Timber Model (GTM) [@sohngen2001], to better assess the outcomes of several likely scenarios.
 The GTAP model facilitates a clear analysis of the impact of various shocks on things like agricultural prices and trade volumes, but the standard model does not well account for land-use changes, and the GTAP Land Use and Land Cover (LULC) database has limitations due to its infrequent updates.

### 2023-06-01 paper.Rmd
Update paper.Rmd

--- 2023-06-01 paper.Rmd
@@
 
-This project contributes to the literature by melding together two widely accepted global models, the Global Trade Analysis Project (GTAP) computable general equilibrium model and the Global Timber Model (GTM), to better assess the outcomes of several likely scenarios.
+This project will assess the impact of various likely scenarios by melding together two widely accepted global models, the Global Trade Analysis Project (GTAP) computable general equilibrium model [@aguilar2022] and the Global Timber Model (GTM) [@sohngen2001].
+The GTAP model has traditionally been used to analyze the impact of various shocks on things like agricultural prices and trade volumes.
+However, the standard model does not well account for land-use changes, and the GTAP Land Use and Land Cover (LULC) database has limitations due to its infrequent updates.
+The GTM is quite helpful in getting around these issues, as it is a dynamic optimization model that can examine the impact of afforestation or avoided deforestation on the carbon sink.
+By joining the two models and passing outputs back and forth between the models, any limitations of either model can be minimized and more robust results, therefore, can be attained.
 

### 2023-06-01 paper.Rmd
Update paper.Rmd

@@
 Scenarios
-1) Full attainment of UN strategic plan for forests 2030 (UNFF):  This says that all deforestation will halt and there will be a 3% (120 million hectare) increase in forest area worldwide by 2030 [@sewell2020].
-2) Full attainment of AFR100:  This says that there will be a 100 million hectare increase in forest area just in Sub-Saharan Africa by 2030 [@sewell2020].
-3) Full attainment of goals of Bonn Challenge and New York Declaration on Forests:  This says that all deforestation will halt and there will be a 350 million hectare increase in forest area worldwide by 2030 [@sewell2020].
+1) Full attainment of AFR100 pledges as part of Bonn Challenge:  This says that there will be a 127.8 million hectare increase in forest area just in Sub-Saharan Africa by 2030 [@bonn2023].
+To see the list of countries and the hectares pledged, please view the Excel document "Bonn Challenge-Africa".
+2) Partial attainment of AFR100 pledges as part of Bonn Challenge:  This scenario analyzes full attainment for nations with no internal wars and no current wars in neighboring lands and 0% attainment in all other regions.
+That results in a 127.8 million hectare increase in forest area by 2030 [@bonn2023].
+3) Full attainment of UN strategic plan for forests 2030 (UNFF):  This says that all deforestation will halt and there will be a 3% (120 million hectare) increase in forest area worldwide by 2030 [@sewell2020].
+4) 
+5) Full attainment of goals of Bonn Challenge and New York Declaration on Forests:  This says that all deforestation will halt and there will be a 350 million hectare increase in forest area worldwide by 2030 [@sewell2020].

### 2023-06-01 paper.Rmd
Update paper.Rmd

@@
 Scenarios
-1) Full attainment of AFR100 pledges as part of Bonn Challenge:  This says that there will be a 127.8 million hectare increase in forest area just in Sub-Saharan Africa by 2030 [@bonn2023].
-To see the list of countries and the hectares pledged, please view the Excel document "Bonn Challenge-Africa".
-2) Partial attainment of AFR100 pledges as part of Bonn Challenge:  This scenario analyzes full attainment for nations with no internal wars and no current wars in neighboring lands and 0% attainment in all other regions.
-That results in a 127.8 million hectare increase in forest area by 2030 [@bonn2023].
+1) Full attainment of AFR100 pledges as part of Bonn Challenge:  This says that there will be a 127.81 million hectare increase in forest area just in Sub-Saharan Africa by 2030 [@bonn2023].
+To see the list of countries and the hectares pledged, please view Scenario 1 of the Excel document "Bonn Challenge-Africa".
+2) Partial attainment of AFR100 pledges as part of Bonn Challenge:  This scenario analyzes full attainment for nations with no internal conflicts and 0% attainment in all other regions.
+That results in a 54.95 million hectare increase in forest area by 2030 [@bonn2023]  [@acled2023].
+To see the list of countries and the hectares pledged, please view Scenario 2 of the Excel document "Bonn Challenge-Africa".
 3) Full attainment of UN strategic plan for forests 2030 (UNFF):  This says that all deforestation will halt and there will be a 3% (120 million hectare) increase in forest area worldwide by 2030 [@sewell2020].

### 2023-06-01 paper.Rmd
Update paper.Rmd

@@
 To see the list of countries and the hectares pledged, please view Scenario 1 of the Excel document "Bonn Challenge-Africa".
-2) Partial attainment of AFR100 pledges as part of Bonn Challenge:  This scenario analyzes full attainment for nations with no internal conflicts and 0% attainment in all other regions.
-That results in a 54.95 million hectare increase in forest area by 2030 [@bonn2023]  [@acled2023].
+2) Partial attainment of AFR100 pledges as part of Bonn Challenge:  This scenario analyzes full attainment for nations with conflict severity level less than or equal to limited and 0% attainment for regions with a higher conflict severity rating.
+Using this measure, there would only be a 54.95 million hectare increase in forest area by 2030 [@bonn2023]  [@acled2023].
 To see the list of countries and the hectares pledged, please view Scenario 2 of the Excel document "Bonn Challenge-Africa".

### 2023-06-01 paper.Rmd
Update paper.Rmd

@@
 To see the list of countries and the hectares pledged, please view Scenario 2 of the Excel document "Bonn Challenge-Africa".
-3) Full attainment of UN strategic plan for forests 2030 (UNFF):  This says that all deforestation will halt and there will be a 3% (120 million hectare) increase in forest area worldwide by 2030 [@sewell2020].
-4) 
-5) Full attainment of goals of Bonn Challenge and New York Declaration on Forests:  This says that all deforestation will halt and there will be a 350 million hectare increase in forest area worldwide by 2030 [@sewell2020].
+3) Full attainment of Latin America pledges as part of Bonn Challenge: This says that there will be a 33.64 million hectare increase in forest area by 2030 [@bonn2023].
+To see the list of countries and the hectares pledged, please view the Latin America tab of the Excel document "Bonn Challenge-Other Regions".
+4) Full attainment of tropical nations' pledges as part of Bonn Challenge: This says that there will be a 182.30 million hectare increase in forest area by 2030 [@bonn2023].
+To see the list of countries and the hectares pledged, please view the Tropical Forest tab of the Excel document "Bonn Challenge-Other Regions".
+Nations that lie in both tropical and sub-tropical zones and taken to be tropical.
+5) Full attainment of UN strategic plan for forests 2030 (UNFF): This says that all deforestation will halt and there will be a 3% (~120 million hectare) increase in forest area worldwide by 2030 [@sewell2020].
+This will be modeled as a uniform 3% increase in forested area across all nations.
+6) Expansion of UN strategic plan for forests: This says that all deforestation will halt and there will be a 15% (~600 million hectare) increase in forest area worldwide by 2030 [@sewell2020].
+This will be modeled as a uniform 15% increase in forested area across all nations.

### 2023-06-12 paper.Rmd
Update paper.Rmd

@@
 title: "Implications of afforestation policies for global markets"
+date: "`r format(Sys.time(), '%d, %B %Y')`"
 author: "Eric Davis, Maros Ivanic, Brent Sohngen"

### 2023-06-12 paper.Rmd
Update paper.Rmd

--- 2023-06-01 paper.Rmd
@@
 title: "Implications of afforestation policies for global markets"
+date: "`r format(Sys.time(), '%B %d, %Y')`"
 author: "Eric Davis, Maros Ivanic, Brent Sohngen"

### 2023-06-23 paper.Rmd
A few fixes to Eri's first commit'

--- 2023-06-01 paper.Rmd
@@
 bibliography: references.bib
+header-includes:
+    - \usepackage{setspace}\doublespacing
 abstract: "The projected growth of the population and incomes as implied by various shared socioeconomic pathways is expected to create  pressures on converting forests into farmland as food prices increase.
@@
 
-\doublespacing
+# Introduction
 
-# Introduction
 In 2020, there were 4.1 billion hectares of forestland globally that cover roughly 31.4% of the land [@fao2023].
@@
 
-Scenarios
+# Scenarios
+
 1) Full attainment of AFR100 pledges as part of Bonn Challenge: This says that there will be a 127.81 million hectare increase in forest area just in Sub-Saharan Africa by 2030 [@bonn2023].
@@
 This will be modeled as a uniform 15% increase in forested area across all nations.
+
+# References

### 2023-06-24 paper.Rmd
Update paper.Rmd

@@
 In 2020, there were 4.1 billion hectares of forestland globally that cover roughly 31.4% of the land [@fao2023].
-These forests play an important role in maintaining and growing the land-based carbon sink, currently sequestering about 30% of GHG emmissions [@roe2019].
+These forests play an important role in maintaining and growing the land-based carbon sink, currently sequestering about 30% of GHG emissions [@roe2019].
 With increasingly negative climate impacts, the pressure to enhance the terrestrial sink is growing, and one of the tools to achieve this is afforestation.
-There is growing attention being placed globally to afforestation efforts, which is highligted by the UN's declaration of the years from 2021 through 2030 as the UN Decade of Ecosystem Resotration [@sewell2020], and many of the Nationally Determined Contributions (NDCs) countries have submitted, which detail their plans to keep warming under the 2C threshold, rely to a significant degree on afforestation projects [@roe2019].
+There is growing attention being placed globally to afforestation efforts, which is highlighted by the UN's declaration of the years from 2021 through 2030 as the UN Decade of Ecosystem Restoration [@sewell2020], and many of the Nationally Determined Contributions (NDCs) countries have submitted, which detail their plans to keep warming under the 2C threshold, rely to a significant degree on afforestation projects [@roe2019].
 This is because of forests proven ability to sequester carbon not only in the stems of trees [@davis2022] but in their soils as well [@guo2021].
 
-The global gains that can be realized from afforestion, however, are uncertain, with broad estimates of cumulative sequestration ranging from 42 gigatonnes to 700 gigatonnes by the end of the century [@veldman2019]  [@bastin201]  [@Humpenöder2014]  [@sathaye2006]  [@doelman2020].
+The global gains that can be realized from afforestation, however, are uncertain, with broad estimates of cumulative sequestration ranging from 42 gigatonnes to 700 gigatonnes by the end of the century [@veldman2019]  [@bastin201]  [@Humpenöder2014]  [@sathaye2006]  [@doelman2020].
 In addition, the amount of land conversion that would be required also varies greatly, with estimates ranging from 0.3 billion hectares to 2.8 billion hectares [@veldman2019]  [@bastin201]  [@Humpenöder2014]  [@doelman2020].
@@
 
-Moreover, afforestation appears to not provide the same benefits in all regions of the world.
+Moreover, afforestation appears not to provide the same benefits in all regions of the world.
 In particular, boreal forests appear to provide the smallest net benefits, as the carbon sequestrations benefits of conifer forests are mitigated by the reduction in albedo in winter [@mykleby2017] [@li2015].
@@
 It is estimated that these two regions hold at least 50% of the potential global gain while regions such as Northern Africa and the Middle East show little promise as forest growth rates are quite low [@doelman2020].
-Thus, it is promising that the African Forest Landscape Restoration Initiative (AFR100), which aims to afforest 100 million hectares by 2030, has roughly 70% of its committments coming from Sub-Saharan nations [@auda2023].
+Thus, it is promising that the African Forest Landscape Restoration Initiative (AFR100), which aims to afforest 100 million hectares by 2030, has roughly 70% of its commitments coming from Sub-Saharan nations [@auda2023].
 While many of these nations may be well-suited in biophysical terms for afforestation, there is concern about how long-lasting any investment may be due to proximate zones of political instability.
 There are other nations that have a big potential for gains from afforestation, but they also face challenges.
-India, for example, has a great potential to increase its carbon sequestration through afforestation, but due to its masssive population and food security worries, its NDC does not entertain the idea of agricultural land being repurposed and instead primarily considers targeting wastelands that may not be well suited for afforestation [@amjath2019].
+India, for example, has a great potential to increase its carbon sequestration through afforestation, but due to its massive population and food security worries, its NDC does not entertain the idea of agricultural land being repurposed and instead primarily considers targeting wastelands that may not be well suited for afforestation [@amjath2019].
 

### 2023-06-24 paper.Rmd
Update paper.Rmd

@@
 abstract: "The projected growth of the population and incomes as implied by various shared socioeconomic pathways is expected to create  pressures on converting forests into farmland as food prices increase.
-Depending on the specific path taken, SSP1 through SSP5, and the continuation of the current trends in agricultural productivity growth, the market impacts are expected to vary but generally resulting in increases in food prices.
-Policies aimed at reversing the deforestation trend by afforesting crop land would result in severe negative market impacts such as  higher food prices, especially in low agriculutrla productivity/high population growht scenarios."
+Depending on the specific path taken, SSP1 through SSP5, and the continuation of the current trends in agricultural productivity growth, the market impacts are expected to vary but generally result in increases in food prices.
+Policies aimed at reversing the deforestation trend by afforesting cropland would result in severe negative market impacts such as  higher food prices, especially in low agricultural productivity/high population growth scenarios."
 ---
@@
 
+A transition from deforestation to reforestation, in general, induce the displacement of land use abroad despite the uncertain roles of international trade due to the heterogeneity of regions/countries (Meyfroidt, Rudel and Lambin, 2010; Lambin and Meyfroidt, 2011).
+DeFries et al.
+(2010) found a strong correlation between forest loss and exports of agricultural products.
+On the other hand, Turner et al.
+(2005) investigated the role of trade on forests in the context of the Free Trade Area of the Americas (FTAA) using the spatial partial-equilibrium Global Forest Products Model and found that FTAA has a positive impact on forest resources.
+
 # Scenarios

### 2023-06-24 paper.Rmd
Update paper.Rmd

@@
 
-A transition from deforestation to reforestation, in general, induce the displacement of land use abroad despite the uncertain roles of international trade due to the heterogeneity of regions/countries (Meyfroidt, Rudel and Lambin, 2010; Lambin and Meyfroidt, 2011).
-DeFries et al.
-(2010) found a strong correlation between forest loss and exports of agricultural products.
+A transition from deforestation to reforestation, in general, induces the displacement of land use abroad despite the uncertain roles of international trade due to the heterogeneity of regions/countries [@meyfroidt2010] [@lambin2011] DeFries et al.
+[@defries2010] found a strong correlation between forest loss and exports of agricultural products.
 On the other hand, Turner et al.
-(2005) investigated the role of trade on forests in the context of the Free Trade Area of the Americas (FTAA) using the spatial partial-equilibrium Global Forest Products Model and found that FTAA has a positive impact on forest resources.
+[@turner2005] investigated the role of trade on forests in the context of the Free Trade Area of the Americas (FTAA) using the spatial partial-equilibrium Global Forest Products Model and found that FTAA has a positive impact on forest resources.
 

### 2023-07-03 paper.Rmd
Update paper.Rmd

@@
 The GTM is quite helpful in getting around these issues, as it is a dynamic optimization model that can examine the impact of afforestation or avoided deforestation on the carbon sink.
-By joining the two models and passing outputs back and forth between the models, any limitations of either model can be minimized and more robust results, therefore, can be attained.
+By joining these two models and passing outputs back and forth between the models, any limitations of either model can be minimized and more robust results, therefore, can be attained.
 

### 2023-07-07 paper.Rmd
Update paper.Rmd

@@
 
+The Land Use and Land Cover Database within the GTAP framework builds upon the global land cover and land use data, incorporating forestry remote-sensing products (Baldos & Hertel, 2012).
+Using globally spatially explicit land use data, Global Trade Analysis Project Agro-Ecological Zone (GTAP-AEZ) was introduced in order to enhance the GTAP model by incorporating intra- and inter-regional land and land-based greenhouse emissions heterogeneity (Hertel et al.
+2008).
+It includes competition among different crops, grazing, and forest-based uses within agro-ecological zones (AEZs).
+
+The dynamic GTAP model, supported by the GTAP-AEZ database, considers general equilibrium impacts.
+Land market effects, identified as significant in driving results by Stevenson et al.
+(2013), are taken into account.
+The GTAP-AEZ database has been utilized to evaluate the impacts of crop intensification on land use, highlighting the importance of global research and development investment as a strategy to alleviate pressure on natural ecosystems.
+However, it is important to recognize that global aggregates may mask localized shifts that can have implications for ecologically significant areas (Byerlee, Stevenson & Villoria, 2014).
+
+Steinbuks and Hertel (2012) developed the FABLE model (Forestry, Agriculture, Biofuels, Land Use, and Environment) as a global partial equilibrium model to analyze optimal land use within a context of increasing demand for food, bioenergy, forest products, and non-provisioning ecosystem services, while also meeting greenhouse gas targets.
+The model encompasses nine sectors, including agriculture, livestock, food processing, biofuels, energy, forest, timber processing, and ecosystem services.
+While FABLE is a powerful tool for examining the optimal trajectory of various land uses under specific demand assumptions, it primarily focuses on specific sectors of the economy, neglecting general equilibrium effects and socio-economic or household welfare impacts (Steinbuks & Hertel, 2012; Hertel, 2017).
+
+The KLUM@GTAP framework integrates the Kleines Land Use Model (KLUM) with an extended version of GTAP called GTAP-EFL to evaluate the impacts of climate change on cropland allocation.
+GTAP-EFL separates energy factors from intermediate inputs and incorporates them into capital, while also considering CO2 emissions.
+KLUM, a global agricultural land-use model, connects the economy to global crop allocation to maximize producer returns under specific risk assumptions.
+KLUM@GTAP substitutes the land allocation mechanism within GTAP-EFL by utilizing regionally aggregated area changes in cropland determined by KLUM to update cropland shares in GTAP-EFL (Ronneberger et al., 2009).
+
 # Scenarios

### 2023-07-07 paper.Rmd
Update paper.Rmd

--- 2023-07-03 paper.Rmd
@@
 
+The Land Use and Land Cover Database within the GTAP framework builds upon the global land cover and land use data, incorporating forestry remote-sensing products [@baldos2012].
+Using globally spatially explicit land use data, Global Trade Analysis Project Agro-Ecological Zone (GTAP-AEZ) was introduced in order to enhance the GTAP model by incorporating intra- and inter-regional land and land-based greenhouse emissions heterogeneity [@hertel2008].
+It includes competition among different crops, grazing, and forest-based uses within agro-ecological zones (AEZs).
+
+The dynamic GTAP model, supported by the GTAP-AEZ database, considers general equilibrium impacts.
+Land market effects, identified as significant in driving results by Stevenson et al.
+[@stevenson2013], are taken into account.
+The GTAP-AEZ database has been utilized to evaluate the impacts of crop intensification on land use, highlighting the importance of global research and development investment as a strategy to alleviate pressure on natural ecosystems.
+However, it is important to recognize that global aggregates may mask localized shifts that can have implications for ecologically significant areas [@byerlee2014].
+
+Steinbuks and Hertel [@steinbuks2012] developed the FABLE model (Forestry, Agriculture, Biofuels, Land Use, and Environment) as a global partial equilibrium model to analyze optimal land use within a context of increasing demand for food, bioenergy, forest products, and non-provisioning ecosystem services, while also meeting greenhouse gas targets.
+The model encompasses nine sectors, including agriculture, livestock, food processing, biofuels, energy, forest, timber processing, and ecosystem services.
+While FABLE is a powerful tool for examining the optimal trajectory of various land uses under specific demand assumptions, it primarily focuses on specific sectors of the economy, neglecting general equilibrium effects and socio-economic or household welfare impacts [@steinbuks2012] [@hertel2017].
+
+The KLUM@GTAP framework integrates the Kleines Land Use Model (KLUM) with an extended version of GTAP called GTAP-EFL to evaluate the impacts of climate change on cropland allocation.
+GTAP-EFL separates energy factors from intermediate inputs and incorporates them into capital, while also considering CO2 emissions.
+KLUM, a global agricultural land-use model, connects the economy to global crop allocation to maximize producer returns under specific risk assumptions.
+KLUM@GTAP substitutes the land allocation mechanism within GTAP-EFL by utilizing regionally aggregated area changes in cropland determined by KLUM to update cropland shares in GTAP-EFL [@ronneberger2009].
+
 # Scenarios

### 2023-08-07 paper.Rmd
Update paper.Rmd

@@
 
-In 2020, there were 4.1 billion hectares of forestland globally that cover roughly 31.4% of the land [@fao2023].
-These forests play an important role in maintaining and growing the land-based carbon sink, currently sequestering about 30% of GHG emissions [@roe2019].
+In 2020, there were 4.1 billion hectares of forestland globally covering roughly 31.4% of the land [@fao2023].
+These forests play an important role in maintaining and growing the land-based carbon sink by currently sequestering about 30% of GHG emissions [@roe2019].
+This is because of forests proven ability to sequester carbon not only in the stems of trees [@davis2022] but in their soils as well [@guo2021].
 With increasingly negative climate impacts, the pressure to enhance the terrestrial sink is growing, and one of the tools to achieve this is afforestation.
-There is growing attention being placed globally to afforestation efforts, which is highlighted by the UN's declaration of the years from 2021 through 2030 as the UN Decade of Ecosystem Restoration [@sewell2020], and many of the Nationally Determined Contributions (NDCs) countries have submitted, which detail their plans to keep warming under the 2C threshold, rely to a significant degree on afforestation projects [@roe2019].
-This is because of forests proven ability to sequester carbon not only in the stems of trees [@davis2022] but in their soils as well [@guo2021].
+As such, there is growing attention being placed globally to afforestation efforts, which is highlighted by the UN's declaration of the years from 2021 through 2030 as the UN Decade of Ecosystem Restoration [@sewell2020].
+Many of the Nationally Determined Contributions (NDCs) countries have submitted, which detail their plans to keep warming under the 2C threshold, also rely to a significant degree on afforestation projects [@roe2019].
 
@@
 In addition, the amount of land conversion that would be required also varies greatly, with estimates ranging from 0.3 billion hectares to 2.8 billion hectares [@veldman2019]  [@bastin201]  [@Humpenöder2014]  [@doelman2020].
-In many cases, these results suggest massive losses in pasture and cropland.
-This puts into question the feasibility of attaining these results, as forests are continually threatened by issues such as rising incomes and food prices and population growth that put pressure on policymakers to allow greater conversion of forestland into farmland.
-Previous research though has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017].
+In many cases, these results suggest massive losses in the amount of pasture and cropland that would be available.
+This puts into question the feasibility of attaining these results, as almost 90 percent of deforestation worldwide is undertaken to expand the amount of agricultural land [@fao2021].
+There are many reasons for this.
+For example, as nations develop and disposable incomes rise, consumers often change to more meat-based diets.
+The population growth that is especially high, at present, in Africa also puts pressure on policymakers to allow greater conversion of forestland into farmland.
+Still, previous research has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017].
+
 Still, the manner in which these various and oftentimes competing concerns are addressed will have significant impacts on the global community.

### 2023-08-07 paper.Rmd
Update paper.Rmd

@@
 In 2020, there were 4.1 billion hectares of forestland globally covering roughly 31.4% of the land [@fao2023].
-These forests play an important role in maintaining and growing the land-based carbon sink by currently sequestering about 30% of GHG emissions [@roe2019].
+These forests play an important role in maintaining and growing the land-based carbon sink by  sequestering about 30% of GHG emissions currently [@roe2019].
 This is because of forests proven ability to sequester carbon not only in the stems of trees [@davis2022] but in their soils as well [@guo2021].
@@
 For example, as nations develop and disposable incomes rise, consumers often change to more meat-based diets.
-The population growth that is especially high, at present, in Africa also puts pressure on policymakers to allow greater conversion of forestland into farmland.
+Population growth is also a driver and when it is especially high, which is the case at present throughout most of Africa, this also puts pressure on policymakers to allow greater conversion of forestland into farmland.
 Still, previous research has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017].

### 2023-08-07 paper.Rmd
Update paper.Rmd

@@
 Population growth is also a driver and when it is especially high, which is the case at present throughout most of Africa, this also puts pressure on policymakers to allow greater conversion of forestland into farmland.
+
+Land-use decisions in one countr also tend to have impacts that spill over into many others [@meyfroidt2010] [@lambin2011], and therefore, cumulative change needs to be assessed with global-scale models.
+Steinbuks and Hertel [@steinbuks2012] developed the FABLE model (Forestry, Agriculture, Biofuels, Land Use, and Environment) as a global partial equilibrium model to analyze optimal land use within a context of increasing demand for food, bioenergy, forest products, and non-provisioning ecosystem services, while also meeting greenhouse gas targets.
+The model encompasses nine sectors, including agriculture, livestock, food processing, biofuels, energy, forest, timber processing, and ecosystem services.
+While FABLE is a powerful tool for examining the optimal trajectory of various land uses under specific demand assumptions, it primarily focuses on specific sectors of the economy, neglecting general equilibrium effects and socio-economic or household welfare impacts [@steinbuks2012] [@hertel2017].
+There have been attempts to use the Global Trade Analysis Project (GTAP) computable general equilibrium model to address such questions, but the standard model does not well account for land-use changes.
+The Land Use and Land Cover Database within the GTAP framework, which incorporated forestry remote-sensing products, was thus created [@baldos2012].
+It was strengthened as part of the GTAP Agro-Ecological Zone (GTAP-AEZ) model, with the use of spatially explicit global land use data and through the incorporation of intra- and inter-regional land and land-based greenhouse emissions heterogeneity [@hertel2008].
+It also includes competition among different crops, grazing, and forest-based uses, but it has limitations due to its infrequent updates.
+The KLUM@GTAP framework integrates the Kleines Land Use Model (KLUM) with an extended yet static version of GTAP called GTAP-EFL to evaluate the impacts of climate change on cropland allocation.
+GTAP-EFL separates energy factors from intermediate inputs and incorporates them into capital, while also considering CO2 emissions.
+KLUM, a global agricultural land-use model, connects the economy to global crop allocation to maximize producer returns under specific risk assumptions.
+KLUM@GTAP substitutes the land allocation mechanism within GTAP-EFL by utilizing regionally aggregated area changes in cropland determined by KLUM to update cropland shares in GTAP-EFL [@ronneberger2009].
+
+More recent modeling attempts of land-use change have begun to move away from comparative static analysis.
+For example, the dynamic GTAP model, supported by the GTAP-AEZ database, considers land market effects, which were identified as significant in driving results by Stevenson et al.
+[@stevenson2013], and has been utilized to evaluate the impacts of crop intensification on land use.
+It, however, has been critiqued as it global aggregates may mask localized shifts that can have implications for ecologically significant areas [@byerlee2014].
+This project will attempt to get around these concerns and assess the impact of various likely afforestation scenarios by melding GTAP with the Global Timber Model (GTM) [@sohngen2001].
+The GTM is seen as being quite helpful in getting around these issues, as it is a dynamic optimization model that can examine the impact of afforestation or avoided deforestation on the carbon sink.
+By joining these two models and passing outputs back and forth between the models, any limitations of either model can be minimized and more robust results, therefore, can be attained.
+
 Still, previous research has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017].
 
-Still, the manner in which these various and oftentimes competing concerns are addressed will have significant impacts on the global community.
-
-Moreover, afforestation appears not to provide the same benefits in all regions of the world.
-In particular, boreal forests appear to provide the smallest net benefits, as the carbon sequestrations benefits of conifer forests are mitigated by the reduction in albedo in winter [@mykleby2017] [@li2015].
+Faced with competing land-use demands, afforestation efforts need to be targeted to the regions of the world where they will provide the most benefits.
+Afforestation in boreal zones has been shown to provide the smallest net benefits, as the carbon sequestrations benefits of conifer forests are mitigated by the reduction in albedo in winter [@mykleby2017] [@li2015].
 Coniferous forests have also been shown to increase soil organic carbon less than broadleaf forests [@guo2021].
-With tropical forests, a strong positive result is derived from afforestation, as the carbon sequestration benefit is complemented by a cooling impact due to changes in both albedo and evapotranspiration, and with temperate forests, the impact of afforestation is also shown to be positive, albeit less strongly so due to a winter warming effect [@li2015].
+Conversely, tropical forests have been linked with a strong positive result from afforestation, as the carbon sequestration benefit is complemented by a cooling impact due to changes in both albedo and evapotranspiration, and with temperate forests, the impact of afforestation is also shown to be positive, albeit less strongly so due to a winter warming effect [@li2015].
 
@@
 India, for example, has a great potential to increase its carbon sequestration through afforestation, but due to its massive population and food security worries, its NDC does not entertain the idea of agricultural land being repurposed and instead primarily considers targeting wastelands that may not be well suited for afforestation [@amjath2019].
-
-This project will assess the impact of various likely scenarios by melding together two widely accepted global models, the Global Trade Analysis Project (GTAP) computable general equilibrium model [@aguilar2022] and the Global Timber Model (GTM) [@sohngen2001].
-The GTAP model has traditionally been used to analyze the impact of various shocks on things like agricultural prices and trade volumes.
-However, the standard model does not well account for land-use changes, and the GTAP Land Use and Land Cover (LULC) database has limitations due to its infrequent updates.
-The GTM is quite helpful in getting around these issues, as it is a dynamic optimization model that can examine the impact of afforestation or avoided deforestation on the carbon sink.
-By joining these two models and passing outputs back and forth between the models, any limitations of either model can be minimized and more robust results, therefore, can be attained.
-
-A transition from deforestation to reforestation, in general, induces the displacement of land use abroad despite the uncertain roles of international trade due to the heterogeneity of regions/countries [@meyfroidt2010] [@lambin2011] DeFries et al.
-[@defries2010] found a strong correlation between forest loss and exports of agricultural products.
-On the other hand, Turner et al.
-[@turner2005] investigated the role of trade on forests in the context of the Free Trade Area of the Americas (FTAA) using the spatial partial-equilibrium Global Forest Products Model and found that FTAA has a positive impact on forest resources.
-
-The Land Use and Land Cover Database within the GTAP framework builds upon the global land cover and land use data, incorporating forestry remote-sensing products [@baldos2012].
-Using globally spatially explicit land use data, Global Trade Analysis Project Agro-Ecological Zone (GTAP-AEZ) was introduced in order to enhance the GTAP model by incorporating intra- and inter-regional land and land-based greenhouse emissions heterogeneity [@hertel2008].
-It includes competition among different crops, grazing, and forest-based uses within agro-ecological zones (AEZs).
-
-The dynamic GTAP model, supported by the GTAP-AEZ database, considers general equilibrium impacts.
-Land market effects, identified as significant in driving results by Stevenson et al.
-[@stevenson2013], are taken into account.
-The GTAP-AEZ database has been utilized to evaluate the impacts of crop intensification on land use, highlighting the importance of global research and development investment as a strategy to alleviate pressure on natural ecosystems.
-However, it is important to recognize that global aggregates may mask localized shifts that can have implications for ecologically significant areas [@byerlee2014].
-
-Steinbuks and Hertel [@steinbuks2012] developed the FABLE model (Forestry, Agriculture, Biofuels, Land Use, and Environment) as a global partial equilibrium model to analyze optimal land use within a context of increasing demand for food, bioenergy, forest products, and non-provisioning ecosystem services, while also meeting greenhouse gas targets.
-The model encompasses nine sectors, including agriculture, livestock, food processing, biofuels, energy, forest, timber processing, and ecosystem services.
-While FABLE is a powerful tool for examining the optimal trajectory of various land uses under specific demand assumptions, it primarily focuses on specific sectors of the economy, neglecting general equilibrium effects and socio-economic or household welfare impacts [@steinbuks2012] [@hertel2017].
-
-The KLUM@GTAP framework integrates the Kleines Land Use Model (KLUM) with an extended version of GTAP called GTAP-EFL to evaluate the impacts of climate change on cropland allocation.
-GTAP-EFL separates energy factors from intermediate inputs and incorporates them into capital, while also considering CO2 emissions.
-KLUM, a global agricultural land-use model, connects the economy to global crop allocation to maximize producer returns under specific risk assumptions.
-KLUM@GTAP substitutes the land allocation mechanism within GTAP-EFL by utilizing regionally aggregated area changes in cropland determined by KLUM to update cropland shares in GTAP-EFL [@ronneberger2009].
 
@@
 
+DeFries et al.
+[@defries2010] found a strong correlation between forest loss and exports of agricultural products.
+On the other hand, Turner et al.
+[@turner2005] investigated the role of trade on forests in the context of the Free Trade Area of the Americas (FTAA) using the spatial partial-equilibrium Global Forest Products Model and found that FTAA has a positive impact on forest resources.
 # References

### 2023-08-07 paper.Rmd
Update paper.Rmd

@@
 Population growth is also a driver and when it is especially high, which is the case at present throughout most of Africa, this also puts pressure on policymakers to allow greater conversion of forestland into farmland.
+Still, previous research has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017].
+However, as land-use decisions in one country also tend to have impacts that spillover into many others [@meyfroidt2010] [@lambin2011], analysis needs to be conducted at a global scale.
 
-Land-use decisions in one countr also tend to have impacts that spill over into many others [@meyfroidt2010] [@lambin2011], and therefore, cumulative change needs to be assessed with global-scale models.
 Steinbuks and Hertel [@steinbuks2012] developed the FABLE model (Forestry, Agriculture, Biofuels, Land Use, and Environment) as a global partial equilibrium model to analyze optimal land use within a context of increasing demand for food, bioenergy, forest products, and non-provisioning ecosystem services, while also meeting greenhouse gas targets.
@@
 While FABLE is a powerful tool for examining the optimal trajectory of various land uses under specific demand assumptions, it primarily focuses on specific sectors of the economy, neglecting general equilibrium effects and socio-economic or household welfare impacts [@steinbuks2012] [@hertel2017].
-There have been attempts to use the Global Trade Analysis Project (GTAP) computable general equilibrium model to address such questions, but the standard model does not well account for land-use changes.
-The Land Use and Land Cover Database within the GTAP framework, which incorporated forestry remote-sensing products, was thus created [@baldos2012].
-It was strengthened as part of the GTAP Agro-Ecological Zone (GTAP-AEZ) model, with the use of spatially explicit global land use data and through the incorporation of intra- and inter-regional land and land-based greenhouse emissions heterogeneity [@hertel2008].
-It also includes competition among different crops, grazing, and forest-based uses, but it has limitations due to its infrequent updates.
-The KLUM@GTAP framework integrates the Kleines Land Use Model (KLUM) with an extended yet static version of GTAP called GTAP-EFL to evaluate the impacts of climate change on cropland allocation.
+Other efforts have used the Global Trade Analysis Project (GTAP) computable general equilibrium model to address such questions, but its standard model does not well account for land-use changes.
+As a result, the Land Use and Land Cover Database within the GTAP framework, which incorporates forestry remote-sensing products, was created [@baldos2012].
+It was further strengthened as part of the GTAP Agro-Ecological Zone (GTAP-AEZ) model, with the use of spatially explicit global land use data and through the incorporation of intra- and inter-regional land and land-based greenhouse emissions heterogeneity [@hertel2008].
+While the GTAP-AEZ model benefits from the inclusion of competition among different crops, grazing, and forest-based uses, it has limitations, primarily due to its infrequent updates.
+The KLUM@GTAP framework integrated the Kleines Land Use Model (KLUM) with an extended yet static version of GTAP called GTAP-EFL to evaluate the impacts of climate change on cropland allocation.
 GTAP-EFL separates energy factors from intermediate inputs and incorporates them into capital, while also considering CO2 emissions.
@@
 By joining these two models and passing outputs back and forth between the models, any limitations of either model can be minimized and more robust results, therefore, can be attained.
-
-Still, previous research has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017].
 

### 2023-08-07 paper.Rmd
Update paper.Rmd

@@
 Still, previous research has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017].
-However, as land-use decisions in one country also tend to have impacts that spillover into many others [@meyfroidt2010] [@lambin2011], analysis needs to be conducted at a global scale.
+
+Faced with competing land-use demands, afforestation efforts need to be targeted to the regions of the world where they will provide the most benefits.
+Afforestation in boreal zones has been shown to provide the smallest net benefits, as the carbon sequestrations benefits of conifer forests are mitigated by the reduction in albedo in winter [@mykleby2017] [@li2015].
+Coniferous forests have also been shown to increase soil organic carbon less than broadleaf forests [@guo2021].
+Conversely, tropical forests have been linked with a strong positive result from afforestation, as the carbon sequestration benefit is complemented by a cooling impact due to changes in both albedo and evapotranspiration, and with temperate forests, the impact of afforestation is also shown to be positive, albeit less strongly so due to a winter warming effect [@li2015].
+
+South America and Sub-Saharan Africa are two regions that show much potential for afforestation.
+It is estimated that these two regions hold at least 50% of the potential global gain while regions such as Northern Africa and the Middle East show little promise as forest growth rates are quite low [@doelman2020].
+Thus, it is promising that the African Forest Landscape Restoration Initiative (AFR100), which aims to afforest 100 million hectares by 2030, has roughly 70% of its commitments coming from Sub-Saharan nations [@auda2023].
+While many of these nations may be well-suited in biophysical terms for afforestation, there is concern about how long-lasting any investment may be due to proximate zones of political instability.
+There are other nations that have a big potential for gains from afforestation, but they also face challenges.
+India, for example, has a great potential to increase its carbon sequestration through afforestation, but due to its massive population and food security worries, its NDC does not entertain the idea of agricultural land being repurposed and instead primarily considers targeting wastelands that may not be well suited for afforestation [@amjath2019].
+Moreover, as land-use decisions in one country also tend to have impacts that spillover into many others [@meyfroidt2010] [@lambin2011], analysis needs to be conducted at a global scale.
 
@@
 [@stevenson2013], and has been utilized to evaluate the impacts of crop intensification on land use.
-It, however, has been critiqued as it global aggregates may mask localized shifts that can have implications for ecologically significant areas [@byerlee2014].
-This project will attempt to get around these concerns and assess the impact of various likely afforestation scenarios by melding GTAP with the Global Timber Model (GTM) [@sohngen2001].
-The GTM is seen as being quite helpful in getting around these issues, as it is a dynamic optimization model that can examine the impact of afforestation or avoided deforestation on the carbon sink.
+It, however, has been critiqued, as global aggregates may mask localized shifts that can have implications for ecologically significant areas [@byerlee2014].
+This project will attempt to get around these concerns and assess the impact of various likely afforestation scenarios by melding GTAP with the Global Timber Model (GTM), which is a dynamic optimization model that can examine the impact of afforestation or avoided deforestation on the carbon sink [@sohngen2001].
 By joining these two models and passing outputs back and forth between the models, any limitations of either model can be minimized and more robust results, therefore, can be attained.
-
-Faced with competing land-use demands, afforestation efforts need to be targeted to the regions of the world where they will provide the most benefits.
-Afforestation in boreal zones has been shown to provide the smallest net benefits, as the carbon sequestrations benefits of conifer forests are mitigated by the reduction in albedo in winter [@mykleby2017] [@li2015].
-Coniferous forests have also been shown to increase soil organic carbon less than broadleaf forests [@guo2021].
-Conversely, tropical forests have been linked with a strong positive result from afforestation, as the carbon sequestration benefit is complemented by a cooling impact due to changes in both albedo and evapotranspiration, and with temperate forests, the impact of afforestation is also shown to be positive, albeit less strongly so due to a winter warming effect [@li2015].
-
-South America and Sub-Saharan Africa are two regions that show much potential for afforestation.
-It is estimated that these two regions hold at least 50% of the potential global gain while regions such as Northern Africa and the Middle East show little promise as forest growth rates are quite low [@doelman2020].
-Thus, it is promising that the African Forest Landscape Restoration Initiative (AFR100), which aims to afforest 100 million hectares by 2030, has roughly 70% of its commitments coming from Sub-Saharan nations [@auda2023].
-While many of these nations may be well-suited in biophysical terms for afforestation, there is concern about how long-lasting any investment may be due to proximate zones of political instability.
-There are other nations that have a big potential for gains from afforestation, but they also face challenges.
-India, for example, has a great potential to increase its carbon sequestration through afforestation, but due to its massive population and food security worries, its NDC does not entertain the idea of agricultural land being repurposed and instead primarily considers targeting wastelands that may not be well suited for afforestation [@amjath2019].
 

### 2024-01-22 paper.Rmd
Initial edits

@@
 author: "Eric Davis, Maros Ivanic, Brent Sohngen"
-output: bookdown::pdf_document2
+output: 
+  bookdown::pdf_document2:
+    keep_tex: TRUE
 toc: FALSE
@@
 abstract: "The projected growth of the population and incomes as implied by various shared socioeconomic pathways is expected to create  pressures on converting forests into farmland as food prices increase.
-Depending on the specific path taken, SSP1 through SSP5, and the continuation of the current trends in agricultural productivity growth, the market impacts are expected to vary but generally result in increases in food prices.
-Policies aimed at reversing the deforestation trend by afforesting cropland would result in severe negative market impacts such as  higher food prices, especially in low agricultural productivity/high population growth scenarios."
+Depending on the specific path taken, SSP1 through SSP5, and the continuation of the current trends in agricultural productivity growth, the market impacts are expected to vary but generally resulting in increases in food prices.
+Policies aimed at reversing the deforestation trend by afforesting crop land would result in severe negative market impacts such as  higher food prices, especially in low agriculutrla productivity/high population growht scenarios."
+editor_options: 
+  chunk_output_type: console
 ---
@@
 
-In 2020, there were 4.1 billion hectares of forestland globally covering roughly 31.4% of the land [@fao2023].
-These forests play an important role in maintaining and growing the land-based carbon sink by  sequestering about 30% of GHG emissions currently [@roe2019].
+In 2020, there were 4.1 billion hectares of forestland globally that cover roughly 31.4% of the land [@fao2023].
+These forests play an important role in maintaining and growing the land-based carbon sink, currently sequestering about 30% of GHG emmissions [@roe2019].
+With increasingly negative climate impacts, the pressure to enhance the terrestrial sink is growing, and one of the tools to achieve this is afforestation.
+There is growing attention being placed globally to afforestation efforts, which is highligted by the UN's declaration of the years from 2021 through 2030 as the UN Decade of Ecosystem Resotration [@sewell2020], and many of the Nationally Determined Contributions (NDCs) countries have submitted, which detail their plans to keep warming under the 2C threshold, rely to a significant degree on afforestation projects [@roe2019].
 This is because of forests proven ability to sequester carbon not only in the stems of trees [@davis2022] but in their soils as well [@guo2021].
-With increasingly negative climate impacts, the pressure to enhance the terrestrial sink is growing, and one of the tools to achieve this is afforestation.
-As such, there is growing attention being placed globally to afforestation efforts, which is highlighted by the UN's declaration of the years from 2021 through 2030 as the UN Decade of Ecosystem Restoration [@sewell2020].
-Many of the Nationally Determined Contributions (NDCs) countries have submitted, which detail their plans to keep warming under the 2C threshold, also rely to a significant degree on afforestation projects [@roe2019].
 
-The global gains that can be realized from afforestation, however, are uncertain, with broad estimates of cumulative sequestration ranging from 42 gigatonnes to 700 gigatonnes by the end of the century [@veldman2019]  [@bastin201]  [@Humpenöder2014]  [@sathaye2006]  [@doelman2020].
-In addition, the amount of land conversion that would be required also varies greatly, with estimates ranging from 0.3 billion hectares to 2.8 billion hectares [@veldman2019]  [@bastin201]  [@Humpenöder2014]  [@doelman2020].
-In many cases, these results suggest massive losses in the amount of pasture and cropland that would be available.
-This puts into question the feasibility of attaining these results, as almost 90 percent of deforestation worldwide is undertaken to expand the amount of agricultural land [@fao2021].
-There are many reasons for this.
-For example, as nations develop and disposable incomes rise, consumers often change to more meat-based diets.
-Population growth is also a driver and when it is especially high, which is the case at present throughout most of Africa, this also puts pressure on policymakers to allow greater conversion of forestland into farmland.
-Still, previous research has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017].
+The global gains that can be realized from afforestion, however, are uncertain, with broad estimates of cumulative sequestration ranging from 42 gigatonnes to 700 gigatonnes by the end of the century [@veldman2019]  [@bastin201]  [@humpenoder2014]  [@sathaye2006]  [@doelman2020].
+In addition, the amount of land conversion that would be required also varies greatly, with estimates ranging from 0.3 billion hectares to 2.8 billion hectares [@veldman2019]  [@bastin201]  [@humpenoder2014]  [@doelman2020].
+In many cases, these results suggest massive losses in pasture and cropland.
+This puts into question the feasibility of attaining these results, as forests are continually threatened by issues such as rising incomes and food prices and population growth that put pressure on policymakers to allow greater conversion of forestland into farmland.
+Previous research though has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017].
+Still, the manner in which these various and oftentimes competing concerns are addressed will have significant impacts on the global community.
 
-Faced with competing land-use demands, afforestation efforts need to be targeted to the regions of the world where they will provide the most benefits.
-Afforestation in boreal zones has been shown to provide the smallest net benefits, as the carbon sequestrations benefits of conifer forests are mitigated by the reduction in albedo in winter [@mykleby2017] [@li2015].
+Moreover, afforestation appears to not provide the same benefits in all regions of the world.
+In particular, boreal forests appear to provide the smallest net benefits, as the carbon sequestrations benefits of conifer forests are mitigated by the reduction in albedo in winter [@mykleby2017] [@li2015].
 Coniferous forests have also been shown to increase soil organic carbon less than broadleaf forests [@guo2021].
-Conversely, tropical forests have been linked with a strong positive result from afforestation, as the carbon sequestration benefit is complemented by a cooling impact due to changes in both albedo and evapotranspiration, and with temperate forests, the impact of afforestation is also shown to be positive, albeit less strongly so due to a winter warming effect [@li2015].
+With tropical forests, a strong positive result is derived from afforestation, as the carbon sequestration benefit is complemented by a cooling impact due to changes in both albedo and evapotranspiration, and with temperate forests, the impact of afforestation is also shown to be positive, albeit less strongly so due to a winter warming effect [@li2015].
 
@@
 It is estimated that these two regions hold at least 50% of the potential global gain while regions such as Northern Africa and the Middle East show little promise as forest growth rates are quite low [@doelman2020].
-Thus, it is promising that the African Forest Landscape Restoration Initiative (AFR100), which aims to afforest 100 million hectares by 2030, has roughly 70% of its commitments coming from Sub-Saharan nations [@auda2023].
+Thus, it is promising that the African Forest Landscape Restoration Initiative (AFR100), which aims to afforest 100 million hectares by 2030, has roughly 70% of its committments coming from Sub-Saharan nations [@auda2023].
 While many of these nations may be well-suited in biophysical terms for afforestation, there is concern about how long-lasting any investment may be due to proximate zones of political instability.
 There are other nations that have a big potential for gains from afforestation, but they also face challenges.
-India, for example, has a great potential to increase its carbon sequestration through afforestation, but due to its massive population and food security worries, its NDC does not entertain the idea of agricultural land being repurposed and instead primarily considers targeting wastelands that may not be well suited for afforestation [@amjath2019].
-Moreover, as land-use decisions in one country also tend to have impacts that spillover into many others [@meyfroidt2010] [@lambin2011], analysis needs to be conducted at a global scale.
+India, for example, has a great potential to increase its carbon sequestration through afforestation, but due to its masssive population and food security worries, its NDC does not entertain the idea of agricultural land being repurposed and instead primarily considers targeting wastelands that may not be well suited for afforestation [@amjath2019].
 
-Steinbuks and Hertel [@steinbuks2012] developed the FABLE model (Forestry, Agriculture, Biofuels, Land Use, and Environment) as a global partial equilibrium model to analyze optimal land use within a context of increasing demand for food, bioenergy, forest products, and non-provisioning ecosystem services, while also meeting greenhouse gas targets.
-The model encompasses nine sectors, including agriculture, livestock, food processing, biofuels, energy, forest, timber processing, and ecosystem services.
-While FABLE is a powerful tool for examining the optimal trajectory of various land uses under specific demand assumptions, it primarily focuses on specific sectors of the economy, neglecting general equilibrium effects and socio-economic or household welfare impacts [@steinbuks2012] [@hertel2017].
-Other efforts have used the Global Trade Analysis Project (GTAP) computable general equilibrium model to address such questions, but its standard model does not well account for land-use changes.
-As a result, the Land Use and Land Cover Database within the GTAP framework, which incorporates forestry remote-sensing products, was created [@baldos2012].
-It was further strengthened as part of the GTAP Agro-Ecological Zone (GTAP-AEZ) model, with the use of spatially explicit global land use data and through the incorporation of intra- and inter-regional land and land-based greenhouse emissions heterogeneity [@hertel2008].
-While the GTAP-AEZ model benefits from the inclusion of competition among different crops, grazing, and forest-based uses, it has limitations, primarily due to its infrequent updates.
-The KLUM@GTAP framework integrated the Kleines Land Use Model (KLUM) with an extended yet static version of GTAP called GTAP-EFL to evaluate the impacts of climate change on cropland allocation.
-GTAP-EFL separates energy factors from intermediate inputs and incorporates them into capital, while also considering CO2 emissions.
-KLUM, a global agricultural land-use model, connects the economy to global crop allocation to maximize producer returns under specific risk assumptions.
-KLUM@GTAP substitutes the land allocation mechanism within GTAP-EFL by utilizing regionally aggregated area changes in cropland determined by KLUM to update cropland shares in GTAP-EFL [@ronneberger2009].
-
-More recent modeling attempts of land-use change have begun to move away from comparative static analysis.
-For example, the dynamic GTAP model, supported by the GTAP-AEZ database, considers land market effects, which were identified as significant in driving results by Stevenson et al.
-[@stevenson2013], and has been utilized to evaluate the impacts of crop intensification on land use.
-It, however, has been critiqued, as global aggregates may mask localized shifts that can have implications for ecologically significant areas [@byerlee2014].
-This project will attempt to get around these concerns and assess the impact of various likely afforestation scenarios by melding GTAP with the Global Timber Model (GTM), which is a dynamic optimization model that can examine the impact of afforestation or avoided deforestation on the carbon sink [@sohngen2001].
-By joining these two models and passing outputs back and forth between the models, any limitations of either model can be minimized and more robust results, therefore, can be attained.
+This project will assess the impact of various likely scenarios by melding together two widely accepted global models, the Global Trade Analysis Project (GTAP) computable general equilibrium model [@aguilar2022] and the Global Timber Model (GTM) [@sohngen2001].
+The GTAP model has traditionally been used to analyze the impact of various shocks on things like agricultural prices and trade volumes.
+However, the standard model does not well account for land-use changes, and the GTAP Land Use and Land Cover (LULC) database has limitations due to its infrequent updates.
+The GTM is quite helpful in getting around these issues, as it is a dynamic optimization model that can examine the impact of afforestation or avoided deforestation on the carbon sink.
+By joining the two models and passing outputs back and forth between the models, any limitations of either model can be minimized and more robust results, therefore, can be attained.
 
@@
 To see the list of countries and the hectares pledged, please view Scenario 1 of the Excel document "Bonn Challenge-Africa".
+
 2) Partial attainment of AFR100 pledges as part of Bonn Challenge: This scenario analyzes full attainment for nations with conflict severity level less than or equal to limited and 0% attainment for regions with a higher conflict severity rating.
@@
 To see the list of countries and the hectares pledged, please view Scenario 2 of the Excel document "Bonn Challenge-Africa".
+
 3) Full attainment of Latin America pledges as part of Bonn Challenge: This says that there will be a 33.64 million hectare increase in forest area by 2030 [@bonn2023].
 To see the list of countries and the hectares pledged, please view the Latin America tab of the Excel document "Bonn Challenge-Other Regions".
+
 4) Full attainment of tropical nations' pledges as part of Bonn Challenge: This says that there will be a 182.30 million hectare increase in forest area by 2030 [@bonn2023].
@@
 Nations that lie in both tropical and sub-tropical zones and taken to be tropical.
+
 5) Full attainment of UN strategic plan for forests 2030 (UNFF): This says that all deforestation will halt and there will be a 3% (~120 million hectare) increase in forest area worldwide by 2030 [@sewell2020].
 This will be modeled as a uniform 3% increase in forested area across all nations.
+
 6) Expansion of UN strategic plan for forests: This says that all deforestation will halt and there will be a 15% (~600 million hectare) increase in forest area worldwide by 2030 [@sewell2020].
@@
 
-DeFries et al.
-[@defries2010] found a strong correlation between forest loss and exports of agricultural products.
-On the other hand, Turner et al.
-[@turner2005] investigated the role of trade on forests in the context of the Free Trade Area of the Americas (FTAA) using the spatial partial-equilibrium Global Forest Products Model and found that FTAA has a positive impact on forest resources.
 # References

### 2024-01-23 paper.Rmd
Fixed the conflict in the abstract

--- 2023-08-07 paper.Rmd
@@
 
-The global gains that can be realized from afforestation, however, are uncertain, with broad estimates of cumulative sequestration ranging from 42 gigatonnes to 700 gigatonnes by the end of the century [@veldman2019]  [@bastin201]  [@Humpenöder2014]  [@sathaye2006]  [@doelman2020].
-In addition, the amount of land conversion that would be required also varies greatly, with estimates ranging from 0.3 billion hectares to 2.8 billion hectares [@veldman2019]  [@bastin201]  [@Humpenöder2014]  [@doelman2020].
+The global gains that can be realized from afforestation, however, are uncertain, with broad estimates of cumulative sequestration ranging from 42 gigatonnes to 700 gigatonnes by the end of the century [@veldman2019]  [@bastin201]  [@humpenoder2014]  [@sathaye2006]  [@doelman2020].
+In addition, the amount of land conversion that would be required also varies greatly, with estimates ranging from 0.3 billion hectares to 2.8 billion hectares [@veldman2019]  [@bastin201]  [@humpenoder2014]  [@doelman2020].
 In many cases, these results suggest massive losses in the amount of pasture and cropland that would be available.
@@
 There are other nations that have a big potential for gains from afforestation, but they also face challenges.
-India, for example, has a great potential to increase its carbon sequestration through afforestation, but due to its massive population and food security worries, its NDC does not entertain the idea of agricultural land being repurposed and instead primarily considers targeting wastelands that may not be well suited for afforestation [@amjath2019].
+India, for example, has a great potential to increase its carbon sequestration through afforestation, but due to its massive population and food security worries, its NDC does not entertain the idea of agricultural land being repurposed and instead primarily considers targeting wa    stelands that may not be well suited for afforestation [@amjath2019].
 Moreover, as land-use decisions in one country also tend to have impacts that spillover into many others [@meyfroidt2010] [@lambin2011], analysis needs to be conducted at a global scale.
@@
 
-DeFries et al.
 [@defries2010] found a strong correlation between forest loss and exports of agricultural products.
-On the other hand, Turner et al.
-[@turner2005] investigated the role of trade on forests in the context of the Free Trade Area of the Americas (FTAA) using the spatial partial-equilibrium Global Forest Products Model and found that FTAA has a positive impact on forest resources.
+On the other hand, [@turner2005] investigated the role of trade on forests in the context of the Free Trade Area of the Americas (FTAA) using the spatial partial-equilibrium Global Forest Products Model and found that FTAA has a positive impact on forest resources.
+
 # References

### 2024-01-24 paper.Rmd
Description of combining the model

@@
 
-The global gains that can be realized from afforestation, however, are uncertain, with broad estimates of cumulative sequestration ranging from 42 gigatonnes to 700 gigatonnes by the end of the century [@veldman2019]  [@bastin201]  [@humpenoder2014]  [@sathaye2006]  [@doelman2020].
-In addition, the amount of land conversion that would be required also varies greatly, with estimates ranging from 0.3 billion hectares to 2.8 billion hectares [@veldman2019]  [@bastin201]  [@humpenoder2014]  [@doelman2020].
+The global gains that can be realized from afforestation, however, are uncertain, with broad estimates of cumulative sequestration ranging from 42 gigatonnes to 700 gigatonnes by the end of the century [@veldman2019]  [@bastin2019]  [@humpenoder2014]  [@sathaye2006]  [@doelman2020].
+In addition, the amount of land conversion that would be required also varies greatly, with estimates ranging from 0.3 billion hectares to 2.8 billion hectares [@veldman2019]  [@bastin2019]  [@humpenoder2014]  [@doelman2020].
 In many cases, these results suggest massive losses in the amount of pasture and cropland that would be available.
@@
 There are other nations that have a big potential for gains from afforestation, but they also face challenges.
-India, for example, has a great potential to increase its carbon sequestration through afforestation, but due to its massive population and food security worries, its NDC does not entertain the idea of agricultural land being repurposed and instead primarily considers targeting wa    stelands that may not be well suited for afforestation [@amjath2019].
+India, for example, has a great potential to increase its carbon sequestration through afforestation, but due to its massive population and food security worries, its NDC does not entertain the idea of agricultural land being repurposed and instead primarily considers targeting wastelands that may not be well suited for afforestation [@amjath2019].
 Moreover, as land-use decisions in one country also tend to have impacts that spillover into many others [@meyfroidt2010] [@lambin2011], analysis needs to be conducted at a global scale.
@@
 By joining these two models and passing outputs back and forth between the models, any limitations of either model can be minimized and more robust results, therefore, can be attained.
+
+# Connecting the GTAP and GTM models
+
+The standard GTAP model represents a leading CGE model of the global economy with a specific focus on trade linkages.
+Its general equilibrium nature means that all markets that are considered in the model are cleared with market-clearing prices.
+The extensive coverage of sectors---the latest version of the GTAP database includes 68 sectors---makes it difficult to model sectors individually; instead the model applies nested CES production functions to all sectors, differentiating them by parameters only.
+To model one sector, e.g., forestry, differently from the rest would require major redesign of the model.
+While specific versions of the GTAP model are definitely possible but extremely common as they require to be be fully parametrized for all countries, which makes the option of adapting the GTAP model for a one-off application impractical.
+To analyze carbon sequestration in GTAP, which would require a lot of effort, we instead create a connection between the standard GTAP model and the GTM model which has been specifically designed to analyze the relationship among land rents, forested land cover and carbon sequestration through forests, among other things.
+
+Establishing a connection between the standard GTAP model and the GTM model is a two-step process.
+First, we decide on the set of common variables between the two models in those markets that are included in both models and their closure, i.e., which variables will be determined endogenously by one model and held as exogenous for the other model.
+In our case, we selected the return on land to be modeled by the GTAP model and fed into the GTM model as exogenous, while the amount of land available for crop and livestock production, while exogenous to the GTAP model, is endogenously determined by the GTM model.
+In the second, step we establish the connection between the models on the software level.
+This step requires that the GTAP model, written in GEMPACK can operate with the inputs from the GTM model, written in GAMS, and submit the outputs back to the GTM model.
+In order to allow for the two quite dated technologies, GEMPACK and GAMS, which were developed long time age before software integration become important or possible, we created a set of user functions in R that allow executing both models, collecting their outputs and an creating  parameter files for both systems automatically.
+
+With R being able to operate both the GTAP and GTM models automatically, we are able to formulate the combined solution of the both models as a fairly simple math program that is solved in R:
+
+\begin{equation}
+\mbox{GTAP}:V_1\to V_2
+\end{equation}
+
+\begin{equation}
+\mbox{GTM}:V_2\to V_1
+\end{equation}
+
+\begin{equation}
+s.t. V_2 = V_1
+\end{equation}
+
+where $V_1$ is a vector of all variables in both systems, i.e., all variables in GTAP and all variables in GTM, that is solved by the GTAP model (represented by function \mbox{GTAP}) into a vector of variables $V_2$.
+At the same time, when $V_2$ in processed by the GTM, its solution is $V_1$.
+To solve the system simultaneously, we require that $V_1$ and $V_2$ be equal.
+In our case, the vector of variables $V_1$ includes all GTAP variables and the land change variables, which are exogenous to the GTAP model; the output vector of variables $V_2$ contains the change in the rents, which is treated as exogenous in the GTM model.
+With several iterations, the system converges to the land change and land rent change that is consistent across the models, i.e., the land change results in the land rent changes that, when entered into the GTM model, result in the same land changes as assumed by the GTAP model.
 

### 2024-02-20 paper.Rmd
Update paper.Rmd

@@
 date: "`r format(Sys.time(), '%B %d, %Y')`"
-author: "Eric Davis, Maros Ivanic, Brent Sohngen"
+author: "Eric C. Davis, Maros Ivanic, and Brent Sohngen"
 output: 
@@
 
-The standard GTAP model represents a leading CGE model of the global economy with a specific focus on trade linkages.
+The standard GTAP model is a leading CGE representation of the global economy and its trade linkages.
 Its general equilibrium nature means that all markets that are considered in the model are cleared with market-clearing prices.
-The extensive coverage of sectors---the latest version of the GTAP database includes 68 sectors---makes it difficult to model sectors individually; instead the model applies nested CES production functions to all sectors, differentiating them by parameters only.
-To model one sector, e.g., forestry, differently from the rest would require major redesign of the model.
-While specific versions of the GTAP model are definitely possible but extremely common as they require to be be fully parametrized for all countries, which makes the option of adapting the GTAP model for a one-off application impractical.
-To analyze carbon sequestration in GTAP, which would require a lot of effort, we instead create a connection between the standard GTAP model and the GTM model which has been specifically designed to analyze the relationship among land rents, forested land cover and carbon sequestration through forests, among other things.
+There are 68 market sectors covered in the latest version of the GTAP database includes  sectors.
+As the model applies nested CES production functions to all sectors, differentiating them by parameters only, it is difficult to model a sector individually.
+To model one sector, in this case forestry, differently from the rest would require a major redesign of the model.
+While not impossible, we chose to create a connection between the standard GTAP model and the GTM model, as GTM has been specifically designed to analyze the relationship among land rents, forested land cover, and carbon sequestration through forests.
+Doing so obviated the need to reparametrize the GTAP model for all countries.
 
-Establishing a connection between the standard GTAP model and the GTM model is a two-step process.
-First, we decide on the set of common variables between the two models in those markets that are included in both models and their closure, i.e., which variables will be determined endogenously by one model and held as exogenous for the other model.
+Establishing a connection between the standard GTAP model and the GTM model was a two-step process.
+First, we identified the set of common variables that are included in both models and their closures.
+That is, we selected which variables would be determined endogenously by one model and held as exogenous for the other model.
 In our case, we selected the return on land to be modeled by the GTAP model and fed into the GTM model as exogenous, while the amount of land available for crop and livestock production, while exogenous to the GTAP model, is endogenously determined by the GTM model.

### 2024-02-20 paper.Rmd
Update paper.Rmd

@@
 First, we identified the set of common variables that are included in both models and their closures.
-That is, we selected which variables would be determined endogenously by one model and held as exogenous for the other model.
-In our case, we selected the return on land to be modeled by the GTAP model and fed into the GTM model as exogenous, while the amount of land available for crop and livestock production, while exogenous to the GTAP model, is endogenously determined by the GTM model.
-In the second, step we establish the connection between the models on the software level.
-This step requires that the GTAP model, written in GEMPACK can operate with the inputs from the GTM model, written in GAMS, and submit the outputs back to the GTM model.
-In order to allow for the two quite dated technologies, GEMPACK and GAMS, which were developed long time age before software integration become important or possible, we created a set of user functions in R that allow executing both models, collecting their outputs and an creating  parameter files for both systems automatically.
+That is, we selected which variables would be determined endogenously by one model and held as exogenous in the other model.
+In our case, we modeled the return on land using the GTAP model and fed those results into the GTM model as exogenous.
+Then, the amount of land available for crop and livestock production was endogenously determined by the GTM model and fed into the GTAP model as an exogenous input.
+In the second step, we established the connection between the models on the software level.
+This step required that the GTAP model, which is written in GEMPACK, be able to handle with the inputs from the GTM model, which is written in GAMS, and submit the outputs back to the GTM model.
+As GEMPACK and GAMS were developed long before software integration became important or even possible, we had to create a set of user functions in R that executed both models, collected their outputs, and an created parameter files automatically.
 

### 2024-02-24 paper.Rmd

@@
 This step required that the GTAP model, which is written in GEMPACK, be able to handle with the inputs from the GTM model, which is written in GAMS, and submit the outputs back to the GTM model.
-As GEMPACK and GAMS were developed long before software integration became important or even possible, we had to create a set of user functions in R that executed both models, collected their outputs, and an created parameter files automatically.
+As GEMPACK and GAMS were developed long before software integration became important or even possible, we had to create a set of user functions in R that executed both models, collected their outputs, and created parameter files automatically.
 
-With R being able to operate both the GTAP and GTM models automatically, we are able to formulate the combined solution of the both models as a fairly simple math program that is solved in R:
+Once R was able to operate both models automatically, we were then able to formulate the combined solution as a fairly simple math program that is solved in R:
 
@@
 
-where $V_1$ is a vector of all variables in both systems, i.e., all variables in GTAP and all variables in GTM, that is solved by the GTAP model (represented by function \mbox{GTAP}) into a vector of variables $V_2$.
-At the same time, when $V_2$ in processed by the GTM, its solution is $V_1$.
+where $V_1$ is a vector of all the variables in both systems, i.e., all variables in GTAP and all variables in GTM.
+Of import, $V_1$ includes the land change variables, which are exogenous to the GTAP model.
+$V_1$ is then used by the GTAP model (represented by function \mbox{GTAP}) and its outputs are passed to GTM as a vector of variables $V_2$.
+This output vector contains the change in rents, which is treated as exogenous in the GTM model.
+GTM then processes the $V_2$ variables as inputs and passes its solution back to GTAP as $V_1$.
 To solve the system simultaneously, we require that $V_1$ and $V_2$ be equal.
-In our case, the vector of variables $V_1$ includes all GTAP variables and the land change variables, which are exogenous to the GTAP model; the output vector of variables $V_2$ contains the change in the rents, which is treated as exogenous in the GTM model.
-With several iterations, the system converges to the land change and land rent change that is consistent across the models, i.e., the land change results in the land rent changes that, when entered into the GTM model, result in the same land changes as assumed by the GTAP model.
+With several iterations, the system converges to the change in land and land rent that is consistent across the models, i.e., the land change results in the land rent changes that, when entered into the GTM model, result in the same land changes as assumed by the GTAP model.
 

### 2024-02-25 paper.Rmd
Update paper.Rmd

@@
 It, however, has been critiqued, as global aggregates may mask localized shifts that can have implications for ecologically significant areas [@byerlee2014].
-This project will attempt to get around these concerns and assess the impact of various likely afforestation scenarios by melding GTAP with the Global Timber Model (GTM), which is a dynamic optimization model that can examine the impact of afforestation or avoided deforestation on the carbon sink [@sohngen2001].
-By joining these two models and passing outputs back and forth between the models, any limitations of either model can be minimized and more robust results, therefore, can be attained.
+In addition, in the standard GTAP model, it is difficult to model a sector individually.
+There are 68 market sectors covered in the latest version of the GTAP database.
+The GTAP model's general equilibrium nature means that all these markets are cleared with market-clearing prices.
+As the model applies nested CES production functions to all sectors and differentiates them by parameters only, to model one sector, in this case forestry, differently from the rest would require a major redesign of the model.
+While not impossible, this project will attempt to get around these concerns and assess the impact of various likely land-use scenarios by melding GTAP with the Global Timber Model (GTM), which is a dynamic optimization model that has been specifically designed to analyze the relationship among land rents, forested land cover, and carbon sequestration through forests and thus is better able to examine the impact of afforestation or avoided deforestation on the carbon sink [@sohngen2001].
+Joining these two models and passing outputs back and forth between the models should enable any limitations of either model to be minimized and more robust results, therefore, to be attained.
 
 # Connecting the GTAP and GTM models
-
-The standard GTAP model is a leading CGE representation of the global economy and its trade linkages.
-Its general equilibrium nature means that all markets that are considered in the model are cleared with market-clearing prices.
-There are 68 market sectors covered in the latest version of the GTAP database includes  sectors.
-As the model applies nested CES production functions to all sectors, differentiating them by parameters only, it is difficult to model a sector individually.
-To model one sector, in this case forestry, differently from the rest would require a major redesign of the model.
-While not impossible, we chose to create a connection between the standard GTAP model and the GTM model, as GTM has been specifically designed to analyze the relationship among land rents, forested land cover, and carbon sequestration through forests.
-Doing so obviated the need to reparametrize the GTAP model for all countries.
 

### 2024-02-25 paper.Rmd
Update paper.Rmd

@@
 
+1a) On aggregate zero percent net deforestation Globally
+
 1) Full attainment of AFR100 pledges as part of Bonn Challenge: This says that there will be a 127.81 million hectare increase in forest area just in Sub-Saharan Africa by 2030 [@bonn2023].
@@
 
+# Results
+We first examined the impact of a scenario where the rate of deforestation globally was reduced to zero through the implementation of a uniform global tax rate on the conversion of forestland to agricultural land.
+A tax rate of roughly 70 percent was required to achieve this.
+With this policy in place, most regions of the world would see significant declines in the amount of agricultural land in 2030 relative to the baseline scenario where land conversion was not taxed.
+For the United States, agricultural land would be 2.1 percent less (Table 1).
+Larger decreases would happen in important regions such as Brazil (-9.7 percent), the rest of South America (-14.7 percent), Southeast Asia (-10.0 percent), and Sub-Saharah Africa (-4.8 percent).
+
+Table 1
+	                              ∆ (%)
+Central America	                -16.9
+East Asia	                      -14.4
+North Africa & the Middle East	-11.1
+Southeast Asia	                -10.0
+Russia	                        -9.8
+Brazil	                        -9.7
+Rest of South America	          -14.7
+South Asia	                    -9.1
+Canada	                        -8.1
+Oceania	                        -7.5
+European Union	                -6.4
+Sub-Saharan Africa	            -4.8
+EUII	                          -3.3
+China	                          -2.5
+United States	                  -2.1
+Japan	                          11.4
+
 [@defries2010] found a strong correlation between forest loss and exports of agricultural products.

### 2024-02-25 paper.Rmd
Update paper.Rmd

@@
 
+These reductions in the amount of agricultural land prompt changes in the quantity of agricultural goods produced for many regions.
+For example, East Asia and Southeast Asia see the largest declines in percentage terms in agricultural production with decreases of 5.9 percent and 2.8 percent, respectively (Table 2).
+Surprisingly, Brazil and the United States, despite the derease in available agricultural land compared to the baseline 2030 scenario, see increased agricultural output of 0.2 percent and 0.8 percent, respectively.
+This may be tied to the sharp declines in other regions prompting greater production for export.
+
+Table 2
+			∆ (%)
+East Asia		-5.9
+Southeast Asia		-2.8
+Central America		-2.7
+South Asia		-2.1
+Rest of South America	-1.6
+North Africa & the Middle East	-1.2
+Sub-Saharan Africa	-0.5
+China			-0.1
+Russia			0.1
+Brazil			0.2
+EUII			0.3
+European Union		0.4
+United States		0.8
+Oceania			0.9
+Canada			1.3
+Japan			1.5
+
+When we examine how this impacts exports of agricultural goods, we find that Japan and the United States appear to be the biggest beneficiaries of this policy with their exports increasing 13.5 percent and 5.0 percent, respectively (table 3).
+The regions where agricultural exports drop most significantly are South Asia, Southeast Asia, and Central America who see decreases of 11.0 percent, 7.3 percent, and 6.9 percent, respectively.
+
+Table 3
+				∆ (%)
+				Agriculture
+Japan				13.5
+United States			5.0
+Canada				2.6
+Russia				2.4
+Oceania				2.2
+EUII				2.2
+China				1.5
+European Union			1.4
+Brazil				1.4
+Sub-Saharan Africa		0.0
+North Africa & the Middle East	-2.3
+Rest of South America		-3.8
+Central America			-6.9
+Southeast Asia			-7.3
+South Asia			-11.0
+East Asia			-19.8
+
+Overall, this policy, while avoiding deforestation, has a rather significant negative economic impact on all regions except Japan and Canada who see an increase in consumer well-being of $977.1 million and $166.9 million, respectively (table 4).
+Consumer well-being is a measurement of equivalent variation and here measures the benefit/harm consumers experience in terms of income as a result of the policy changes.
+The United States, along with 10 other regions, see well-being drop by more than $1 billion.
+The most negatively impacted regions are South Asia (-28.0 billion), China (-14.5 billion), and Southeast Asia (-13.6 billion).
+In total, such a policy is estimated to decrease global consumer well-being by $92.1 billion.
+
+Table 4
+				∆ ($)
+Japan				977,125,000
+Canada				166,937,500
+Oceania				-169,625,000
+EUII				-588,500,000
+Brazil				-735,125,000
+United States			-1,085,000,000
+Russia				-1,152,375,000
+Rest of South America		-3,848,625,000
+Central America			-4,471,500,000
+North Africa & the Middle East	-5,012,125,000
+East Asia			-5,065,625,000
+European Union			-7,152,500,000
+Sub-Saharan Africa		-7,850,500,000
+Southeast Asia			-13,601,000,000
+China				-14,452,000,000
+South Asia			-28,024,000,000
+
 [@defries2010] found a strong correlation between forest loss and exports of agricultural products.
 On the other hand, [@turner2005] investigated the role of trade on forests in the context of the Free Trade Area of the Americas (FTAA) using the spatial partial-equilibrium Global Forest Products Model and found that FTAA has a positive impact on forest resources.
-
 # References

### 2024-02-25 paper.Rmd
Update paper.Rmd

@@
 
+Much of this loss in well-being may be the result of higher prices.
+Japan is the only nation that sees any positive price impacts resulting from this policy, with the price of agricultural goods decreasing 0.2 percent (table 5).
+All other regions, for both agricultural and non-agricultural products, see prices rise between 0.1 and 2.8 percent.
+
+Table 5
+									∆ (%)				
+	Agricultural Goods		Forest Products		Lumber		Paper Products		Other Non-Agricultural Goods
+Oceania					0.6	0.3	0.3	0.3	0.3
+China					0.7	0.3	0.2	0.2	0.3
+Japan					-0.2	0.2	0.2	0.2	0.2
+East Asia				2.8	0.2	0.2	0.2	0.2
+Southeast Asia				1.5	0.2	0.2	0.2	0.2
+South Asia				1.6	0.2	0.2	0.1	0.2
+Canada					0.4	0.3	0.3	0.3	0.4
+United States				0.4	0.3	0.3	0.3	0.3
+Central America				1.3	0.2	0.2	0.2	0.2
+Brazil					0.7	0.3	0.3	0.3	0.3
+Rest of South America			1.1	0.2	0.2	0.2	0.2
+European Union				0.5	0.2	0.2	0.2	0.2
+EUII					0.6	0.3	0.2	0.2	0.2
+Russia					0.6	0.2	0.2	0.2	0.2
+Sub-Saharan Africa			0.9	0.3	0.2	0.2	0.3
+North Africa & the Middle East		1.1	0.2	0.2	0.2	0.2
+
 [@defries2010] found a strong correlation between forest loss and exports of agricultural products.

### 2024-02-26 paper.Rmd
Update paper.Rmd

--- 2024-02-25 paper.Rmd
@@
 
-Table 1
+Table 1 qo-Land
 	                         ∆ (%)
@@
 
-Table 2
+Table 2 qo-Ag
 			∆ (%)
@@
 
-Table 3
+Table 3 qxw-Ag
 				∆ (%)
@@
 
-Table 4
+Table 4 EV
 				∆ ($)
@@
 
+Much of this loss in well-being may be the result of higher prices.
+Japan is the only nation that sees any positive price impacts resulting from this policy, with the price of agricultural goods decreasing 0.2 percent (table 5).
+All other regions, for both agricultural and non-agricultural products, see prices rise between 0.1 and 2.8 percent.
+
+Table 5 pm
+									∆ (%)				
+	Agricultural Goods		Forest Products		Lumber		Paper Products		Other Non-Agricultural Goods
+Oceania					0.6	0.3	0.3	0.3	0.3
+China					0.7	0.3	0.2	0.2	0.3
+Japan					-0.2	0.2	0.2	0.2	0.2
+East Asia				2.8	0.2	0.2	0.2	0.2
+Southeast Asia				1.5	0.2	0.2	0.2	0.2
+South Asia				1.6	0.2	0.2	0.1	0.2
+Canada					0.4	0.3	0.3	0.3	0.4
+United States				0.4	0.3	0.3	0.3	0.3
+Central America				1.3	0.2	0.2	0.2	0.2
+Brazil					0.7	0.3	0.3	0.3	0.3
+Rest of South America			1.1	0.2	0.2	0.2	0.2
+European Union				0.5	0.2	0.2	0.2	0.2
+EUII					0.6	0.3	0.2	0.2	0.2
+Russia					0.6	0.2	0.2	0.2	0.2
+Sub-Saharan Africa			0.9	0.3	0.2	0.2	0.3
+North Africa & the Middle East		1.1	0.2	0.2	0.2	0.2
+
+Land price seems weird. I have it decreasing????
+
 [@defries2010] found a strong correlation between forest loss and exports of agricultural products.

### 2024-02-26 paper.Rmd
Update paper.Rmd

@@
 While not impossible, this project will attempt to get around these concerns and assess the impact of various likely land-use scenarios by melding GTAP with the Global Timber Model (GTM), which is a dynamic optimization model that has been specifically designed to analyze the relationship among land rents, forested land cover, and carbon sequestration through forests and thus is better able to examine the impact of afforestation or avoided deforestation on the carbon sink [@sohngen2001].
-Joining these two models and passing outputs back and forth between the models should enable any limitations of either model to be minimized and more robust results, therefore, to be attained.
+The latest versions of both models were used.
+For the GTAP model, this meant using the 11b data which has 2017 as a base year.
+This data was then updated to the base year (2023) using data from the Centre for Prospective Studies and International Information (CEPII) to account for the projected changes in gross domestic product (GDP), population, and productivity for each country in the world.
 
@@
 
-Establishing a connection between the standard GTAP model and the GTM model was a two-step process.
+By joining these two models and passing outputs back and forth between the models, any limitations of either model should be minimized and more robust results, therefore, should be attained.
+Establishing a connection between the standard GTAP model and the GTM model required  a two-step process.
 First, we identified the set of common variables that are included in both models and their closures.
@@
 
-# Scenarios
+# Modeling the impact of possible measures to mitigate deforestation/promote afforestation
+
+To understand the impact of potential regional and global afforestation actions, an important first step was creating a realistic baseline scenario that depicts the likely changes between over the decade under study: 2023-2033.
+This was done by taking into account key drivers of change like growth in population and GDP.
+Without accounting for this, results would likely have been biased as they would have failed to account for the increasing pressure to convert forests into agricultural land.
+Data for this were drawn from sources such as CEPII and the United Nations.
+
+When examining the likely impacts of the baseline changes alone, it is estimated that by 2033, significant deforestation would have occured in many regions of the world. For example, in Brazil...
 

### 2024-02-26 paper.Rmd
Update paper.Rmd

@@
 
-1a) On aggregate zero percent net deforestation Globally
-
-1) Full attainment of AFR100 pledges as part of Bonn Challenge: This says that there will be a 127.81 million hectare increase in forest area just in Sub-Saharan Africa by 2030 [@bonn2023].
-To see the list of countries and the hectares pledged, please view Scenario 1 of the Excel document "Bonn Challenge-Africa".
-
-2) Partial attainment of AFR100 pledges as part of Bonn Challenge: This scenario analyzes full attainment for nations with conflict severity level less than or equal to limited and 0% attainment for regions with a higher conflict severity rating.
-Using this measure, there would only be a 54.95 million hectare increase in forest area by 2030 [@bonn2023]  [@acled2023].
-To see the list of countries and the hectares pledged, please view Scenario 2 of the Excel document "Bonn Challenge-Africa".
-
-3) Full attainment of Latin America pledges as part of Bonn Challenge: This says that there will be a 33.64 million hectare increase in forest area by 2030 [@bonn2023].
-To see the list of countries and the hectares pledged, please view the Latin America tab of the Excel document "Bonn Challenge-Other Regions".
-
-4) Full attainment of tropical nations' pledges as part of Bonn Challenge: This says that there will be a 182.30 million hectare increase in forest area by 2030 [@bonn2023].
-To see the list of countries and the hectares pledged, please view the Tropical Forest tab of the Excel document "Bonn Challenge-Other Regions".
-Nations that lie in both tropical and sub-tropical zones and taken to be tropical.
-
-5) Full attainment of UN strategic plan for forests 2030 (UNFF): This says that all deforestation will halt and there will be a 3% (~120 million hectare) increase in forest area worldwide by 2030 [@sewell2020].
-This will be modeled as a uniform 3% increase in forested area across all nations.
-
-6) Expansion of UN strategic plan for forests: This says that all deforestation will halt and there will be a 15% (~600 million hectare) increase in forest area worldwide by 2030 [@sewell2020].
-This will be modeled as a uniform 15% increase in forested area across all nations.
-
-# Results
-We first examined the impact of a scenario where the rate of deforestation globally was reduced to zero through the implementation of a uniform global tax rate on the conversion of forestland to agricultural land.
-A tax rate of roughly 70 percent was required to achieve this.
-With this policy in place, most regions of the world would see significant declines in the amount of agricultural land in 2030 relative to the baseline scenario where land conversion was not taxed.
+Then, we examined what it would take given the pressures of a still growing global population with the concomitant increasing appetite for foodstuffs to end, on net, the practice of deforestation globally.
+Results suggest that it would require a tax of roughly 70.3 percent to be assessed on the value of the converted land for the deforestation rate to fall globally to roughly zero.
+The economic impact of such a policy measure would not be limited to the cost of the tax, it would also change impact food production and global trade.
+For example, with this policy in place, most regions of the world would see significant declines in the amount of agricultural land in 2033 relative to the baseline scenario where land conversion was not taxed.
 For the United States, agricultural land would be 2.1 percent less (Table 1).
-Larger decreases would happen in important regions such as Brazil (-9.7 percent), the rest of South America (-14.7 percent), Southeast Asia (-10.0 percent), and Sub-Saharah Africa (-4.8 percent).
+Larger decreases would happen in other major agricultural production regions such as China (-2.5 percent), South Asia (-8.0 percent), Brazil (-8.5 percent), and Russia (-9.3 percent).
 
@@
 				∆ (%)
-Central America	                -16.9
-East Asia	                -14.4
-North Africa & the Middle East	-11.1
-Southeast Asia	                -10.0
-Russia	                         -9.8
-Brazil	                         -9.7
-Rest of South America	        -14.7
-South Asia	                 -9.1
-Canada	                         -8.1
-Oceania	                         -7.5
-European Union	                 -6.4
-Sub-Saharan Africa	         -4.8
-EUII	                         -3.3
+East Asia			-14.6
+Central America			-14.5
+North Africa & the Middle East	-10.3
+Russia				-9.3
+Southeast Asia			-9.0
+Brazil				-8.5
+Rest of South America		-11.7
+South Asia			-8.0
+Canada				-7.9
+West & Central Europe 		-6.2
+Oceania				-5.9
+Sub-Saharan Africa		-4.1
 China				-2.5
-United States	                 -2.1
-Japan	                         11.4
+United States			-2.2
+Japan				17.5
 
 These reductions in the amount of agricultural land prompt changes in the quantity of agricultural goods produced for many regions.
-For example, East Asia and Southeast Asia see the largest declines in percentage terms in agricultural production with decreases of 5.9 percent and 2.8 percent, respectively (Table 2).
-Surprisingly, Brazil and the United States, despite the derease in available agricultural land compared to the baseline 2030 scenario, see increased agricultural output of 0.2 percent and 0.8 percent, respectively.
+For example, East Asia and Southeast Asia see the largest declines in percentage terms in agricultural production with decreases of 5.1 percent and 2.0 percent, respectively (Table 2).
+Surprisingly, Brazil and the United States, despite the derease in available agricultural land compared to the baseline 2033 scenario, see increased agricultural output of 0.2 percent and 0.6 percent, respectively.
 This may be tied to the sharp declines in other regions prompting greater production for export.
@@
 				∆ (%)
-East Asia		-5.9
-Southeast Asia		-2.8
-Central America		-2.7
-South Asia		-2.1
-Rest of South America	-1.6
-North Africa & the Middle East	-1.2
-Sub-Saharan Africa	-0.5
+				Agriculture
+East Asia			-5.1
+Southeast Asia			-2.0
+Central America			-2.0
+South Asia			-1.1
+North Africa & the Middle East	-0.9
+Sub-Saharan Africa		-0.3
 China				-0.1
@@
 Brazil				0.2
-EUII			0.3
-European Union		0.4
-United States		0.8
-Oceania			0.9
+Rest of South America		-1.2
+West & Central Europe 		0.4
+United States			0.6
+Oceania				0.7
 Canada				1.3
-Japan			1.5
+Japan				1.4
 
@@
 				Agriculture
-Japan				13.5
-United States			5.0
-Canada				2.6
-Russia				2.4
-Oceania				2.2
-EUII				2.2
-China				1.5
-European Union			1.4
-Brazil				1.4
-Sub-Saharan Africa		0.0
-North Africa & the Middle East	-2.3
-Rest of South America		-3.8
-Central America			-6.9
-Southeast Asia			-7.3
-South Asia			-11.0
-East Asia			-19.8
+East Asia			-5.1
+Southeast Asia			-2.0
+Central America			-2.0
+South Asia			-1.1
+North Africa & the Middle East	-0.9
+Sub-Saharan Africa		-0.3
+China				-0.1
+Russia				0.1
+Brazil				0.2
+Rest of South America		-1.2
+West & Central Europe 		0.4
+United States			0.6
+Oceania				0.7
+Canada				1.3
+Japan				1.4
 
@@
 
+1a) On aggregate zero percent net deforestation Globally
+
+1) Full attainment of AFR100 pledges as part of Bonn Challenge: This says that there will be a 127.81 million hectare increase in forest area just in Sub-Saharan Africa by 2030 [@bonn2023].
+To see the list of countries and the hectares pledged, please view Scenario 1 of the Excel document "Bonn Challenge-Africa".
+
+2) Partial attainment of AFR100 pledges as part of Bonn Challenge: This scenario analyzes full attainment for nations with conflict severity level less than or equal to limited and 0% attainment for regions with a higher conflict severity rating.
+Using this measure, there would only be a 54.95 million hectare increase in forest area by 2030 [@bonn2023]  [@acled2023].
+To see the list of countries and the hectares pledged, please view Scenario 2 of the Excel document "Bonn Challenge-Africa".
+
+3) Full attainment of Latin America pledges as part of Bonn Challenge: This says that there will be a 33.64 million hectare increase in forest area by 2030 [@bonn2023].
+To see the list of countries and the hectares pledged, please view the Latin America tab of the Excel document "Bonn Challenge-Other Regions".
+
+4) Full attainment of tropical nations' pledges as part of Bonn Challenge: This says that there will be a 182.30 million hectare increase in forest area by 2030 [@bonn2023].
+To see the list of countries and the hectares pledged, please view the Tropical Forest tab of the Excel document "Bonn Challenge-Other Regions".
+Nations that lie in both tropical and sub-tropical zones and taken to be tropical.
+
+5) Full attainment of UN strategic plan for forests 2030 (UNFF): This says that all deforestation will halt and there will be a 3% (~120 million hectare) increase in forest area worldwide by 2030 [@sewell2020].
+This will be modeled as a uniform 3% increase in forested area across all nations.
+
+6) Expansion of UN strategic plan for forests: This says that all deforestation will halt and there will be a 15% (~600 million hectare) increase in forest area worldwide by 2030 [@sewell2020].
+This will be modeled as a uniform 15% increase in forested area across all nations.
+
 [@defries2010] found a strong correlation between forest loss and exports of agricultural products.

### 2024-02-26 paper.Rmd
Update paper.Rmd

@@
 
-When we examine how this impacts exports of agricultural goods, we find that Japan and the United States appear to be the biggest beneficiaries of this policy with their exports increasing 13.5 percent and 5.0 percent, respectively (table 3).
-The regions where agricultural exports drop most significantly are South Asia, Southeast Asia, and Central America who see decreases of 11.0 percent, 7.3 percent, and 6.9 percent, respectively.
+When we examine how this impacts exports of agricultural goods, we find that Japan and the United States appear to be the biggest beneficiaries of this policy with their exports increasing 8.8 percent and 3.5 percent, respectively (table 3).
+The regions where agricultural exports drop most significantly are East Asia, South Asia, Southeast Asia, and Central America who see decreases of 14.5 percent, 5.9 percent, 5.1 percent,and 5.0 percent, respectively.
 
@@
 				Agriculture
-East Asia			-5.1
-Southeast Asia			-2.0
-Central America			-2.0
-South Asia			-1.1
-North Africa & the Middle East	-0.9
-Sub-Saharan Africa		-0.3
-China				-0.1
-Russia				0.1
-Brazil				0.2
-Rest of South America		-1.2
-West & Central Europe 		0.4
-United States			0.6
-Oceania				0.7
-Canada				1.3
-Japan				1.4
+Japan				8.78
+United States			3.49
+Canada				3.13
+China				1.96
+Oceania				1.67
+Russia				1.52
+West & Central Europe 		1.16
+Brazil				1.03
+Rest of South America		-2.67
+Sub-Saharan Africa		0.02
+North Africa & the Middle East	-1.52
+Central America		-5.02
+Southeast Asia		-5.14
+South Asia		-5.94
+East Asia		-14.45
 
@@
 East Asia			-5,065,625,000
-European Union			-7,152,500,000
+West & Central Europe		-7,152,500,000
 Sub-Saharan Africa		-7,850,500,000
@@
 Much of this loss in well-being may be the result of higher prices.
-Japan is the only nation that sees any positive price impacts resulting from this policy, with the price of agricultural goods decreasing 0.2 percent (table 5).
-All other regions, for both agricultural and non-agricultural products, see prices rise between 0.1 and 2.8 percent.
+Japan is the only nation that sees any positive price impacts resulting from this policy, with the price of agricultural goods decreasing 0.3 percent (table 5).
+All other regions, for both agricultural and non-agricultural products, see prices rise between 0.3 and 5.1 percent.
 
@@
 	Agricultural Goods		Forest Products		Lumber		Paper Products		Other Non-Agricultural Goods
-Oceania					0.6	0.3	0.3	0.3	0.3
-China					0.7	0.3	0.2	0.2	0.3
-Japan					-0.2	0.2	0.2	0.2	0.2
-East Asia				2.8	0.2	0.2	0.2	0.2
-Southeast Asia				1.5	0.2	0.2	0.2	0.2
-South Asia				1.6	0.2	0.2	0.1	0.2
-Canada					0.4	0.3	0.3	0.3	0.4
-United States				0.4	0.3	0.3	0.3	0.3
-Central America				1.3	0.2	0.2	0.2	0.2
-Brazil					0.7	0.3	0.3	0.3	0.3
-Rest of South America			1.1	0.2	0.2	0.2	0.2
-European Union				0.5	0.2	0.2	0.2	0.2
-EUII					0.6	0.3	0.2	0.2	0.2
-Russia					0.6	0.2	0.2	0.2	0.2
-Sub-Saharan Africa			0.9	0.3	0.2	0.2	0.3
-North Africa & the Middle East		1.1	0.2	0.2	0.2	0.2
+Oceania						1.2	0.4	0.5	0.5	0.5
+China						1.1	0.4	0.4	0.4	0.4
+Japan						-0.3	0.4	0.4	0.4	0.4
+East Asia					5.1	0.3	0.3	0.3	0.4
+Southeast Asia					2.8	0.3	0.3	0.4	0.3
+South Asia					3.1	0.3	0.3	0.3	0.3
+Canada						0.7	0.5	0.5	0.5	0.5
+United States					0.7	0.5	0.5	0.5	0.5
+Central America					2.5	0.3	0.3	0.4	0.3
+Brazil						1.2	0.5	0.5	0.5	0.5
+Rest of South America				2.1	0.3	0.3	0.3	0.3
+West & Central Europe 				0.9	0.4	0.4	0.4	0.4
+Russia						1.1	0.4	0.4	0.4	0.4
+Sub-Saharan Africa				1.5	0.4	0.4	0.4	0.4
+North Africa & the Middle East			2.0	0.4	0.4	0.4	0.4
 

### 2024-02-27 paper.Rmd
Update paper.Rmd

@@
 To understand the impact of potential regional and global afforestation actions, an important first step was creating a realistic baseline scenario that depicts the likely changes between over the decade under study: 2023-2033.
-This was done by taking into account key drivers of change like growth in population and GDP.
+Data for this were drawn from sources such as CEPII and the United Nations.
+Using this data, we were able to control for the impact of key drivers of change like growth in population and GDP.
 Without accounting for this, results would likely have been biased as they would have failed to account for the increasing pressure to convert forests into agricultural land.
-Data for this were drawn from sources such as CEPII and the United Nations.
 
@@
 
-Then, we examined what it would take given the pressures of a still growing global population with the concomitant increasing appetite for foodstuffs to end, on net, the practice of deforestation globally.
+Then, we examined what it would take given the pressures of a still growing global population and an increasing appetite for foodstuffs for the practice of deforestation to end, on net, globally.
 Results suggest that it would require a tax of roughly 70.3 percent to be assessed on the value of the converted land for the deforestation rate to fall globally to roughly zero.

### 2024-02-27 paper.Rmd
Update paper.Rmd

@@
 Then, we examined what it would take given the pressures of a still growing global population and an increasing appetite for foodstuffs for the practice of deforestation to end, on net, globally.
-Results suggest that it would require a tax of roughly 70.3 percent to be assessed on the value of the converted land for the deforestation rate to fall globally to roughly zero.
-The economic impact of such a policy measure would not be limited to the cost of the tax, it would also change impact food production and global trade.
+Results suggest that it would require a tax of roughly 70.3 percent to be assessed on the value of converted land for the deforestation rate to fall globally to roughly zero.
+The economic impact of such a policy measure would not be limited to the cost of the tax, it would also change impact food production and global trade flows.
 For example, with this policy in place, most regions of the world would see significant declines in the amount of agricultural land in 2033 relative to the baseline scenario where land conversion was not taxed.
-For the United States, agricultural land would be 2.1 percent less (Table 1).
+For the United States, there would be 2.1 percent less agricultural land (Table 1).
 Larger decreases would happen in other major agricultural production regions such as China (-2.5 percent), South Asia (-8.0 percent), Brazil (-8.5 percent), and Russia (-9.3 percent).
@@
 
-These reductions in the amount of agricultural land prompt changes in the quantity of agricultural goods produced for many regions.
-For example, East Asia and Southeast Asia see the largest declines in percentage terms in agricultural production with decreases of 5.1 percent and 2.0 percent, respectively (Table 2).
-Surprisingly, Brazil and the United States, despite the derease in available agricultural land compared to the baseline 2033 scenario, see increased agricultural output of 0.2 percent and 0.6 percent, respectively.
+These reductions in the amount of agricultural land would then impact the quantity of agricultural goods that could be produced.
+For example, the 14.6 percent decrease in agrciultural land in East Asia and the 9.0 percent decrease in agrciultural land in Southeast Asia would result in large declines in percentage terms in agricultural production of 5.1 percent and 2.0 percent, respectively (Table 2).
+Surprisingly, Brazil and the United States, despite the derease in available agricultural land compared to the baseline 2033 scenario, are estimated to see increased agricultural output of 0.2 percent and 0.6 percent, respectively.
 This may be tied to the sharp declines in other regions prompting greater production for export.
@@
 
-Overall, this policy, while avoiding deforestation, has a rather significant negative economic impact on all regions except Japan and Canada who see an increase in consumer well-being of $977.1 million and $166.9 million, respectively (table 4).
+Overall, this policy, while creating the positive result of avoiding deforestation, has a rather significant negative economic impact on all regions except Japan and Canada who see an increase in consumer well-being of $977.1 million and $166.9 million, respectively (table 4).
 Consumer well-being is a measurement of equivalent variation and here measures the benefit/harm consumers experience in terms of income as a result of the policy changes.
 The United States, along with 10 other regions, see well-being drop by more than $1 billion.
-The most negatively impacted regions are South Asia (-28.0 billion), China (-14.5 billion), and Southeast Asia (-13.6 billion).
-In total, such a policy is estimated to decrease global consumer well-being by $92.1 billion.
+The most negatively impacted regions are South Asia (-$28.0 billion), China (-$14.5 billion), and Southeast Asia (-$13.6 billion).
+In total, this tax-driven policy is estimated to decrease global consumer well-being in 2033 by $92.1 billion relative to the scenario where no tax on forestland conversion was in place.
 
@@
 Oceania				-169,625,000
-EUII				-588,500,000
 Brazil				-735,125,000
+Rest of South America		-3,848,625,000
 United States			-1,085,000,000
 Russia				-1,152,375,000
-Rest of South America		-3,848,625,000
 Central America			-4,471,500,000
@@
 Much of this loss in well-being may be the result of higher prices.
-Japan is the only nation that sees any positive price impacts resulting from this policy, with the price of agricultural goods decreasing 0.3 percent (table 5).
+Japan is the only nation that sees any price impacts that are beneficial for consumers resulting from this policy, with the price of agricultural goods decreasing 0.3 percent (table 5).
 All other regions, for both agricultural and non-agricultural products, see prices rise between 0.3 and 5.1 percent.
@@
 Land price seems weird. I have it decreasing????
+
+# Discussion
+
+This analysis shows that, at an estimated tax rate of 70.3 percent of the value of each acre of land that is converted to agricultural land, global net deforestation would drop to roughly zero.
+In the United States, as USDA-NASS (Land Values 2022 Summary (August 2022)) valued the average acre of farmland in 2022 at $3,800, the tax would be $2,671.4/acre.
+In Brazil, CEIC valued the average price of Brazilian land at roughly 2,500 Brazilian Reais, which would put the tax at roughly $203 per acre.
+Governments would very likely be wary of farmer discontent were they to apply, but they would also have to take into account the larger impacts on their economies.
+Restricting agricultural land, as shown above, would for the most part restrict output and increase prices.
+It is uncertain whether the gains from the increase in prices would be sufficient to offset the loss of potential revenue that would have come from expanded agricultural operations.
+As governments might be reluctant to impose such taxes on their own citizens, it is possible that policymakers who wished to see deforestation stopped could impose import tariffs equal to the tax rate to achieve the same result.
 

### 2024-02-27 paper.Rmd
Update paper.Rmd

@@
 
+We also show the utility of joining two major economic models.
+GTAP has long been valued for its ability to provide robust information on global trade flows and prices, and GTM has been quite helpful is understanding the impact of forestry decisions on the carbon sink and land rents.
+By combining these models through the use of R and allowing the models to pass inputs and outputs back and forth iteratively, the benefits of both models have been maintained and their weaknesses greatly minimized.
+This advance holds great promise for advancement on a variety of fronts to researchers and policymakers interested in climate change, agricultural trade, and their interdependencies.
+
 1a) On aggregate zero percent net deforestation Globally

### 2024-02-27 paper.Rmd
Update paper.Rmd

@@
 For example, as nations develop and disposable incomes rise, consumers often change to more meat-based diets.
-Population growth is also a driver and when it is especially high, which is the case at present throughout most of Africa, this also puts pressure on policymakers to allow greater conversion of forestland into farmland.
-Still, previous research has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017].
+Population growth is also a driver.
+Currently, the United Nations' medium variant population projections suggest that between 2023 and 2033 the worlds' population will grow by 8.8 percent, a gain of 707.1 million people [@un2024].
+And where population growth is especially high, which is the case at present throughout most of Africa (2023-2033 population growth: 25.0%: 360.6 million), this puts extreme pressure on policymakers to allow greater conversion of forestland into farmland.
+Still, some research has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017].
 

### 2024-02-27 paper.Rmd
Update paper.Rmd

@@
 
-Faced with competing land-use demands, afforestation efforts need to be targeted to the regions of the world where they will provide the most benefits.
-Afforestation in boreal zones has been shown to provide the smallest net benefits, as the carbon sequestrations benefits of conifer forests are mitigated by the reduction in albedo in winter [@mykleby2017] [@li2015].
+Afforestation efforts are also thought to not provide equal benefits in every region.
+For example, afforestation in boreal zones has been shown to provide the smallest net benefits, as the carbon sequestrations benefits of conifer forests are mitigated by the reduction in albedo in winter [@mykleby2017] [@li2015].
 Coniferous forests have also been shown to increase soil organic carbon less than broadleaf forests [@guo2021].
@@
 
-South America and Sub-Saharan Africa are two regions that show much potential for afforestation.
+South America and Sub-Saharan Africa are thus two regions that show much potential for afforestation.
 It is estimated that these two regions hold at least 50% of the potential global gain while regions such as Northern Africa and the Middle East show little promise as forest growth rates are quite low [@doelman2020].
@@
 There are other nations that have a big potential for gains from afforestation, but they also face challenges.
-India, for example, has a great potential to increase its carbon sequestration through afforestation, but due to its massive population and food security worries, its NDC does not entertain the idea of agricultural land being repurposed and instead primarily considers targeting wastelands that may not be well suited for afforestation [@amjath2019].
+India, for example, has a great potential to increase its carbon sequestration through afforestation, but its massive population and food security worries limit its ability to consider such actions.
+The United Nations suggests that India's population will grow slightly less fast than the world average between 2023 and 2033 (+8.5%), but due to its sizeable base, that will equate to an increase in population of 120.5 million.
+Thus, its NDC does not entertain the idea of agricultural land being repurposed and instead primarily considers targeting wastelands that may not be well suited for afforestation [@amjath2019].
 Moreover, as land-use decisions in one country also tend to have impacts that spillover into many others [@meyfroidt2010] [@lambin2011], analysis needs to be conducted at a global scale.

### 2024-02-27 paper.Rmd
Update paper.Rmd

@@
 
-Steinbuks and Hertel [@steinbuks2012] developed the FABLE model (Forestry, Agriculture, Biofuels, Land Use, and Environment) as a global partial equilibrium model to analyze optimal land use within a context of increasing demand for food, bioenergy, forest products, and non-provisioning ecosystem services, while also meeting greenhouse gas targets.
-The model encompasses nine sectors, including agriculture, livestock, food processing, biofuels, energy, forest, timber processing, and ecosystem services.
-While FABLE is a powerful tool for examining the optimal trajectory of various land uses under specific demand assumptions, it primarily focuses on specific sectors of the economy, neglecting general equilibrium effects and socio-economic or household welfare impacts [@steinbuks2012] [@hertel2017].
+Toward this end, Steinbuks and Hertel [@steinbuks2012] developed their global partial equilibrium FABLE model (Forestry, Agriculture, Biofuels, Land Use, and Environment) to identify the optimal land use choices given the competing demands of meeting greenhouse gas targets and satisfying demand for food, bioenergy, forest products, and ecosystem services.
+While FABLE is a powerful tool for examining the optimal trajectory of various land uses under specific demand assumptions, it is focused on specific sectors of the economy and neglects general equilibrium and household welfare effects [@steinbuks2012] [@hertel2017].
 Other efforts have used the Global Trade Analysis Project (GTAP) computable general equilibrium model to address such questions, but its standard model does not well account for land-use changes.
-As a result, the Land Use and Land Cover Database within the GTAP framework, which incorporates forestry remote-sensing products, was created [@baldos2012].
-It was further strengthened as part of the GTAP Agro-Ecological Zone (GTAP-AEZ) model, with the use of spatially explicit global land use data and through the incorporation of intra- and inter-regional land and land-based greenhouse emissions heterogeneity [@hertel2008].
+This prompted the creation of the Land Use and Land Cover Database within the GTAP framework, which incorporates forestry remote-sensing products [@baldos2012].
+The GTAP Agro-Ecological Zone (GTAP-AEZ) model further strengthened the GTAP model's ability to analyze land use changes, with the use of spatially explicit global land use data and through the incorporation of intra- and inter-regional land and land-based greenhouse emissions heterogeneity [@hertel2008].
 While the GTAP-AEZ model benefits from the inclusion of competition among different crops, grazing, and forest-based uses, it has limitations, primarily due to its infrequent updates.
-The KLUM@GTAP framework integrated the Kleines Land Use Model (KLUM) with an extended yet static version of GTAP called GTAP-EFL to evaluate the impacts of climate change on cropland allocation.
+The KLUM@GTAP framework integrated the Kleines Land Use Model (KLUM) with an extended yet static version of GTAP called GTAP-EFL to evaluate the impact of climate change on cropland allocation.
 GTAP-EFL separates energy factors from intermediate inputs and incorporates them into capital, while also considering CO2 emissions.
@@
 For example, the dynamic GTAP model, supported by the GTAP-AEZ database, considers land market effects, which were identified as significant in driving results by Stevenson et al.
-[@stevenson2013], and has been utilized to evaluate the impacts of crop intensification on land use.
+[@stevenson2013].
+This model has been utilized to evaluate the impact of crop intensification on land use.
 It, however, has been critiqued, as global aggregates may mask localized shifts that can have implications for ecologically significant areas [@byerlee2014].
@@
 As the model applies nested CES production functions to all sectors and differentiates them by parameters only, to model one sector, in this case forestry, differently from the rest would require a major redesign of the model.
-While not impossible, this project will attempt to get around these concerns and assess the impact of various likely land-use scenarios by melding GTAP with the Global Timber Model (GTM), which is a dynamic optimization model that has been specifically designed to analyze the relationship among land rents, forested land cover, and carbon sequestration through forests and thus is better able to examine the impact of afforestation or avoided deforestation on the carbon sink [@sohngen2001].
+While not impossible, this project will attempt to get around these concerns and assess the impact of various likely land-use scenarios by melding the dynamic GTAP model with the Global Timber Model (GTM), which is a dynamic optimization model that has been specifically designed to analyze the relationship among land rents, forested land cover, and carbon sequestration through forests and thus is better able to examine the impact of afforestation or avoided deforestation on the carbon sink [@sohngen2001].
 The latest versions of both models were used.
-For the GTAP model, this meant using the 11b data which has 2017 as a base year.
+For the GTAP model, this meant using data which has 2017 as its base year.
 This data was then updated to the base year (2023) using data from the Centre for Prospective Studies and International Information (CEPII) to account for the projected changes in gross domestic product (GDP), population, and productivity for each country in the world.
+Similar steps were taken to update and recalibrate the GTM model.
 

### 2024-02-27 paper.Rmd
Update paper.Rmd

@@
 
-By joining these two models and passing outputs back and forth between the models, any limitations of either model should be minimized and more robust results, therefore, should be attained.
-Establishing a connection between the standard GTAP model and the GTM model required  a two-step process.
-First, we identified the set of common variables that are included in both models and their closures.
-That is, we selected which variables would be determined endogenously by one model and held as exogenous in the other model.
-In our case, we modeled the return on land using the GTAP model and fed those results into the GTM model as exogenous.
-Then, the amount of land available for crop and livestock production was endogenously determined by the GTM model and fed into the GTAP model as an exogenous input.
+The rationale for joining these two models is that, by passing outputs back and forth between the models until an dual-state equilibrium is achieved, any limitations of either model should be minimized and more robust results should, therefore, be attained.
+Establishing a connection between the  GTAP model and the GTM model required a two-step process.
+First, we had to identify the set of variables that are common to both models and their closures.
+From those, we selected which variables would be determined endogenously by one model and held as exogenous in the other model.
+In our case, we decided to model the return on land using the GTAP model and feed those results into the GTM model as exogenous.
+The GTM model then endogenously determined the amount of land available for crop and livestock production, and we fed that back into the GTAP model as an exogenous input.
 In the second step, we established the connection between the models on the software level.
-This step required that the GTAP model, which is written in GEMPACK, be able to handle with the inputs from the GTM model, which is written in GAMS, and submit the outputs back to the GTM model.
-As GEMPACK and GAMS were developed long before software integration became important or even possible, we had to create a set of user functions in R that executed both models, collected their outputs, and created parameter files automatically.
+This step required that the GTAP model, which is written in GEMPACK, be able to handle inputs from the GTM model, which is written in GAMS, and submit the outputs back to the GTM model.
+As GEMPACK and GAMS were developed long before software integration became important or even possible, we settled on the use of R [@R2021] to create the user functions that executed both models, collected their outputs, and automatically created parameter files.
 
-Once R was able to operate both models automatically, we were then able to formulate the combined solution as a fairly simple math program that is solved in R:
+Once R was able to control the interactions between both models, we were then able to formulate the combined solution as a fairly simple math program that is solved in R:
 

### 2024-02-27 paper.Rmd
Update paper.Rmd

@@
 For the GTAP model, this meant using data which has 2017 as its base year.
-This data was then updated to the base year (2023) using data from the Centre for Prospective Studies and International Information (CEPII) to account for the projected changes in gross domestic product (GDP), population, and productivity for each country in the world.
+This data was then updated to the base year (2023) using data from the Centre for Prospective Studies and International Information (CEPII) [@CEPII2024] to account for the projected changes in gross domestic product (GDP), population, and productivity for each country in the world.
 Similar steps were taken to update and recalibrate the GTM model.
@@
 
-To understand the impact of potential regional and global afforestation actions, an important first step was creating a realistic baseline scenario that depicts the likely changes between over the decade under study: 2023-2033.
-Data for this were drawn from sources such as CEPII and the United Nations.
+To understand the impact of potential regional and global afforestation actions, an important first step was creating a realistic baseline scenario that depicts the likely changes over the decade under study: 2023-2033.
+Data for this were drawn from sources including CEPII and the United Nations.
 Using this data, we were able to control for the impact of key drivers of change like growth in population and GDP.

### 2024-02-27 paper.Rmd
Update paper.Rmd

@@
 For example, the 14.6 percent decrease in agrciultural land in East Asia and the 9.0 percent decrease in agrciultural land in Southeast Asia would result in large declines in percentage terms in agricultural production of 5.1 percent and 2.0 percent, respectively (Table 2).
-Surprisingly, Brazil and the United States, despite the derease in available agricultural land compared to the baseline 2033 scenario, are estimated to see increased agricultural output of 0.2 percent and 0.6 percent, respectively.
-This may be tied to the sharp declines in other regions prompting greater production for export.
+Surprisingly, Brazil and the United States, despite their dereases in available agricultural land compared to the baseline 2033 scenario, are estimated to see increased agricultural output of 0.2 percent and 0.6 percent, respectively.
+This may be tied to the sharp declines in other regions prompting greater intensity of production to satisfy external demand.
 
@@
 
-When we examine how this impacts exports of agricultural goods, we find that Japan and the United States appear to be the biggest beneficiaries of this policy with their exports increasing 8.8 percent and 3.5 percent, respectively (table 3).
+When we examine how this impacts exports of agricultural goods, we find that exports from Japan and the United States grow the most, increasing 8.8 percent and 3.5 percent, respectively (table 3).
 The regions where agricultural exports drop most significantly are East Asia, South Asia, Southeast Asia, and Central America who see decreases of 14.5 percent, 5.9 percent, 5.1 percent,and 5.0 percent, respectively.
@@
 
-Overall, this policy, while creating the positive result of avoiding deforestation, has a rather significant negative economic impact on all regions except Japan and Canada who see an increase in consumer well-being of $977.1 million and $166.9 million, respectively (table 4).
+Overall, this tax policy, which results in zero net deforestation globally, has a rather significant negative economic impact on all regions except Japan and Canada who see an increase in consumer well-being of $977.1 million and $166.9 million, respectively (table 4).
 Consumer well-being is a measurement of equivalent variation and here measures the benefit/harm consumers experience in terms of income as a result of the policy changes.

### 2024-02-27 paper.Rmd
Update paper.Rmd

@@
 
-When examining the likely impacts of the baseline changes alone, it is estimated that by 2033, significant deforestation would have occured in many regions of the world. For example, in Brazil...
+When examining the likely impacts of the baseline changes alone, it is estimated that between 2023 and 2033, 144.2 million hectares (3.6%) of global forests will be converted to agricultural lands with the largest decline expected to happen in Sub-Saharan Africa, which is estimated to lose 39.7 million hectares (7.0%) of its forestland (Table 1).
+Southeast Asia, Brazil, and the rest of South America also see forestland shrink significantly, with decreases of 22.6 million hectares (9.3%), 19.0 million hectares (3.8%), and 23.5 million hectares (6.7%), respectively.
+Conversely, the United States is estimated to add 1.3 million hectares (0.5%) to its forestland.
 
-Then, we examined what it would take given the pressures of a still growing global population and an increasing appetite for foodstuffs for the practice of deforestation to end, on net, globally.
-Results suggest that it would require a tax of roughly 70.3 percent to be assessed on the value of converted land for the deforestation rate to fall globally to roughly zero.
-The economic impact of such a policy measure would not be limited to the cost of the tax, it would also change impact food production and global trade flows.
+Table 1. Forestland by region in 2023 and estimated change by 2033
+	Russia	Sub-Saharan Africa Brazil Canada Rest of South America	United States Southeast Asia Oceania EU	China Central America South Asia Other Europe North Africa & the Middle East Japan East Asia Total
+Forestland in 2023	838.1	564.7	498.2	412.8	348.5	248.4	243.9	199.2	186.7	164.6	94.6	47.8	40.2	33.8	23.7	14.5	3,959.5
+Forestland in 2033	828.7	525.0	479.2	407.9	325.0	249.7	221.3	196.1	175.5	164.8	91.2	42.6	36.1	30.9	26.9	14.6	3,815.3
+∆ (ha)			-9.5	-39.7	-19.0	-5.0	-23.5	1.3	-22.6	-3.1	-11.1	0.1	-3.4	-5.3	-4.1	-2.9	3.2	0.1	-144.2
+∆ (%)			-1.1	-7.0	-3.8	-1.2	-6.7	0.5	-9.3	-1.5	-6.0	0.1	-3.6	-11.0	-10.1	-8.5	13.7	0.7	-3.6
+
+Against this backdrop, we examined what it would take to end, on net, the practice of deforestation globally.
+Results suggest that it would require a tax of roughly 70.3 percent to be assessed on the value of converted land for the deforestation rate to fall globally to zero.
+The economic impact of such a policy measure would not be limited to the cost of the tax, it would also impact food production and global trade flows.
 For example, with this policy in place, most regions of the world would see significant declines in the amount of agricultural land in 2033 relative to the baseline scenario where land conversion was not taxed.
-For the United States, there would be 2.1 percent less agricultural land (Table 1).
+For the United States, there would be 2.1 percent less agricultural land (Table 2).
 Larger decreases would happen in other major agricultural production regions such as China (-2.5 percent), South Asia (-8.0 percent), Brazil (-8.5 percent), and Russia (-9.3 percent).
 
-Table 1 qo-Land
+Table 2. Percentage change in the amount of agricultural land in 2033 with deforestation mitigation policy in place (Source: GTAP qo-Land)
 				∆ (%)
@@
 These reductions in the amount of agricultural land would then impact the quantity of agricultural goods that could be produced.
-For example, the 14.6 percent decrease in agrciultural land in East Asia and the 9.0 percent decrease in agrciultural land in Southeast Asia would result in large declines in percentage terms in agricultural production of 5.1 percent and 2.0 percent, respectively (Table 2).
+For example, the 14.6 percent decrease in agrciultural land in East Asia and the 9.0 percent decrease in agrciultural land in Southeast Asia would result in large declines in percentage terms in agricultural production of 5.1 percent and 2.0 percent, respectively (Table 3).
 Surprisingly, Brazil and the United States, despite their dereases in available agricultural land compared to the baseline 2033 scenario, are estimated to see increased agricultural output of 0.2 percent and 0.6 percent, respectively.
@@
 
-Table 2 qo-Ag
+Table 3. Percentage change in the amount of agricultural land in 2033 with deforestation mitigation policy in place (Source: GTAP qo-Land)qo-Ag
 				∆ (%)
@@
 
-When we examine how this impacts exports of agricultural goods, we find that exports from Japan and the United States grow the most, increasing 8.8 percent and 3.5 percent, respectively (table 3).
+When we examine how this impacts exports of agricultural goods, we find that exports from Japan and the United States grow the most, increasing 8.8 percent and 3.5 percent, respectively (Table 4).
 The regions where agricultural exports drop most significantly are East Asia, South Asia, Southeast Asia, and Central America who see decreases of 14.5 percent, 5.9 percent, 5.1 percent,and 5.0 percent, respectively.
 
-Table 3 qxw-Ag
+Table 4. Percentage change in the amount of agricultural land in 2033 with deforestation mitigation policy in place (Source: GTAP qo-Land)qxw-Ag
 				∆ (%)

### 2024-02-27 paper.Rmd
Update paper.Rmd

@@
 Table 1. Forestland by region in 2023 and estimated change by 2033
-	Russia	Sub-Saharan Africa Brazil Canada Rest of South America	United States Southeast Asia Oceania EU	China Central America South Asia Other Europe North Africa & the Middle East Japan East Asia Total
+	Russia	Sub-Saharan Africa Brazil Canada Rest of South America	United States Southeast Asia Oceania Western & Central Europe China Central America South Asia Other Europe North Africa & the Middle East Japan East Asia Total
 Forestland in 2023	838.1	564.7	498.2	412.8	348.5	248.4	243.9	199.2	186.7	164.6	94.6	47.8	40.2	33.8	23.7	14.5	3,959.5
@@
 ∆ (%)			-1.1	-7.0	-3.8	-1.2	-6.7	0.5	-9.3	-1.5	-6.0	0.1	-3.6	-11.0	-10.1	-8.5	13.7	0.7	-3.6
+Note: Southeast Asia includes Brunei Darussalam, Cambodia, Indonesia, Laos, Malaysia, Myanmar, Philippines, Singapore, Thailand, Timor-Leste, and Vietnam.
+Oceania includes Australia, Cook Islands, Fiji, French Polynesia, Kiribati, Marshall Islands, Micronesia, Nauru, New Caledonia, New Zealand, Palau, Papua New Guinea, Samoa (Western Samoa), Solomon Islands, Tonga, Tuvalu, and Vanuatu.
+Rest of South America includes
+Argentina, Bolivia, Czech Republic, Chile, Colombia, Ecuador, the Falkland Islands, French Guiana, Guyana, Paraguay, Peru, South Georgia, the South Sandwich Islands, Suriname, Uruguay, and Venezuela.
+Western & Central Europe includes  Austria, Belgium, Bulgaria, Croatia, Denmark, Estonia, Finland, France, Hungary, Iceland, Italy, Ireland, Germany, Greece, Latvia, Liechtenstein, Lithuania, Luxembourg, Malta, Monaco, Montenegro, Netherlands, North Macedonia, Norway, Poland, Portugal, Romania, Serbia, Slovakia, Slovenia, Spain, Sweden, Switzerland, Turkey, and the United Kingdom.
+Other Europe includes Albania, Armenia, Azerbaijan, Belarus, Bosnia and Herzegovina, Georgia,  Kosovo Kazakhstan, Kyrgyzstan, Moldova, Tajikistan, and Ukraine.
 

### 2024-02-27 paper.Rmd
Update paper.Rmd

@@
 
-Table 3. Percentage change in the amount of agricultural land in 2033 with deforestation mitigation policy in place (Source: GTAP qo-Land)qo-Ag
+Table 3. Percentage change in the amount of agricultural goods produced in 2033 with deforestation mitigation policy in place (Source: GTAP qo-Ag)
 				∆ (%)
@@
 
-Table 4. Percentage change in the amount of agricultural land in 2033 with deforestation mitigation policy in place (Source: GTAP qo-Land)qxw-Ag
+Table 4. Percentage change in the amount of agricultural exports in 2033, by region, with deforestation mitigation policy in place (Source: GTAP qxw-Ag)
 				∆ (%)
@@
 
-Overall, this tax policy, which results in zero net deforestation globally, has a rather significant negative economic impact on all regions except Japan and Canada who see an increase in consumer well-being of $977.1 million and $166.9 million, respectively (table 4).
+Overall, this tax policy, which results in zero net deforestation globally, has a rather significant negative economic impact on all regions except Japan and Canada who see an increase in consumer well-being of $977.1 million and $166.9 million, respectively (Table 5).
 Consumer well-being is a measurement of equivalent variation and here measures the benefit/harm consumers experience in terms of income as a result of the policy changes.
@@
 
-Table 4 EV
+Table 5. Change in consumer well-being in 2033, by region, in dollars, with deforestation mitigation policy in place (Source: GTAP EV) 
 				∆ ($)
@@
 Much of this loss in well-being may be the result of higher prices.
-Japan is the only nation that sees any price impacts that are beneficial for consumers resulting from this policy, with the price of agricultural goods decreasing 0.3 percent (table 5).
+Japan is the only nation that sees any price impacts that are beneficial for consumers resulting from this policy, with the price of agricultural goods decreasing 0.3 percent (Table 6).
 All other regions, for both agricultural and non-agricultural products, see prices rise between 0.3 and 5.1 percent.
 
-Table 5 pm
+Table 6. Change in market price of various goods in 2033, by region, with deforestation mitigation policy in place (Source: GTAP pm) 
 									∆ (%)				

### 2024-02-27 paper.Rmd
Update paper.Rmd

@@
 We also show the utility of joining two major economic models.
-GTAP has long been valued for its ability to provide robust information on global trade flows and prices, and GTM has been quite helpful is understanding the impact of forestry decisions on the carbon sink and land rents.
+GTAP has long been valued for its ability to provide robust information on global trade flows and prices, and GTM has proven its utility in understanding the impact of forestry decisions on the carbon sink and land rents.
 By combining these models through the use of R and allowing the models to pass inputs and outputs back and forth iteratively, the benefits of both models have been maintained and their weaknesses greatly minimized.
-This advance holds great promise for advancement on a variety of fronts to researchers and policymakers interested in climate change, agricultural trade, and their interdependencies.
+This advance holds great promise for advancement on a variety of fronts for researchers and policymakers interested in climate change, agricultural trade, and their interdependencies.
+
+Unused
 

### 2024-02-27 paper.Rmd
My changes

@@
 
-```{r setup, include=FALSE}
+```{r setup, include=FALSE, cache = TRUE, echo = FALSE}
 knitr::opts_chunk$set(echo = TRUE)
@@
 
-The global gains that can be realized from afforestation, however, are uncertain, with broad estimates of cumulative sequestration ranging from 42 gigatonnes to 700 gigatonnes by the end of the century [@veldman2019]  [@bastin2019]  [@humpenoder2014]  [@sathaye2006]  [@doelman2020].
-In addition, the amount of land conversion that would be required also varies greatly, with estimates ranging from 0.3 billion hectares to 2.8 billion hectares [@veldman2019]  [@bastin2019]  [@humpenoder2014]  [@doelman2020].
+The global gains that can be realized from afforestation, however, are uncertain, with broad estimates of cumulative sequestration ranging from 42 gigatonnes to 700 gigatonnes by the end of the century [@veldman2019; @bastin2019; @humpenoder2014; @sathaye2006; @doelman2020].
+In addition, the amount of land conversion that would be required also varies greatly, with estimates ranging from 0.3 billion hectares to 2.8 billion hectares [@veldman2019; @bastin2019; @humpenoder2014; @doelman2020].
 In many cases, these results suggest massive losses in the amount of pasture and cropland that would be available.
@@
 Currently, the United Nations' medium variant population projections suggest that between 2023 and 2033 the worlds' population will grow by 8.8 percent, a gain of 707.1 million people [@un2024].
-And where population growth is especially high, which is the case at present throughout most of Africa (2023-2033 population growth: 25.0%: 360.6 million), this puts extreme pressure on policymakers to allow greater conversion of forestland into farmland.
+And where population growth is especially high, which is the case at present throughout most of Africa (2023--2033 population growth: 25.0%: 360.6 million), this puts extreme pressure on policymakers to allow greater conversion of forestland into farmland.
 Still, some research has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017].
@@
 Afforestation efforts are also thought to not provide equal benefits in every region.
-For example, afforestation in boreal zones has been shown to provide the smallest net benefits, as the carbon sequestrations benefits of conifer forests are mitigated by the reduction in albedo in winter [@mykleby2017] [@li2015].
+For example, afforestation in boreal zones has been shown to provide the smallest net benefits, as the carbon sequestrations benefits of conifer forests are mitigated by the reduction in albedo in winter [@mykleby2017; @li2015].
 Coniferous forests have also been shown to increase soil organic carbon less than broadleaf forests [@guo2021].
@@
 India, for example, has a great potential to increase its carbon sequestration through afforestation, but its massive population and food security worries limit its ability to consider such actions.
-The United Nations suggests that India's population will grow slightly less fast than the world average between 2023 and 2033 (+8.5%), but due to its sizeable base, that will equate to an increase in population of 120.5 million.
-Thus, its NDC does not entertain the idea of agricultural land being repurposed and instead primarily considers targeting wastelands that may not be well suited for afforestation [@amjath2019].
-Moreover, as land-use decisions in one country also tend to have impacts that spillover into many others [@meyfroidt2010] [@lambin2011], analysis needs to be conducted at a global scale.
+The United Nations suggests that India's population will grow slightly less fast than the world average between 2023 and 2033 (+8.5%), but due to its sizable base, that will equate to an increase in population of 120.5 million.
+Thus, its NDC does not entertain the idea of agricultural land being re-purposed and instead primarily considers targeting wastelands that may not be well suited for afforestation [@amjath2019].
+Moreover, as land-use decisions in one country also tend to have impacts that spillover into many others [@meyfroidt2010; @lambin2011], analysis needs to be conducted at a global scale.
 
-Toward this end, Steinbuks and Hertel [@steinbuks2012] developed their global partial equilibrium FABLE model (Forestry, Agriculture, Biofuels, Land Use, and Environment) to identify the optimal land use choices given the competing demands of meeting greenhouse gas targets and satisfying demand for food, bioenergy, forest products, and ecosystem services.
-While FABLE is a powerful tool for examining the optimal trajectory of various land uses under specific demand assumptions, it is focused on specific sectors of the economy and neglects general equilibrium and household welfare effects [@steinbuks2012] [@hertel2017].
+Toward this end, @steinbuks2012 developed their global partial equilibrium FABLE model (Forestry, Agriculture, Biofuels, Land Use, and Environment) to identify the optimal land use choices given the competing demands of meeting greenhouse gas targets and satisfying demand for food, bioenergy, forest products, and ecosystem services.
+While FABLE is a powerful tool for examining the optimal trajectory of various land uses under specific demand assumptions, it is focused on specific sectors of the economy and neglects general equilibrium and household welfare effects [@steinbuks2012; @hertel2017].
 Other efforts have used the Global Trade Analysis Project (GTAP) computable general equilibrium model to address such questions, but its standard model does not well account for land-use changes.
@@
 More recent modeling attempts of land-use change have begun to move away from comparative static analysis.
-For example, the dynamic GTAP model, supported by the GTAP-AEZ database, considers land market effects, which were identified as significant in driving results by Stevenson et al.
-[@stevenson2013].
+For example, the dynamic GTAP model, supported by the GTAP-AEZ database, considers land market effects, which were identified as significant in driving results by @stevenson2013.
 This model has been utilized to evaluate the impact of crop intensification on land use.
@@
 
-# Modeling the impact of possible measures to mitigate deforestation/promote afforestation
+# Deforestation under the baseline projections
 
-To understand the impact of potential regional and global afforestation actions, an important first step was creating a realistic baseline scenario that depicts the likely changes over the decade under study: 2023-2033.
-Data for this were drawn from sources including CEPII and the United Nations.
-Using this data, we were able to control for the impact of key drivers of change like growth in population and GDP.
-Without accounting for this, results would likely have been biased as they would have failed to account for the increasing pressure to convert forests into agricultural land.
+To understand the impact of potential regional and global afforestation actions, an important first step was creating a realistic baseline scenario that depicts the likely changes over the decade under study, 2023--2033.
+Data for this were drawn from CEPII [@CEPII2024].
+To build our scenario, we used two key projected variables by CEPII: population and GDP.
+The CEPII projections do not contain any information on land use.
+To build a more complete baseline, we also include the projections on forested land change included in the Global Timber Model (GTM).
+Using the GTAP model in connection with the GTM model (to assure that the changes in land are consistent between the two models), we calculate the overall change in factor quantity/productivity growth required to achieve the targeted values in 2033.
+In Table \@ref(tab:baseline-growth) we show the projected population and GDP growths per region.
 
-When examining the likely impacts of the baseline changes alone, it is estimated that between 2023 and 2033, 144.2 million hectares (3.6%) of global forests will be converted to agricultural lands with the largest decline expected to happen in Sub-Saharan Africa, which is estimated to lose 39.7 million hectares (7.0%) of its forestland (Table 1).
+```{r echo = FALSE}
+regLabels = c(
+  "US" = 'US',
+  "CHINA" = 'China',
+  "BRAZIL" = 'Brazil',
+  "CANADA" = 'Canada',
+  "RUSSIA" = 'Russia',
+  "EU ANNEX I" = 'EU I',
+  "EU NON ANNEX I" = 'EU II',
+  "SOUTH ASIA" = 'South Asia',
+  "CENT AMER." = 'Central America',
+  "RSAM" = 'Rest of Latin America',
+  "SSAF" = 'Sub-Saharan Africa',
+  "SE ASIA" = 'Southeast Asia',
+  "OCEANIA" = 'Oceania',
+  "JAPAN" = 'Japan',
+  "AFME" = 'North Africa/Middle East',
+  "E ASIA" = 'East Asia'
+)
+
+gtapCountries = HARr::read_har('asset/gtap/flexagg/gtm/gsdgset.har')$h1
+
+gtmCountries = c(
+  'US',
+  'CHINA',
+  'BRAZIL',
+  'CANADA',
+  'RUSSIA',
+  'EU ANNEX I',
+  'EU NON ANNEX I',
+  'SOUTH ASIA',
+  'CENT AMER.',
+  'RSAM',
+  'SSAF',
+  'SE ASIA',
+  'OCEANIA',
+  'JAPAN',
+  'AFME',
+  'E ASIA'
+)
+
+gGTAP = gtapCountries
+names(gGTAP)=c("OCEANIA","CHINA","JAPAN","E ASIA", "SE ASIA", "SOUTH ASIA", "CANADA","US",  "CENT AMER.","BRAZIL", "RSAM",  "EU ANNEX I", 
+               "EU NON ANNEX I",  "RUSSIA", "SSAF", "AFME", "ROW")
+
+gGTM = names(gGTAP)
+names(gGTM)= gGTAP
+
+baselineGTAP = HARr::read_SL4('asset/GTAP/run_baseline.sl4')
+
+out = cbind('Real GDP (\\%)' = baselineGTAP$qgdp[,'TOTAL'],'Population (\\%)' = baselineGTAP$pop[,'TOTAL'],'Agricultural land (\\%)' = baselineGTAP$qo['land',,'TOTAL'])
+
+rownames(out) = regLabels[gGTM[rownames(out)]]
+
+rownames(out)[length(rownames(out))] = 'ROW'
+
+kableExtra::column_spec(
+  kableExtra::column_spec(kableExtra::kable_styling(
+    knitr::kable(out, caption = 'GDP, population, and agricultural land growth under the baseline scenario (Source: CEPII and GTM)', digits = 1, escape = FALSE, label = 'baseline-growth')
+  ), 1, border_left = TRUE),
+  4,
+  border_right = TRUE
+)
+
+```
+
+Focusing on the absolute area of the forested areas, as generated by the GTM model, we estimate that between 2023 and 2033, 144.2 million hectares (3.6%) of global forests will be converted to agricultural lands with the largest decline expected to happen in Sub-Saharan Africa, which is estimated to lose 39.7 million hectares (7.0%) of its forestland (Table \@ref(tab:baseline-forest-change)).
 Southeast Asia, Brazil, and the rest of South America also see forestland shrink significantly, with decreases of 22.6 million hectares (9.3%), 19.0 million hectares (3.8%), and 23.5 million hectares (6.7%), respectively.
@@
 
-Table 1. Forestland by region in 2023 and estimated change by 2033
-	Russia	Sub-Saharan Africa Brazil Canada Rest of South America	United States Southeast Asia Oceania Western & Central Europe China Central America South Asia Other Europe North Africa & the Middle East Japan East Asia Total
-Forestland in 2023	838.1	564.7	498.2	412.8	348.5	248.4	243.9	199.2	186.7	164.6	94.6	47.8	40.2	33.8	23.7	14.5	3,959.5
-Forestland in 2033	828.7	525.0	479.2	407.9	325.0	249.7	221.3	196.1	175.5	164.8	91.2	42.6	36.1	30.9	26.9	14.6	3,815.3
-∆ (ha)			-9.5	-39.7	-19.0	-5.0	-23.5	1.3	-22.6	-3.1	-11.1	0.1	-3.4	-5.3	-4.1	-2.9	3.2	0.1	-144.2
-∆ (%)			-1.1	-7.0	-3.8	-1.2	-6.7	0.5	-9.3	-1.5	-6.0	0.1	-3.6	-11.0	-10.1	-8.5	13.7	0.7	-3.6
-Note: Southeast Asia includes Brunei Darussalam, Cambodia, Indonesia, Laos, Malaysia, Myanmar, Philippines, Singapore, Thailand, Timor-Leste, and Vietnam.
+```{r echo = FALSE}
+
+baseline = readRDS('history_baseline.rds')
+
+out = baseline$gtmResult$forests[1:2,]
+
+rownames(out) = c('Forestland in 2023','Forestland in 2033')
+
+out = rbind(out, `$\\Delta$ (Mha)` = out[2,]-out[1,],`$\\Delta$ (\\%)` = 100*(out[2,]/out[1,]-1))
+colnames(out) = regLabels[colnames(out)]
+
+out = cbind(out, Total = rowSums(out))
+
+out[4,'Total'] = (out[2 , 'Total']/out[1 , 'Total']-1)*100
+
+kableExtra::column_spec(
+  kableExtra::column_spec(
+    kableExtra::kable_styling(
+      kableExtra::add_footnote(
+        knitr::kable(
+          t(out),
+          escape = FALSE,
+          format = 'latex',
+          digits = 1,
+          label = 'baseline-forest-change',
+          caption = 'Forestland by region in 2023 and estimated change by 2033'
+        ),
+        label = c(
+          'a' = 'Southeast Asia includes Brunei Darussalam, Cambodia, Indonesia, Laos, Malaysia, Myanmar, Philippines, Singapore, Thailand, Timor-Leste, and Vietnam.',
+          'b' = 'Oceania includes Australia, Cook Islands, Fiji, French Polynesia, Kiribati, Marshall Islands, Micronesia, Nauru, New Caledonia, New Zealand, Palau, Papua New Guinea, Samoa (Western Samoa), Solomon Islands, Tonga, Tuvalu, and Vanuatu.',
+          'c' = 'Rest of South America includes Argentina, Bolivia, Czech Republic, Chile, Colombia, Ecuador, the Falkland Islands, French Guiana, Guyana, Paraguay, Peru, South Georgia, the South Sandwich Islands, Suriname, Uruguay, and Venezuela.',
+          d = 'Western & Central Europe includes  Austria, Belgium, Bulgaria, Croatia, Denmark, Estonia, Finland, France, Hungary, Iceland, Italy, Ireland, Germany, Greece, Latvia, Liechtenstein, Lithuania, Luxembourg, Malta, Monaco, Montenegro, Netherlands, North Macedonia, Norway, Poland, Portugal, Romania, Serbia, Slovakia, Slovenia, Spain, Sweden, Switzerland, Turkey, and the United Kingdom.',
+          e = 'Other Europe includes Albania, Armenia, Azerbaijan, Belarus, Bosnia and Herzegovina, Georgia,  Kosovo Kazakhstan, Kyrgyzstan, Moldova, Tajikistan, and Ukraine.'
+        ),
+        threeparttable  = TRUE
+      )
+    ),
+    1,
+    border_left = TRUE
+  ),
+  5,
+  border_right = TRUE
+)
+
+```
+
+<!-- Table 1. Forestland by region in 2023 and estimated change by 2033 -->
+<!-- 	Russia	Sub-Saharan Africa Brazil Canada Rest of South America	United States Southeast Asia Oceania Western & Central Europe China Central America South Asia Other Europe North Africa & the Middle East Japan East Asia Total  -->
+<!-- Forestland in 2023	838.1	564.7	498.2	412.8	348.5	248.4	243.9	199.2	186.7	164.6	94.6	47.8	40.2	33.8	23.7	14.5	3,959.5 -->
+<!-- Forestland in 2033	828.7	525.0	479.2	407.9	325.0	249.7	221.3	196.1	175.5	164.8	91.2	42.6	36.1	30.9	26.9	14.6	3,815.3 -->
+<!-- $\Delta$ (ha)			-9.5	-39.7	-19.0	-5.0	-23.5	1.3	-22.6	-3.1	-11.1	0.1	-3.4	-5.3	-4.1	-2.9	3.2	0.1	-144.2 -->
+<!-- $\Delta$ (%)			-1.1	-7.0	-3.8	-1.2	-6.7	0.5	-9.3	-1.5	-6.0	0.1	-3.6	-11.0	-10.1	-8.5	13.7	0.7	-3.6 -->
+<!-- Note: Southeast Asia includes Brunei Darussalam, Cambodia, Indonesia, Laos, Malaysia, Myanmar, Philippines, Singapore, Thailand, Timor-Leste, and Vietnam.
 Oceania includes Australia, Cook Islands, Fiji, French Polynesia, Kiribati, Marshall Islands, Micronesia, Nauru, New Caledonia, New Zealand, Palau, Papua New Guinea, Samoa (Western Samoa), Solomon Islands, Tonga, Tuvalu, and Vanuatu.
-Rest of South America includes
-Argentina, Bolivia, Czech Republic, Chile, Colombia, Ecuador, the Falkland Islands, French Guiana, Guyana, Paraguay, Peru, South Georgia, the South Sandwich Islands, Suriname, Uruguay, and Venezuela.
+Rest of South America includes  -->
+<!-- Argentina, Bolivia, Czech Republic, Chile, Colombia, Ecuador, the Falkland Islands, French Guiana, Guyana, Paraguay, Peru, South Georgia, the South Sandwich Islands, Suriname, Uruguay, and Venezuela.
 Western & Central Europe includes  Austria, Belgium, Bulgaria, Croatia, Denmark, Estonia, Finland, France, Hungary, Iceland, Italy, Ireland, Germany, Greece, Latvia, Liechtenstein, Lithuania, Luxembourg, Malta, Monaco, Montenegro, Netherlands, North Macedonia, Norway, Poland, Portugal, Romania, Serbia, Slovakia, Slovenia, Spain, Sweden, Switzerland, Turkey, and the United Kingdom.
 Other Europe includes Albania, Armenia, Azerbaijan, Belarus, Bosnia and Herzegovina, Georgia,  Kosovo Kazakhstan, Kyrgyzstan, Moldova, Tajikistan, and Ukraine.
+	  -->
 
-Against this backdrop, we examined what it would take to end, on net, the practice of deforestation globally.
-Results suggest that it would require a tax of roughly 70.3 percent to be assessed on the value of converted land for the deforestation rate to fall globally to zero.
-The economic impact of such a policy measure would not be limited to the cost of the tax, it would also impact food production and global trade flows.
+# Modeling the impact of possible measures to mitigate deforestation/promote afforestation
+
+Against the backdrop of the baseline scenario, we turn to the question of the policy required that could prevent global deforestation in the period of 2023--2033.
+Because the area of forested area is mostly determined by the relative returns of land in agriculture and forestry, in our policy scenario we modify the return to land in agriculture by applying a uniform global tax the use of land in the production of agricultural output.
+
+Solving the two models together shows  that it would require a tax of roughly 70 percent to be assessed on the use of land for the deforestation rate to fall globally to zero.
+The economic impact of such a policy measure would not be limited to the cost of the tax, it would also impact food production and its prices.
 For example, with this policy in place, most regions of the world would see significant declines in the amount of agricultural land in 2033 relative to the baseline scenario where land conversion was not taxed.
 For the United States, there would be 2.1 percent less agricultural land (Table 2).
-Larger decreases would happen in other major agricultural production regions such as China (-2.5 percent), South Asia (-8.0 percent), Brazil (-8.5 percent), and Russia (-9.3 percent).
+Larger decreases would happen in other major agricultural production regions such as China (-2.5 percent), South Asia (-8.0 percent), Brazil (-8.5 percent), and Russia (-9.3 percent) (see Table \@ref(tab:baseline-scenario1)).
 
-Table 2. Percentage change in the amount of agricultural land in 2033 with deforestation mitigation policy in place (Source: GTAP qo-Land)
-				∆ (%)
-East Asia			-14.6
-Central America			-14.5
-North Africa & the Middle East	-10.3
-Russia				-9.3
-Southeast Asia			-9.0
-Brazil				-8.5
-Rest of South America		-11.7
-South Asia			-8.0
-Canada				-7.9
-West & Central Europe 		-6.2
-Oceania				-5.9
-Sub-Saharan Africa		-4.1
-China				-2.5
-United States			-2.2
-Japan				17.5
+```{r echo = FALSE}
+
+baselineGTAP = HARr::read_SL4('asset/GTAP/run_baseline.sl4')
+worldGTAP = HARr::read_SL4('asset/GTAP/run_scenario1.sl4')
+
+out = cbind(
+  'Real GDP (\\%)' = ((1 + worldGTAP$qgdp[, 'TOTAL'] / 100) / (1 + baselineGTAP$qgdp[, 'TOTAL'] /
+                                                                 100) - 1) * 100,
+  'Agr. land (\\%)' = ((1 + worldGTAP$qo["land",, 'TOTAL'] / 100) / (1 + baselineGTAP$qo["land",, 'TOTAL'] /
+                                                                 100) - 1) * 100,
+  'Agr. prices (\\%)' = ((1 + worldGTAP$pm["ag",, 'TOTAL'] / 100) / (1 + baselineGTAP$pm["ag",, 'TOTAL'] /
+                                                                 100) - 1) * 100,
+  'Agr. output (\\%)' = ((1 + worldGTAP$qo["ag",, 'TOTAL'] / 100) / (1 + baselineGTAP$qo["ag",, 'TOTAL'] /
+                                                                 100) - 1) * 100
+)
+rownames(out) = regLabels[gGTM[rownames(out)]]
+
+rownames(out)[length(rownames(out))] = 'ROW'
+
+kableExtra::column_spec(
+  kableExtra::column_spec(kableExtra::kable_styling(
+    knitr::kable(out, caption = 'Changes in key variables relative to baseline under a uniform land tax of 70 \\%', digits = 1, escape = FALSE, label = 'baseline-scenario1')
+  ), 1, border_left = TRUE),
+  5,
+  border_right = TRUE
+)
+
+```
+
+```{r echo = FALSE}
+
+world = readRDS('history_world.rds')
+
+out = world[[8]]$gtmResult$forests[1:2,]
+
+rownames(out) = c('Forestland in 2023','Forestland in 2033')
+
+out = rbind(out, `$\\Delta$ (Mha)` = out[2,]-out[1,],`$\\Delta$ (\\%)` = 100*(out[2,]/out[1,]-1))
+colnames(out) = regLabels[colnames(out)]
+
+out = cbind(out, Total = rowSums(out))
+
+out[4,'Total'] = (out[2 , 'Total']/out[1 , 'Total']-1)*100
+
+kableExtra::column_spec(
+  kableExtra::column_spec(
+    kableExtra::kable_styling(
+        knitr::kable(
+          t(out),
+          escape = FALSE,
+          format = 'latex',
+          digits = 1,
+          caption = 'Percentage change in the amount of forested land in 2033 with deforestation mitigation policy in place (Source: Authors\' calculations)'
+        )
+    ),
+    1,
+    border_left = TRUE
+  ),
+  5,
+  border_right = TRUE
+)
+
+```
+
+<!-- Table 2. Percentage change in the amount of agricultural land in 2033 with deforestation mitigation policy in place (Source: GTAP qo-Land) -->
+<!-- 				$\Delta$ (%) -->
+<!-- East Asia			-14.6 -->
+<!-- Central America			-14.5 -->
+<!-- North Africa & the Middle East	-10.3 -->
+<!-- Russia				-9.3 -->
+<!-- Southeast Asia			-9.0 -->
+<!-- Brazil				-8.5 -->
+<!-- Rest of South America		-11.7 -->
+<!-- South Asia			-8.0 -->
+<!-- Canada				-7.9 -->
+<!-- West & Central Europe 		-6.2 -->
+<!-- Oceania				-5.9 -->
+<!-- Sub-Saharan Africa		-4.1 -->
+<!-- China				-2.5 -->
+<!-- United States			-2.2 -->
+<!-- Japan				17.5 -->
+
+```{r echo= FALSE, fig.cap='Changes in ag trade volumes (red arrows denote reductions, green arrows denote increases)'}
+baselineGTAP = HARr::read_SL4('asset/GTAP/run_baseline.sl4')
+worldGTAP = HARr::read_SL4('asset/GTAP/run_scenario1.sl4')
+baselineGTAPC = HARr::read_har('asset/GTAP/run_baseline.slc', TRUE)
+
+changesInAgTrade = ((1+worldGTAP$qxs/100)/(1+baselineGTAP$qxs/100)-1)['ag',,,'TOTAL'] * baselineGTAPC$viws['ag',,]
+
+circlize::chordDiagram(changesInAgTrade, directional = TRUE, col = ifelse(changesInAgTrade>0,'green','red'),direction.type = "arrows", link.arr.type = 'big.arrow')
+
+```
 
 These reductions in the amount of agricultural land would then impact the quantity of agricultural goods that could be produced.
-For example, the 14.6 percent decrease in agrciultural land in East Asia and the 9.0 percent decrease in agrciultural land in Southeast Asia would result in large declines in percentage terms in agricultural production of 5.1 percent and 2.0 percent, respectively (Table 3).
-Surprisingly, Brazil and the United States, despite their dereases in available agricultural land compared to the baseline 2033 scenario, are estimated to see increased agricultural output of 0.2 percent and 0.6 percent, respectively.
+For example, the 14.6 percent decrease in agricultural land in East Asia and the 9.0 percent decrease in agricultural land in Southeast Asia would result in large declines in percentage terms in agricultural production of 5.1 percent and 2.0 percent, respectively (Table 3).
+Surprisingly, Brazil and the United States, despite their decreases in available agricultural land compared to the baseline 2033 scenario, are estimated to see increased agricultural output of 0.2 percent and 0.6 percent, respectively.
 This may be tied to the sharp declines in other regions prompting greater intensity of production to satisfy external demand.
@@
 Table 3. Percentage change in the amount of agricultural goods produced in 2033 with deforestation mitigation policy in place (Source: GTAP qo-Ag)
-				∆ (%)
+				$\Delta$ (%)
 				Agriculture
@@
 When we examine how this impacts exports of agricultural goods, we find that exports from Japan and the United States grow the most, increasing 8.8 percent and 3.5 percent, respectively (Table 4).
-The regions where agricultural exports drop most significantly are East Asia, South Asia, Southeast Asia, and Central America who see decreases of 14.5 percent, 5.9 percent, 5.1 percent,and 5.0 percent, respectively.
+The regions where agricultural exports drop most significantly are East Asia, South Asia, Southeast Asia, and Central America who see decreases of 14.5 percent, 5.9 percent, 5.1 percent, and 5.0 percent, respectively.
 
 Table 4. Percentage change in the amount of agricultural exports in 2033, by region, with deforestation mitigation policy in place (Source: GTAP qxw-Ag)
-				∆ (%)
+				$\Delta$ (%)
 				Agriculture
@@
 Table 5. Change in consumer well-being in 2033, by region, in dollars, with deforestation mitigation policy in place (Source: GTAP EV) 
-				∆ ($)
+				$\Delta$ ($)
 Japan				977,125,000
@@
 Table 6. Change in market price of various goods in 2033, by region, with deforestation mitigation policy in place (Source: GTAP pm) 
-									∆ (%)				
+									$\Delta$ (%)				
 	Agricultural Goods		Forest Products		Lumber		Paper Products		Other Non-Agricultural Goods
@@
 In the United States, as USDA-NASS (Land Values 2022 Summary (August 2022)) valued the average acre of farmland in 2022 at $3,800, the tax would be $2,671.4/acre.
-In Brazil, CEIC valued the average price of Brazilian land at roughly 2,500 Brazilian Reais, which would put the tax at roughly $203 per acre.
+In Brazil, CEIC valued the average price of Brazilian land at roughly 2,500 Brazilian Reals, which would put the tax at roughly $203 per acre.
 Governments would very likely be wary of farmer discontent were they to apply, but they would also have to take into account the larger impacts on their economies.
@@
 By combining these models through the use of R and allowing the models to pass inputs and outputs back and forth iteratively, the benefits of both models have been maintained and their weaknesses greatly minimized.
-This advance holds great promise for advancement on a variety of fronts for researchers and policymakers interested in climate change, agricultural trade, and their interdependencies.
+This advance holds great promise for advancement on a variety of fronts for researchers and policymakers interested in climate change, agricultural trade, and their inter-dependencies.
 
-Unused
+<!-- Unused -->
 
-1a) On aggregate zero percent net deforestation Globally
+<!-- 1a) On aggregate zero percent net deforestation Globally -->
 
-1) Full attainment of AFR100 pledges as part of Bonn Challenge: This says that there will be a 127.81 million hectare increase in forest area just in Sub-Saharan Africa by 2030 [@bonn2023].
+<!-- 1) Full attainment of AFR100 pledges as part of Bonn Challenge: This says that there will be a 127.81 million hectare increase in forest area just in Sub-Saharan Africa by 2030 [@bonn2023].
 To see the list of countries and the hectares pledged, please view Scenario 1 of the Excel document "Bonn Challenge-Africa".
+-->
 
-2) Partial attainment of AFR100 pledges as part of Bonn Challenge: This scenario analyzes full attainment for nations with conflict severity level less than or equal to limited and 0% attainment for regions with a higher conflict severity rating.
+<!-- 2) Partial attainment of AFR100 pledges as part of Bonn Challenge: This scenario analyzes full attainment for nations with conflict severity level less than or equal to limited and 0% attainment for regions with a higher conflict severity rating.
 Using this measure, there would only be a 54.95 million hectare increase in forest area by 2030 [@bonn2023]  [@acled2023].
 To see the list of countries and the hectares pledged, please view Scenario 2 of the Excel document "Bonn Challenge-Africa".
+-->
 
-3) Full attainment of Latin America pledges as part of Bonn Challenge: This says that there will be a 33.64 million hectare increase in forest area by 2030 [@bonn2023].
+<!-- 3) Full attainment of Latin America pledges as part of Bonn Challenge: This says that there will be a 33.64 million hectare increase in forest area by 2030 [@bonn2023].
 To see the list of countries and the hectares pledged, please view the Latin America tab of the Excel document "Bonn Challenge-Other Regions".
+-->
 
-4) Full attainment of tropical nations' pledges as part of Bonn Challenge: This says that there will be a 182.30 million hectare increase in forest area by 2030 [@bonn2023].
+<!-- 4) Full attainment of tropical nations' pledges as part of Bonn Challenge: This says that there will be a 182.30 million hectare increase in forest area by 2030 [@bonn2023].
 To see the list of countries and the hectares pledged, please view the Tropical Forest tab of the Excel document "Bonn Challenge-Other Regions".
 Nations that lie in both tropical and sub-tropical zones and taken to be tropical.
+-->
 
-5) Full attainment of UN strategic plan for forests 2030 (UNFF): This says that all deforestation will halt and there will be a 3% (~120 million hectare) increase in forest area worldwide by 2030 [@sewell2020].
+<!-- 5) Full attainment of UN strategic plan for forests 2030 (UNFF): This says that all deforestation will halt and there will be a 3% (~120 million hectare) increase in forest area worldwide by 2030 [@sewell2020].
 This will be modeled as a uniform 3% increase in forested area across all nations.
+-->
 
-6) Expansion of UN strategic plan for forests: This says that all deforestation will halt and there will be a 15% (~600 million hectare) increase in forest area worldwide by 2030 [@sewell2020].
+<!-- 6) Expansion of UN strategic plan for forests: This says that all deforestation will halt and there will be a 15% (~600 million hectare) increase in forest area worldwide by 2030 [@sewell2020].
 This will be modeled as a uniform 15% increase in forested area across all nations.
+-->
 
-[@defries2010] found a strong correlation between forest loss and exports of agricultural products.
+<!-- [@defries2010] found a strong correlation between forest loss and exports of agricultural products.
 On the other hand, [@turner2005] investigated the role of trade on forests in the context of the Free Trade Area of the Americas (FTAA) using the spatial partial-equilibrium Global Forest Products Model and found that FTAA has a positive impact on forest resources.
+-->
 # References

### 2024-02-27 paper.Rmd
Update paper.Rmd

@@
 
-Afforestation efforts are also thought to not provide equal benefits in every region.
+It is important to understand where to target afforestation efforts, as they are thought to not provide equal benefits in every region.
 For example, afforestation in boreal zones has been shown to provide the smallest net benefits, as the carbon sequestrations benefits of conifer forests are mitigated by the reduction in albedo in winter [@mykleby2017; @li2015].
@@
 South America and Sub-Saharan Africa are thus two regions that show much potential for afforestation.
-It is estimated that these two regions hold at least 50% of the potential global gain while regions such as Northern Africa and the Middle East show little promise as forest growth rates are quite low [@doelman2020].
+It is estimated that these two regions hold at least 50% of the potential for global gains while regions such as Northern Africa and the Middle East show little promise as forest growth rates are quite low [@doelman2020].
 Thus, it is promising that the African Forest Landscape Restoration Initiative (AFR100), which aims to afforest 100 million hectares by 2030, has roughly 70% of its commitments coming from Sub-Saharan nations [@auda2023].
@@
 
-The rationale for joining these two models is that, by passing outputs back and forth between the models until an dual-state equilibrium is achieved, any limitations of either model should be minimized and more robust results should, therefore, be attained.
+The rationale for joining these two models is that, by passing outputs back and forth between the models until a dual-state equilibrium is achieved, any limitations of either model should be minimized and more robust results should, therefore, be attained.
 Establishing a connection between the  GTAP model and the GTM model required a two-step process.

### 2024-02-27 paper.Rmd
Update paper.Rmd

@@
 )
-
 ```
@@
 )
-
 ```
-
-<!-- Table 1. Forestland by region in 2023 and estimated change by 2033 -->
-<!-- 	Russia	Sub-Saharan Africa Brazil Canada Rest of South America	United States Southeast Asia Oceania Western & Central Europe China Central America South Asia Other Europe North Africa & the Middle East Japan East Asia Total  -->
-<!-- Forestland in 2023	838.1	564.7	498.2	412.8	348.5	248.4	243.9	199.2	186.7	164.6	94.6	47.8	40.2	33.8	23.7	14.5	3,959.5 -->
-<!-- Forestland in 2033	828.7	525.0	479.2	407.9	325.0	249.7	221.3	196.1	175.5	164.8	91.2	42.6	36.1	30.9	26.9	14.6	3,815.3 -->
-<!-- $\Delta$ (ha)			-9.5	-39.7	-19.0	-5.0	-23.5	1.3	-22.6	-3.1	-11.1	0.1	-3.4	-5.3	-4.1	-2.9	3.2	0.1	-144.2 -->
-<!-- $\Delta$ (%)			-1.1	-7.0	-3.8	-1.2	-6.7	0.5	-9.3	-1.5	-6.0	0.1	-3.6	-11.0	-10.1	-8.5	13.7	0.7	-3.6 -->
-<!-- Note: Southeast Asia includes Brunei Darussalam, Cambodia, Indonesia, Laos, Malaysia, Myanmar, Philippines, Singapore, Thailand, Timor-Leste, and Vietnam.
-Oceania includes Australia, Cook Islands, Fiji, French Polynesia, Kiribati, Marshall Islands, Micronesia, Nauru, New Caledonia, New Zealand, Palau, Papua New Guinea, Samoa (Western Samoa), Solomon Islands, Tonga, Tuvalu, and Vanuatu.
-Rest of South America includes  -->
-<!-- Argentina, Bolivia, Czech Republic, Chile, Colombia, Ecuador, the Falkland Islands, French Guiana, Guyana, Paraguay, Peru, South Georgia, the South Sandwich Islands, Suriname, Uruguay, and Venezuela.
-Western & Central Europe includes  Austria, Belgium, Bulgaria, Croatia, Denmark, Estonia, Finland, France, Hungary, Iceland, Italy, Ireland, Germany, Greece, Latvia, Liechtenstein, Lithuania, Luxembourg, Malta, Monaco, Montenegro, Netherlands, North Macedonia, Norway, Poland, Portugal, Romania, Serbia, Slovakia, Slovenia, Spain, Sweden, Switzerland, Turkey, and the United Kingdom.
-Other Europe includes Albania, Armenia, Azerbaijan, Belarus, Bosnia and Herzegovina, Georgia,  Kosovo Kazakhstan, Kyrgyzstan, Moldova, Tajikistan, and Ukraine.
-	  -->
 
@@
 
-<!-- Table 2. Percentage change in the amount of agricultural land in 2033 with deforestation mitigation policy in place (Source: GTAP qo-Land) -->
-<!-- 				$\Delta$ (%) -->
-<!-- East Asia			-14.6 -->
-<!-- Central America			-14.5 -->
-<!-- North Africa & the Middle East	-10.3 -->
-<!-- Russia				-9.3 -->
-<!-- Southeast Asia			-9.0 -->
-<!-- Brazil				-8.5 -->
-<!-- Rest of South America		-11.7 -->
-<!-- South Asia			-8.0 -->
-<!-- Canada				-7.9 -->
-<!-- West & Central Europe 		-6.2 -->
-<!-- Oceania				-5.9 -->
-<!-- Sub-Saharan Africa		-4.1 -->
-<!-- China				-2.5 -->
-<!-- United States			-2.2 -->
-<!-- Japan				17.5 -->
-
 ```{r echo= FALSE, fig.cap='Changes in ag trade volumes (red arrows denote reductions, green arrows denote increases)'}
@@
 This may be tied to the sharp declines in other regions prompting greater intensity of production to satisfy external demand.
-
-Table 3. Percentage change in the amount of agricultural goods produced in 2033 with deforestation mitigation policy in place (Source: GTAP qo-Ag)
-				$\Delta$ (%)
-				Agriculture
-East Asia			-5.1
-Southeast Asia			-2.0
-Central America			-2.0
-South Asia			-1.1
-North Africa & the Middle East	-0.9
-Sub-Saharan Africa		-0.3
-China				-0.1
-Russia				0.1
-Brazil				0.2
-Rest of South America		-1.2
-West & Central Europe 		0.4
-United States			0.6
-Oceania				0.7
-Canada				1.3
-Japan				1.4
 
@@
 
-Table 6. Change in market price of various goods in 2033, by region, with deforestation mitigation policy in place (Source: GTAP pm) 
-									$\Delta$ (%)				
-	Agricultural Goods		Forest Products		Lumber		Paper Products		Other Non-Agricultural Goods
-Oceania						1.2	0.4	0.5	0.5	0.5
-China						1.1	0.4	0.4	0.4	0.4
-Japan						-0.3	0.4	0.4	0.4	0.4
-East Asia					5.1	0.3	0.3	0.3	0.4
-Southeast Asia					2.8	0.3	0.3	0.4	0.3
-South Asia					3.1	0.3	0.3	0.3	0.3
-Canada						0.7	0.5	0.5	0.5	0.5
-United States					0.7	0.5	0.5	0.5	0.5
-Central America					2.5	0.3	0.3	0.4	0.3
-Brazil						1.2	0.5	0.5	0.5	0.5
-Rest of South America				2.1	0.3	0.3	0.3	0.3
-West & Central Europe 				0.9	0.4	0.4	0.4	0.4
-Russia						1.1	0.4	0.4	0.4	0.4
-Sub-Saharan Africa				1.5	0.4	0.4	0.4	0.4
-North Africa & the Middle East			2.0	0.4	0.4	0.4	0.4
-
-Land price seems weird. I have it decreasing????
-
 # Discussion

### 2024-02-27 paper.Rmd
Update paper.Rmd

@@
 
+```{r echo = FALSE}
+
+baselineGTAP = HARr::read_SL4('asset/GTAP/run_baseline.sl4')
+worldGTAP = HARr::read_SL4('asset/GTAP/run_scenario1.sl4')
+
+out = rbind(out, `$\\Delta$ ($)` = worldGTAP$EV-baselineGTAP$EV)
+)
+rownames(out) = regLabels[gGTM[rownames(out)]]
+
+rownames(out)[length(rownames(out))] = 'ROW'
+
+kableExtra::column_spec(
+  kableExtra::column_spec(kableExtra::kable_styling(
+    knitr::kable(out, caption = 'Change in consumer well-being in 2033, by region, in dollars, with deforestation mitigation policy in place', digits = 1, escape = FALSE, label = 'baseline-scenario1')
+  ), 1, border_left = TRUE),
+  5,
+  border_right = TRUE
+)
+
+```
 Table 5. Change in consumer well-being in 2033, by region, in dollars, with deforestation mitigation policy in place (Source: GTAP EV) 
@@
 
+# References
+
 <!-- Unused -->
@@
 -->
-# References

### 2024-02-28 paper.Rmd
Update paper.Rmd

@@
 
-# Deforestation under the baseline projections
+# Changes under the baseline projections
 
@@
 In Table \@ref(tab:baseline-growth) we show the projected population and GDP growths per region.
+GDP increases for all regions of the world with the largest percentage changes occurring in South Asia (+308.6%), China (+189.8%), and Sub-Saharan Africa (+172.9%).
+Population also is projected to grow in most regions with the largest changes happening in Sub-Saharan Africa (+55.7%).
+Japan, Russia, and East Asia are projected to see their populations decline by 10.4%, 4.7%, and 2.6%, respectively.
+As for the projected change in agricultural land, baseline estimates show significant growth in Oceania (+27.6%), Rest of Latin America (+25.8%), Sub-Saharan Africa (+17.4%), Central America (+13.9%), and Brazil (+13.9%).
+Japan is the only nation that sees a significant decline (-34.9%) in agricultural land, which is likely tied in part to the population decline described above.
 

### 2024-02-28 paper.Rmd
Update paper.Rmd

@@
 
-Against the backdrop of the baseline scenario, we turn to the question of the policy required that could prevent global deforestation in the period of 2023--2033.
-Because the area of forested area is mostly determined by the relative returns of land in agriculture and forestry, in our policy scenario we modify the return to land in agriculture by applying a uniform global tax the use of land in the production of agricultural output.
-
+Against the backdrop of the baseline scenario, we turn to the question of what policy could prevent global deforestation in the period of 2023--2033.
+Because the area of forestland is mostly determined by the relative returns of land in agriculture and forestry, in our policy scenario we modify the return to land in agriculture by applying a uniform global tax on the use of land in the production of agricultural output.
 Solving the two models together shows that it would require a tax of roughly 70 percent to be assessed on the use of land for the deforestation rate to fall globally to zero.
@@
 For example, with this policy in place, most regions of the world would see significant declines in the amount of agricultural land in 2033 relative to the baseline scenario where land conversion was not taxed.
-For the United States, there would be 2.1 percent less agricultural land (Table 2).
-Larger decreases would happen in other major agricultural production regions such as China (-2.5 percent), South Asia (-8.0 percent), Brazil (-8.5 percent), and Russia (-9.3 percent) (see Table \@ref(tab:baseline-scenario1)).
+For the United States, there would be 2.1 percent less agricultural land (Table \@ref(tab:baseline-scenario1)).
+Larger decreases would happen in other major agricultural production regions such as China (-2.5 percent), South Asia (-7.9 percent), Brazil (-8.6 percent), and Russia (-9.3 percent).
+
+These reductions in the amount of agricultural land would then impact the quantity of agricultural goods that could be produced.
+For example, the 15.0 percent decrease in agricultural land in East Asia and the 9.0 percent decrease in agricultural land in Southeast Asia would result in large declines in percentage terms in agricultural production of 5.0 percent and 1.7 percent, respectively.
+Surprisingly, Brazil and the United States, despite their decreases in available agricultural land compared to the baseline 2033 scenario, are estimated to see increased agricultural output of 0.6 percent and 1.0 percent, respectively.
+This may be tied to the sharp declines in other regions prompting greater intensity of production to satisfy external demand.
 
@@
 ```
-
-These reductions in the amount of agricultural land would then impact the quantity of agricultural goods that could be produced.
-For example, the 14.6 percent decrease in agricultural land in East Asia and the 9.0 percent decrease in agricultural land in Southeast Asia would result in large declines in percentage terms in agricultural production of 5.1 percent and 2.0 percent, respectively (Table 3).
-Surprisingly, Brazil and the United States, despite their decreases in available agricultural land compared to the baseline 2033 scenario, are estimated to see increased agricultural output of 0.2 percent and 0.6 percent, respectively.
-This may be tied to the sharp declines in other regions prompting greater intensity of production to satisfy external demand.
 

### 2024-02-28 paper.Rmd
Update paper.Rmd

@@
 
+When we examine how this impacts exports of agricultural goods, we find that exports from Japan and the United States grow the most, increasing 8.8 percent and 3.5 percent, respectively (Table 4).
+The regions where agricultural exports drop most significantly are East Asia, South Asia, Southeast Asia, and Central America who see decreases of 14.5 percent, 5.9 percent, 5.1 percent, and 5.0 percent, respectively.
+
 ```{r echo= FALSE, fig.cap='Changes in ag trade volumes (red arrows denote reductions, green arrows denote increases)'}
@@
 
-When we examine how this impacts exports of agricultural goods, we find that exports from Japan and the United States grow the most, increasing 8.8 percent and 3.5 percent, respectively (Table 4).
-The regions where agricultural exports drop most significantly are East Asia, South Asia, Southeast Asia, and Central America who see decreases of 14.5 percent, 5.9 percent, 5.1 percent, and 5.0 percent, respectively.
-
-Table 4. Percentage change in the amount of agricultural exports in 2033, by region, with deforestation mitigation policy in place (Source: GTAP qxw-Ag)
-				$\Delta$ (%)
-				Agriculture
-Japan				8.78
-United States			3.49
-Canada				3.13
-China				1.96
-Oceania				1.67
-Russia				1.52
-West & Central Europe 		1.16
-Brazil				1.03
-Rest of South America		-2.67
-Sub-Saharan Africa		0.02
-North Africa & the Middle East	-1.52
-Central America			-5.02
-Southeast Asia			-5.14
-South Asia			-5.94
-East Asia			-14.45
-
-Overall, this tax policy, which results in zero net deforestation globally, has a rather significant negative economic impact on all regions except Japan and Canada who see an increase in consumer well-being of $977.1 million and $166.9 million, respectively (Table 5).
+Overall, this tax policy, which results in zero net deforestation globally, has a rather significant negative economic impact on all regions except Japan and Canada who see an increase in consumer well-being of $977.1 million and $166.9 million, respectively (Table \@ref(tab:EV)).
 Consumer well-being is a measurement of equivalent variation and here measures the benefit/harm consumers experience in terms of income as a result of the policy changes.
@@
   kableExtra::column_spec(kableExtra::kable_styling(
-    knitr::kable(out, caption = 'Change in consumer well-being in 2033, by region, in dollars, with deforestation mitigation policy in place', digits = 1, escape = FALSE, label = 'baseline-scenario1')
+    knitr::kable(out, caption = 'Change in consumer well-being in 2033, by region, in dollars, with deforestation mitigation policy in place', digits = 1, escape = FALSE, label = 'EV')
   ), 1, border_left = TRUE),
@@
 ```
-Table 5. Change in consumer well-being in 2033, by region, in dollars, with deforestation mitigation policy in place (Source: GTAP EV) 
-				$\Delta$ ($)
-Japan				977,125,000
-Canada				166,937,500
-Oceania				-169,625,000
-Brazil				-735,125,000
-Rest of South America		-3,848,625,000
-United States			-1,085,000,000
-Russia				-1,152,375,000
-Central America			-4,471,500,000
-North Africa & the Middle East	-5,012,125,000
-East Asia			-5,065,625,000
-West & Central Europe		-7,152,500,000
-Sub-Saharan Africa		-7,850,500,000
-Southeast Asia			-13,601,000,000
-China				-14,452,000,000
-South Asia			-28,024,000,000
 

### 2024-02-28 paper.Rmd
Update paper.Rmd

@@
 Much of this loss in well-being may be the result of higher prices.
-Japan is the only nation that sees any price impacts that are beneficial for consumers resulting from this policy, with the price of agricultural goods decreasing 0.3 percent (Table 6).
+Japan is the only nation that sees any price impacts that are beneficial for consumers resulting from this policy, with the price of agricultural goods decreasing 0.3 percent (Table \@ref(tab:pm)).
 All other regions, for both agricultural and non-agricultural products, see prices rise between 0.3 and 5.1 percent.
+
+```{r echo= FALSE}
+baselineGTAP = HARr::read_SL4('asset/GTAP/run_baseline.sl4')
+worldGTAP = HARr::read_SL4('asset/GTAP/run_scenario1.sl4')
+
+out = ((1+worldGTAP$pm/100)/(1+baselineGTAP$pm/100)-1)
+
+colnames(out) = regLabels[colnames(out)]
+out = cbind(out, Total = rowSums(out))
+
+out[4,'Total'] = (out[2 , 'Total']/out[1 , 'Total']-1)*100
+
+kableExtra::column_spec(
+  kableExtra::column_spec(kableExtra::kable_styling(
+    knitr::kable(out, caption = 'Percent change in market prices for agricultural goods in 2033, in dollars, with deforestation mitigation policy in place', digits = 1, escape = FALSE, label = 'pm')
+  ), 1, border_left = TRUE),
+  5,
+  border_right = TRUE
+)
+
+```
 

### 2024-02-28 paper.Rmd
Update paper.Rmd

@@
     - \usepackage{setspace}\doublespacing
-abstract: "The projected growth of the population and incomes as implied by various shared socioeconomic pathways is expected to create pressures on converting forests into farmland as food prices increase.
-Depending on the specific path taken, SSP1 through SSP5, and the continuation of the current trends in agricultural productivity growth, the market impacts are expected to vary but generally result in increases in food prices.
-Policies aimed at reversing the deforestation trend by afforesting cropland would result in severe negative market impacts such as  higher food prices, especially in low agricultural productivity/high population growth scenarios."
+abstract: "The projected growth in population and incomes as implied by various shared socioeconomic pathways is expected to create pressure to convert forestland into farmland.
+At the same time, the increasingly negative climate impacts are expected to generate increasing pressure to enhance the terrestrial carbon sink.
+These goals seem incompatible as reversing the deforestation trend by afforesting cropland might result in negative market impacts such as higher food prices, especially in low agricultural productivity/high population growth scenarios.
+In order to better grasp the interrelationship and the magnitude of harm that might be caused by prioritizing the goal of mitigating deforestation, this analysis connects two respected global economic models: the Global Trade Analysis Project (GTAP) model and the Global Timber Model (GTM).
+"
 ---
@@
 
+# Acknowledgements
+
+The findings and conclusions in this publication are those of the author(s) and should not be construed to represent any official USDA or U.S. Government determination or policy.
+
+# Funding statement
+
+This research was supported [in part] by the U.S. Department of Agriculture, Economic Research Service.
+
 # References

### 2024-02-28 paper.Rmd
Update paper.Rmd

@@
 In order to better grasp the interrelationship and the magnitude of harm that might be caused by prioritizing the goal of mitigating deforestation, this analysis connects two respected global economic models: the Global Trade Analysis Project (GTAP) model and the Global Timber Model (GTM).
-"
+Results from the model suggest that it is possible through fiscal policy to preserve 144.2 million hectares of forestland that otherwise would be converted to agricultural land by 2033.
+There does appear to be a rather significant economic price for doing so, as the avoided deforestation would result in less agricultural output and higher market prices.
+This is estimated to impact the well-being of global consumers by $92.1 billion, which translates to an global average cost of $10.6 per person in 2033."
 ---

### 2024-02-28 paper.Rmd
Update paper.Rmd

@@
 ---
-title: "Implications of afforestation policies for global markets"
+title: "Avoiding deforestation: The implications for global markets"
 date: "`r format(Sys.time(), '%B %d, %Y')`"
@@
 Results from the model suggest that it is possible through fiscal policy to preserve 144.2 million hectares of forestland that otherwise would be converted to agricultural land by 2033.
-There does appear to be a rather significant economic price for doing so, as the avoided deforestation would result in less agricultural output and higher market prices.
+As to the economic price for doing so, the avoided deforestation would in most regions of the world result in less agricultural output and higher market prices.
 This is estimated to impact the well-being of global consumers by $92.1 billion, which translates to an global average cost of $10.6 per person in 2033."
@@
 
-This analysis shows that, at an estimated tax rate of 70.3 percent of the value of each acre of land that is converted to agricultural land, global net deforestation would drop to roughly zero.
-In the United States, as USDA-NASS (Land Values 2022 Summary (August 2022)) valued the average acre of farmland in 2022 at $3,800, the tax would be $2,671.4/acre.
-In Brazil, CEIC valued the average price of Brazilian land at roughly 2,500 Brazilian Reals, which would put the tax at roughly $203 per acre.
-Governments would very likely be wary of farmer discontent were they to apply, but they would also have to take into account the larger impacts on their economies.
-Restricting agricultural land, as shown above, would for the most part restrict output and increase prices.
-It is uncertain whether the gains from the increase in prices would be sufficient to offset the loss of potential revenue that would have come from expanded agricultural operations.
-As governments might be reluctant to impose such taxes on their own citizens, it is possible that policymakers who wished to see deforestation stopped could impose import tariffs equal to the tax rate to achieve the same result.
+This analysis shows that, at an estimated tax rate of 70.3 percent of the value of each hectare of land that is converted to agricultural land, global net deforestation would drop to roughly zero by 2033, preserving 144.2 million hectares of forestland that otherwise would have been converted to agricultural land.
+In the United States, as USDA-NASS (Land Values 2022 Summary (August 2022)) valued the average hectare of farmland in 2022 at $1,537.8, the tax would equate to $1,081.1/hectare.
+In Brazil, CEIC valued the average price of Brazilian land at roughly 2,500 Brazilian Reals, which would put the tax at roughly $82.2 per hectare.
+The economic price of this would likely be transfered to consumers through less agricultural output and higher market prices.
+Overall, such an action is estimated to reduce consumer well-being by $92.1 billion or $10.6 per every person in 2033.
+Governments might be wary of farmer discontent were they to take such fiscal actions and thus might be reluctant to impose such taxes on them.
+Thus, policymakers who wish to see deforestation stopped might have to resort to import tariffs on products produced on land converted from forestland for agricultural use.
 

### 2024-02-28 paper.Rmd
Update paper.Rmd

@@
 abstract: "The projected growth in population and incomes as implied by various shared socioeconomic pathways is expected to create pressure to convert forestland into farmland.
-At the same time, the increasingly negative climate impacts are expected to generate increasing pressure to enhance the terrestrial carbon sink.
-These goals seem incompatible as reversing the deforestation trend by afforesting cropland might result in negative market impacts such as higher food prices, especially in low agricultural productivity/high population growth scenarios.
+At the same time, the increasingly negative climate impacts are expected to generate further pressure to enhance the terrestrial carbon sink.
+These goals are potentially incompatible as reversing the deforestation trend by afforesting cropland might result in negative market impacts such as higher food prices, especially in low agricultural productivity/high population growth scenarios.
 In order to better grasp the interrelationship and the magnitude of harm that might be caused by prioritizing the goal of mitigating deforestation, this analysis connects two respected global economic models: the Global Trade Analysis Project (GTAP) model and the Global Timber Model (GTM).
@@
 As to the economic price for doing so, the avoided deforestation would in most regions of the world result in less agricultural output and higher market prices.
-This is estimated to impact the well-being of global consumers by $92.1 billion, which translates to an global average cost of $10.6 per person in 2033."
+This is estimated to impact the well-being of global consumers by $92.1 billion, which translates to a global average cost of $10.6 per person in 2033."
 ---
@@
 Currently, the United Nations' medium variant population projections suggest that between 2023 and 2033 the worlds' population will grow by 8.8 percent, a gain of 707.1 million people [@un2024].
-And where population growth is especially high, which is the case at present throughout most of Africa (2023--2033 population growth: 25.0%: 360.6 million), this puts extreme pressure on policymakers to allow greater conversion of forestland into farmland.
+And where population growth is especially high, which is the case at present throughout most of Africa (2023 to 2033 population growth: 25.0%: 360.6 million), this puts extreme pressure on policymakers to allow greater conversion of forestland into farmland.
 Still, some research has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017].
 
-It is important to understand where to target afforestation efforts, as they are thought to not provide equal benefits in every region.
+If such efforts are to be made, it is important to understand where to target them, as afforestation is thought to not provide equal benefits in every region.
 For example, afforestation in boreal zones has been shown to provide the smallest net benefits, as the carbon sequestrations benefits of conifer forests are mitigated by the reduction in albedo in winter [@mykleby2017; @li2015].
@@
 
-Toward this end, @steinbuks2012 developed their global partial equilibrium FABLE model (Forestry, Agriculture, Biofuels, Land Use, and Environment) to identify the optimal land use choices given the competing demands of meeting greenhouse gas targets and satisfying demand for food, bioenergy, forest products, and ecosystem services.
+Several notable modeling attempts have been made in this direction.
+@steinbuks2012 developed their global partial equilibrium FABLE model (Forestry, Agriculture, Biofuels, Land Use, and Environment) to identify the optimal land use choices given the competing demands of meeting greenhouse gas targets and satisfying demand for food, bioenergy, forest products, and ecosystem services.
 While FABLE is a powerful tool for examining the optimal trajectory of various land uses under specific demand assumptions, it is focused on specific sectors of the economy and neglects general equilibrium and household welfare effects [@steinbuks2012; @hertel2017].
@@
 The KLUM@GTAP framework integrated the Kleines Land Use Model (KLUM) with an extended yet static version of GTAP called GTAP-EFL to evaluate the impact of climate change on cropland allocation.
-GTAP-EFL separates energy factors from intermediate inputs and incorporates them into capital, while also considering CO2 emissions.
-KLUM, a global agricultural land-use model, connects the economy to global crop allocation to maximize producer returns under specific risk assumptions.
+KLUM is a global agricultural land-use model that connects the economy to global crop allocation to maximize producer returns under specific risk assumptions, and GTAP-EFL separates energy factors from intermediate inputs and incorporates them into capital, while also considering CO2 emissions.
 KLUM@GTAP substitutes the land allocation mechanism within GTAP-EFL by utilizing regionally aggregated area changes in cropland determined by KLUM to update cropland shares in GTAP-EFL [@ronneberger2009].
 
-More recent modeling attempts of land-use change have begun to move away from comparative static analysis.
+More recent modeling attempts of land-use change have begun to move away from comparative static analysis though.
 For example, the dynamic GTAP model, supported by the GTAP-AEZ database, considers land market effects, which were identified as significant in driving results by @stevenson2013.
@@
 Data for this were drawn from CEPII [@CEPII2024].
-To build our scenario, we used two key projected variables by CEPII: population and GDP.
+To build our scenario, we use two of their variables that we determined to be key: population and GDP.
 The CEPII projections do not contain any information on land use.
@@
 Using the GTAP model in connection with the GTM model (to assure that the changes in land are consistent between the two models), we calculate the overall change in factor quantity/productivity growth required to achieve the targeted values in 2033.
-In Table \@ref(tab:baseline-growth) we show the projected population and GDP growths per region.
+In Table \@ref(tab:baseline-growth), we show the projected population and GDP growth per region.
 GDP increases for all regions of the world with the largest percentage changes occurring in South Asia (+308.6%), China (+189.8%), and Sub-Saharan Africa (+172.9%).
@@
 Japan, Russia, and East Asia are projected to see their populations decline by 10.4%, 4.7%, and 2.6%, respectively.
-As for the projected change in agricultural land, baseline estimates show significant growth in Oceania (+27.6%), Rest of Latin America (+25.8%), Sub-Saharan Africa (+17.4%), Central America (+13.9%), and Brazil (+13.9%).
-Japan is the only nation that sees a significant decline (-34.9%) in agricultural land, which is likely tied in part to the population decline described above.
+As for the projected change in agricultural land, our baseline prjoections show significant growth in Oceania (+27.6%), Rest of Latin America (+25.8%), Sub-Saharan Africa (+17.4%), Central America (+13.9%), and Brazil (+13.9%).
+Japan is the only nation that sees a significant decline (-34.9%) in agricultural land, which is likely tied in part to the population decline just described.
 

### 2024-02-28 paper.Rmd
my changes?

@@
     - \usepackage{setspace}\doublespacing
-abstract: "The projected growth in population and incomes as implied by various shared socioeconomic pathways is expected to create pressure to convert forestland into farmland.
+abstract: "The projected growth in population and incomes is expected to create pressure to convert forestland into farmland.
 At the same time, the increasingly negative climate impacts are expected to generate further pressure to enhance the terrestrial carbon sink.
-These goals are potentially incompatible as reversing the deforestation trend by afforesting cropland might result in negative market impacts such as higher food prices, especially in low agricultural productivity/high population growth scenarios.
-In order to better grasp the interrelationship and the magnitude of harm that might be caused by prioritizing the goal of mitigating deforestation, this analysis connects two respected global economic models: the Global Trade Analysis Project (GTAP) model and the Global Timber Model (GTM).
-Results from the model suggest that it is possible through fiscal policy to preserve 144.2 million hectares of forestland that otherwise would be converted to agricultural land by 2033.
+Even though these goals are  incompatible as reversing the deforestation trend by afforesting cropland would result in negative market impacts such as higher food prices, using the GTAP and GTM models, we find that these impacts would be relatively small if the goal of  preserving 144.2 million hectares of forestland that otherwise would be converted to agricultural land by 2033 is achieved through a tax on using land in agricultural production.
 As to the economic price for doing so, the avoided deforestation would in most regions of the world result in less agricultural output and higher market prices.
@@
 regLabels = c(
-  "US" = 'US',
+  "US" = 'United States',
   "CHINA" = 'China',
@@
   "RUSSIA" = 'Russia',
-  "EU ANNEX I" = 'EU I',
+  "EU ANNEX I" = 'West/Central Europe',
   "EU NON ANNEX I" = 'EU II',
@@
   "CENT AMER." = 'Central America',
-  "RSAM" = 'Rest of Latin America',
+  "RSAM" = 'Other Latin America',
   "SSAF" = 'Sub-Saharan Africa',
-  "SE ASIA" = 'Southeast Asia',
+  "SE ASIA" = 'SE Asia',
   "OCEANIA" = 'Oceania',
@@
 names(gGTM)= gGTAP
+
+gGTM['ROW']='ROW'
+regLabels['ROW']='ROW'
 
@@
           escape = FALSE,
-          format = 'latex',
+          #format = 'latex',
           digits = 1,
@@
 
-circlize::chordDiagram(changesInAgTrade, directional = TRUE, col = ifelse(changesInAgTrade>0,'green','red'),direction.type = "arrows", link.arr.type = 'big.arrow')
+colnames(changesInAgTrade) = rownames(changesInAgTrade) = regLabels[gGTM[c("ROW", "china", "japan", "ROW", "seasia", "ROW", "ROW", "usa", "ROW", "brazil", "samerica", "eui", "ROW", 
+"ROW", "ssa", "ROW", "ROW")]]
+
+changesInAgTrade = matsbyname::aggregate_byname(changesInAgTrade)
+
+diag(changesInAgTrade)=0
+
+#round(changesInAgTrade,0)
+
+circlize::chordDiagram(changesInAgTrade, directional = TRUE, col = ifelse(changesInAgTrade>0,'green','red'),direction.type = c("arrows"), link.arr.type = 'big.arrow')
 
@@
 
-out = rbind(out, `$\\Delta$ ($)` = worldGTAP$EV-baselineGTAP$EV)
+out = cbind(
+  `Baseline (B\\$)` = baselineGTAP$ev[, 'TOTAL'] / 1000,
+  `Scenario (B\\$)` = worldGTAP$ev[, 'TOTAL'] / 1000,
+  `$\\Delta$ (B\\$)` = (worldGTAP$ev[, 'TOTAL'] - baselineGTAP$ev[, 'TOTAL']) /
+    1000
 )
+
 rownames(out) = regLabels[gGTM[rownames(out)]]
 
-rownames(out)[length(rownames(out))] = 'ROW'
+#rownames(out)[length(rownames(out))] = 'ROW'
+
+out = rbind(out, 'Total'=colSums(out))
 
@@
   kableExtra::column_spec(kableExtra::kable_styling(
-    knitr::kable(out, caption = 'Change in consumer well-being in 2033, by region, in dollars, with deforestation mitigation policy in place', digits = 1, escape = FALSE, label = 'EV')
+    knitr::kable(out, caption = 'Change in consumer well-being in 2033, by region, in billions dollars, with deforestation mitigation policy in place', digits = 1, escape = FALSE, label = 'EV',format.args = list(big.mark = ","))
   ), 1, border_left = TRUE),
-  5,
+  4,
   border_right = TRUE
@@
 ```{r echo= FALSE}
+
 baselineGTAP = HARr::read_SL4('asset/GTAP/run_baseline.sl4')
@@
 
-out = ((1+worldGTAP$pm/100)/(1+baselineGTAP$pm/100)-1)
+out = data.frame('Ag prices'=(1+worldGTAP$pm['ag',,'TOTAL']/100)/(1+baselineGTAP$pm['ag',,'TOTAL']/100)-1, check.names = FALSE) *100
 
-colnames(out) = regLabels[colnames(out)]
-out = cbind(out, Total = rowSums(out))
-
-out[4,'Total'] = (out[2 , 'Total']/out[1 , 'Total']-1)*100
+rownames(out) = regLabels[gGTM[rownames(out)]]
+#out = cbind(out, Total = rowSums(out))
+#out[4,'Total'] = (out[2 , 'Total']/out[1 , 'Total']-1)*100
 
@@
   kableExtra::column_spec(kableExtra::kable_styling(
-    knitr::kable(out, caption = 'Percent change in market prices for agricultural goods in 2033, in dollars, with deforestation mitigation policy in place', digits = 1, escape = FALSE, label = 'pm')
+    knitr::kable(out, caption = 'Percent change in market prices for agricultural goods in 2033, with deforestation mitigation policy in place, relative to baseline', digits = 1, escape = FALSE, label = 'pm')
   ), 1, border_left = TRUE),
-  5,
+  2,
   border_right = TRUE
@@
 This advance holds great promise for advancement on a variety of fronts for researchers and policymakers interested in climate change, agricultural trade, and their inter-dependencies.
+
+# Conclusions
 

### 2024-02-28 paper.Rmd
Added conclusions

@@
 ---
-title: "Avoiding deforestation: The implications for global markets"
+title: "Avoiding global deforestation by taxing land in agricultural production: The implications for global markets"
 date: "`r format(Sys.time(), '%B %d, %Y')`"
@@
 ```
+
+\vspace{1in}
+
+**The findings and conclusions in this publication are those of the author(s) and should not be construed to represent any official USDA or U.S. Government determination or policy.**
+
+**This research was supported in part by the U.S. Department of Agriculture, Economic Research Service.**
+
+\newpage
 
@@
 
-The rationale for joining these two models is that, by passing outputs back and forth between the models until a dual-state equilibrium is achieved, any limitations of either model should be minimized and more robust results should, therefore, be attained.
-Establishing a connection between the GTAP model and the GTM model required a two-step process.
-First, we had to identify the set of variables that are common to both models and their closures.
-From those, we selected which variables would be determined endogenously by one model and held as exogenous in the other model.
-In our case, we decided to model the return on land using the GTAP model and feed those results into the GTM model as exogenous.
-The GTM model then endogenously determined the amount of land available for crop and livestock production, and we fed that back into the GTAP model as an exogenous input.
-In the second step, we established the connection between the models on the software level.
-This step required that the GTAP model, which is written in GEMPACK, be able to handle inputs from the GTM model, which is written in GAMS, and submit the outputs back to the GTM model.
-As GEMPACK and GAMS were developed long before software integration became important or even possible, we settled on the use of R [@R2021] to create the user functions that executed both models, collected their outputs, and automatically created parameter files.
+Both the GTAP and GTM models represent widely used models for policy analysis.
+In the case of GTAP, the model is used to describe the implications of various policy instruments for national and globla markets, including prices, output and trade.
+The GTM model is, on the other hand, capable to translating the impacts of population and global growth, and the changes in land returns, to changes in global forested areas and calculate the impacts on total carbon contained in forests.
 
-Once R was able to control the interactions between both models, we were then able to formulate the combined solution as a fairly simple math program that is solved in R:
+Even though the GTAP and GTM models capture important aspects of the global economy and the biophysical world, they do not contain certain behaviors.
+In the case of GTAP, the model assumes that the amount of agricultural land is fixed, while the GTM considers the returns to land fixed.
+By combining the two models, we can fill each model missing behavior for a more complete description of the global impacts.
 
-\begin{equation}
-\mbox{GTAP}:V_1\to V_2
-\end{equation}
+The combination of the two models means that we use the GTAP model to create the link between the amounts of available land and its returns for the benefit of the GTM model, while at the same time we use the GTM model to create the response of  agricultural land quantity to returns in land returns for the benefit of the GTAP model.
+By solving both models together, we can greatly expand the descriptive power of both models.
 
-\begin{equation}
-\mbox{GTM}:V_2\to V_1
-\end{equation}
+We achieve the combined solution of GTAP and GTM by passing outputs back and forth between the models until a dual-state equilibrium is achieved.
+Establishing a connection between the GTAP model and the GTM model requires that we  model the return on land using the GTAP model and feed those results into the GTM model as exogenous inputs.
+The GTM model then endogenously determines the amount of land available for crop and livestock production, and we feed that back into the GTAP model as an exogenous input.
 
-\begin{equation}
-s.t. V_2 = V_1
-\end{equation}
+On the software level, we establish the connection between the models using an R script, which is capable of handling other software, including generating required inputs, executing it and reading in any outputs.
+This is required to operate the GTAP model, which is written in GEMPACK, and the GTM model, which is written in GAMS.
+As GEMPACK and GAMS were developed long before software integration became important or even possible, we use much more modern technologies made available through the R software  [@R2021] to create the user functions that execut both models, collect their outputs, and automatically create any required input files.
 
-where $V_1$ is a vector of all the variables in both systems, i.e., all variables in GTAP and all variables in GTM.
+<!-- Once R was able to control the interactions between both models, we were then able to formulate the combined solution as a fairly simple math program that is solved in R: -->
+
+<!-- \begin{equation} -->
+<!-- \mbox{GTAP}:V_1\to V_2 -->
+<!-- \end{equation} -->
+
+<!-- \begin{equation} -->
+<!-- \mbox{GTM}:V_2\to V_1 -->
+<!-- \end{equation} -->
+
+<!-- \begin{equation} -->
+<!-- s.t. V_2 = V_1 -->
+<!-- \end{equation} -->
+
+<!-- where $V_1$ is a vector of all the variables in both systems, i.e., all variables in GTAP and all variables in GTM.
 Of import, $V_1$ includes the land change variables, which are exogenous to the GTAP model.
@@
 With several iterations, the system converges to the change in land and land rent that is consistent across the models, i.e., the land change results in the land rent changes that, when entered into the GTM model, result in the same land changes as assumed by the GTAP model.
+-->
 
@@
 
-# Acknowledgements
+According to CEPII's projections, the growth of the global economy in the next ten years is expected to result in tremendous increases in global and regional GDP accompanied with a moderate growth in population.
+Another set of projections by GTM, however, suggests that this growth would also mean significant reductions in the world's forests that would be converted to agricultural land, resulting in a loss of over 100 million hectares of forest before 2033.
 
-The findings and conclusions in this publication are those of the author(s) and should not be construed to represent any official USDA or U.S. Government determination or policy.
+By combining two well known global models---the GTAP model that describes the world markets and trade, and the GTAM model that describes the impact of economic and population growth along with land rents on the amounts of forested areas---we were able to estimate the size of the tax on agricultural land needed to prevent deforestation in the coming ten years through reducing the returns to land when used in agriculture.
+We find that even though the tax is substantial at about 70 percent, it results in only very small reductions in the projected growth in global GDP and welfare.
+We also observe that the reduction in available agricultural land relative to the baseline growth results in reduced agricultural output and higher prices, which may require additional policy action to avoid negative impacts in terms of food security.
 
-# Funding statement
+Finally, we note that a uniform global tax reducing the returns to land to the point of preventing global deforestation would produce highly differentiated impacts around the world's regions.
+Some regions, such as China, Russia, Canada and the United States would increase their forested areas as a result of the tax, while the regions with the greatest deforestation pressures, such as Sub-Saharan Africa, Brazil and Southeast Asia would only reduce the rate of their deforestation, for a zero change in the global forest area.
 
-This research was supported [in part] by the U.S. Department of Agriculture, Economic Research Service.
+\newpage
 

### 2024-02-28 paper.tex
Added conclusions

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{setspace}\doublespacing
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Avoiding global deforestation by taxing land in agricultural production: The implications for global markets},
  pdfauthor={Eric C. Davis, Maros Ivanic, and Brent Sohngen},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Avoiding global deforestation by taxing land in agricultural production: The implications for global markets}
\author{Eric C. Davis, Maros Ivanic, and Brent Sohngen}
\date{February 28, 2024}

\begin{document}
\maketitle
\begin{abstract}
The projected growth in population and incomes is expected to create pressure to convert forestland into farmland. At the same time, the increasingly negative climate impacts are expected to generate further pressure to enhance the terrestrial carbon sink. Even though these goals are incompatible as reversing the deforestation trend by afforesting cropland would result in negative market impacts such as higher food prices, using the GTAP and GTM models, we find that these impacts would be relatively small if the goal of preserving 144.2 million hectares of forestland that otherwise would be converted to agricultural land by 2033 is achieved through a tax on using land in agricultural production. As to the economic price for doing so, the avoided deforestation would in most regions of the world result in less agricultural output and higher market prices. This is estimated to impact the well-being of global consumers by \$92.1 billion, which translates to a global average cost of \$10.6 per person in 2033.
\end{abstract}

\vspace{1in}

\textbf{The findings and conclusions in this publication are those of the author(s) and should not be construed to represent any official USDA or U.S. Government determination or policy.}

\textbf{This research was supported in part by the U.S. Department of Agriculture, Economic Research Service.}

\newpage

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

In 2020, there were 4.1 billion hectares of forestland globally covering roughly 31.4\% of the land (FAO and UNEP 2020). These forests play an important role in maintaining and growing the land-based carbon sink by sequestering about 30\% of GHG emissions currently (Roe et al. 2019). This is because of forests proven ability to sequester carbon not only in the stems of trees (Davis, Sohngen, and Lewis 2022) but in their soils as well (Guo et al. 2021). With increasingly negative climate impacts, the pressure to enhance the terrestrial sink is growing, and one of the tools to achieve this is afforestation. As such, there is growing attention being placed globally to afforestation efforts, which is highlighted by the UN's declaration of the years from 2021 through 2030 as the UN Decade of Ecosystem Restoration (Sewell, Esch, and Lowenhardt 2020). Many of the Nationally Determined Contributions (NDCs) countries have submitted, which detail their plans to keep warming under the 2C threshold, also rely to a significant degree on afforestation projects (Roe et al. 2019).

The global gains that can be realized from afforestation, however, are uncertain, with broad estimates of cumulative sequestration ranging from 42 gigatonnes to 700 gigatonnes by the end of the century (Bastin et al. 2019a, 2019b; Humpenöder et al. 2014; Sathaye et al. 2006; Doelman et al. 2020). In addition, the amount of land conversion that would be required also varies greatly, with estimates ranging from 0.3 billion hectares to 2.8 billion hectares (Bastin et al. 2019a, 2019b; Humpenöder et al. 2014; Doelman et al. 2020). In many cases, these results suggest massive losses in the amount of pasture and cropland that would be available. This puts into question the feasibility of attaining these results, as almost 90 percent of deforestation worldwide is undertaken to expand the amount of agricultural land (FAO 2021). There are many reasons for this. For example, as nations develop and disposable incomes rise, consumers often change to more meat-based diets. Population growth is also a driver. Currently, the United Nations' medium variant population projections suggest that between 2023 and 2033 the worlds' population will grow by 8.8 percent, a gain of 707.1 million people (Nations 2023). And where population growth is especially high, which is the case at present throughout most of Africa (2023 to 2033 population growth: 25.0\%: 360.6 million), this puts extreme pressure on policymakers to allow greater conversion of forestland into farmland. Still, some research has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land (Griscoma et al. 2017).

If such efforts are to be made, it is important to understand where to target them, as afforestation is thought to not provide equal benefits in every region. For example, afforestation in boreal zones has been shown to provide the smallest net benefits, as the carbon sequestrations benefits of conifer forests are mitigated by the reduction in albedo in winter (Mykleby, Snyder, and Twine 2017; Li et al. 2015). Coniferous forests have also been shown to increase soil organic carbon less than broadleaf forests (Guo et al. 2021). Conversely, tropical forests have been linked with a strong positive result from afforestation, as the carbon sequestration benefit is complemented by a cooling impact due to changes in both albedo and evapotranspiration, and with temperate forests, the impact of afforestation is also shown to be positive, albeit less strongly so due to a winter warming effect (Li et al. 2015).

South America and Sub-Saharan Africa are thus two regions that show much potential for afforestation. It is estimated that these two regions hold at least 50\% of the potential for global gains while regions such as Northern Africa and the Middle East show little promise as forest growth rates are quite low (Doelman et al. 2020). Thus, it is promising that the African Forest Landscape Restoration Initiative (AFR100), which aims to afforest 100 million hectares by 2030, has roughly 70\% of its commitments coming from Sub-Saharan nations (AUDA-NEPAD 2023). While many of these nations may be well-suited in biophysical terms for afforestation, there is concern about how long-lasting any investment may be due to proximate zones of political instability. There are other nations that have a big potential for gains from afforestation, but they also face challenges. India, for example, has a great potential to increase its carbon sequestration through afforestation, but its massive population and food security worries limit its ability to consider such actions. The United Nations suggests that India's population will grow slightly less fast than the world average between 2023 and 2033 (+8.5\%), but due to its sizable base, that will equate to an increase in population of 120.5 million. Thus, its NDC does not entertain the idea of agricultural land being re-purposed and instead primarily considers targeting wastelands that may not be well suited for afforestation (Amjath-Babu et al. 2019). Moreover, as land-use decisions in one country also tend to have impacts that spillover into many others (Meyfroidt, Rudel, and Lambin 2010; Lambin and Meyfroidt 2011), analysis needs to be conducted at a global scale.

Several notable modeling attempts have been made in this direction. Steinbuks and Hertel (2012) developed their global partial equilibrium FABLE model (Forestry, Agriculture, Biofuels, Land Use, and Environment) to identify the optimal land use choices given the competing demands of meeting greenhouse gas targets and satisfying demand for food, bioenergy, forest products, and ecosystem services. While FABLE is a powerful tool for examining the optimal trajectory of various land uses under specific demand assumptions, it is focused on specific sectors of the economy and neglects general equilibrium and household welfare effects (Steinbuks and Hertel 2012; Hertel 2017). Other efforts have used the Global Trade Analysis Project (GTAP) computable general equilibrium model to address such questions, but its standard model does not well account for land-use changes. This prompted the creation of the Land Use and Land Cover Database within the GTAP framework, which incorporates forestry remote-sensing products (Baldos and Hertel 2012). The GTAP Agro-Ecological Zone (GTAP-AEZ) model further strengthened the GTAP model's ability to analyze land use changes, with the use of spatially explicit global land use data and through the incorporation of intra- and inter-regional land and land-based greenhouse emissions heterogeneity (Hertel et al. 2008). While the GTAP-AEZ model benefits from the inclusion of competition among different crops, grazing, and forest-based uses, it has limitations, primarily due to its infrequent updates. The \href{mailto:KLUM@GTAP}{\nolinkurl{KLUM@GTAP}} framework integrated the Kleines Land Use Model (KLUM) with an extended yet static version of GTAP called GTAP-EFL to evaluate the impact of climate change on cropland allocation. KLUM is a global agricultural land-use model that connects the economy to global crop allocation to maximize producer returns under specific risk assumptions, and GTAP-EFL separates energy factors from intermediate inputs and incorporates them into capital, while also considering CO2 emissions. \href{mailto:KLUM@GTAP}{\nolinkurl{KLUM@GTAP}} substitutes the land allocation mechanism within GTAP-EFL by utilizing regionally aggregated area changes in cropland determined by KLUM to update cropland shares in GTAP-EFL (Ronneberger and Tol 2009).

More recent modeling attempts of land-use change have begun to move away from comparative static analysis though. For example, the dynamic GTAP model, supported by the GTAP-AEZ database, considers land market effects, which were identified as significant in driving results by Stevenson (2013). This model has been utilized to evaluate the impact of crop intensification on land use. It, however, has been critiqued, as global aggregates may mask localized shifts that can have implications for ecologically significant areas (Byerlee 2014). In addition, in the standard GTAP model, it is difficult to model a sector individually. There are 68 market sectors covered in the latest version of the GTAP database. The GTAP model's general equilibrium nature means that all these markets are cleared with market-clearing prices. As the model applies nested CES production functions to all sectors and differentiates them by parameters only, to model one sector, in this case forestry, differently from the rest would require a major redesign of the model. While not impossible, this project will attempt to get around these concerns and assess the impact of various likely land-use scenarios by melding the dynamic GTAP model with the Global Timber Model (GTM), which is a dynamic optimization model that has been specifically designed to analyze the relationship among land rents, forested land cover, and carbon sequestration through forests and thus is better able to examine the impact of afforestation or avoided deforestation on the carbon sink (Sohngen, Mendelsohn, and Sedjo 2001). The latest versions of both models were used. For the GTAP model, this meant using data which has 2017 as its base year. This data was then updated to the base year (2023) using data from the Centre for Prospective Studies and International Information (CEPII) (Fontagné and Santoni 2022) to account for the projected changes in gross domestic product (GDP), population, and productivity for each country in the world. Similar steps were taken to update and recalibrate the GTM model.

\hypertarget{connecting-the-gtap-and-gtm-models}{%
\section{Connecting the GTAP and GTM models}\label{connecting-the-gtap-and-gtm-models}}

Both the GTAP and GTM models represent widely used models for policy analysis. In the case of GTAP, the model is used to describe the implications of various policy instruments for national and globla markets, including prices, output and trade. The GTM model is, on the other hand, capable to translating the impacts of population and global growth, and the changes in land returns, to changes in global forested areas and calculate the impacts on total carbon contained in forests.

Even though the GTAP and GTM models capture important aspects of the global economy and the biophysical world, they do not contain certain behaviors. In the case of GTAP, the model assumes that the amount of agricultural land is fixed, while the GTM considers the returns to land fixed. By combining the two models, we can fill each model missing behavior for a more complete description of the global impacts.

The combination of the two models means that we use the GTAP model to create the link between the amounts of available land and its returns for the benefit of the GTM model, while at the same time we use the GTM model to create the response of agricultural land quantity to returns in land returns for the benefit of the GTAP model. By solving both models together, we can greatly expand the descriptive power of both models.

We achieve the combined solution of GTAP and GTM by passing outputs back and forth between the models until a dual-state equilibrium is achieved. Establishing a connection between the GTAP model and the GTM model requires that we model the return on land using the GTAP model and feed those results into the GTM model as exogenous inputs. The GTM model then endogenously determines the amount of land available for crop and livestock production, and we feed that back into the GTAP model as an exogenous input.

On the software level, we establish the connection between the models using an R script, which is capable of handling other software, including generating required inputs, executing it and reading in any outputs. This is required to operate the GTAP model, which is written in GEMPACK, and the GTM model, which is written in GAMS. As GEMPACK and GAMS were developed long before software integration became important or even possible, we use much more modern technologies made available through the R software (R Core Team 2021) to create the user functions that execut both models, collect their outputs, and automatically create any required input files.

\hypertarget{changes-under-the-baseline-projections}{%
\section{Changes under the baseline projections}\label{changes-under-the-baseline-projections}}

To understand the impact of potential regional and global afforestation actions, an important first step was creating a realistic baseline scenario that depicts the likely changes over the decade under study, 2023--2033. Data for this were drawn from CEPII (Fontagné and Santoni 2022). To build our scenario, we use two of their variables that we determined to be key: population and GDP. The CEPII projections do not contain any information on land use. To build a more complete baseline, we also include the projections on forested land change included in the Global Timber Model (GTM). Using the GTAP model in connection with the GTM model (to assure that the changes in land are consistent between the two models), we calculate the overall change in factor quantity/productivity growth required to achieve the targeted values in 2033. In Table \ref{tab:baseline-growth}, we show the projected population and GDP growth per region. GDP increases for all regions of the world with the largest percentage changes occurring in South Asia (+308.6\%), China (+189.8\%), and Sub-Saharan Africa (+172.9\%). Population also is projected to grow in most regions with the largest changes happening in Sub-Saharan Africa (+55.7\%). Japan, Russia, and East Asia are projected to see their populations decline by 10.4\%, 4.7\%, and 2.6\%, respectively. As for the projected change in agricultural land, our baseline prjoections show significant growth in Oceania (+27.6\%), Rest of Latin America (+25.8\%), Sub-Saharan Africa (+17.4\%), Central America (+13.9\%), and Brazil (+13.9\%). Japan is the only nation that sees a significant decline (-34.9\%) in agricultural land, which is likely tied in part to the population decline just described.

\begin{table}
\centering
\caption{\label{tab:baseline-growth}GDP, population, and agricultural land growth under the baseline scenario (Source: CEPII and GTM)}
\centering
\begin{tabular}[t]{|>{}l|r|r|>{}r|}
\hline
  & Real GDP (\%) & Population (\%) & Agricultural land (\%)\\
\hline
Oceania & 52.7 & 24.0 & 27.6\\
\hline
China & 189.8 & 0.8 & 0.0\\
\hline
Japan & 20.6 & -10.4 & -34.9\\
\hline
East Asia & 46.0 & -2.6 & -1.1\\
\hline
SE Asia & 138.5 & 15.0 & 11.7\\
\hline
South Asia & 308.6 & 17.8 & 12.8\\
\hline
Canada & 41.3 & 15.2 & 2.6\\
\hline
United States & 38.6 & 10.6 & -0.5\\
\hline
Central America & 74.1 & 18.3 & 16.7\\
\hline
Brazil & 20.8 & 7.8 & 13.9\\
\hline
Other Latin America & 64.8 & 15.8 & 25.8\\
\hline
West/Central Europe & 25.4 & 0.9 & 4.5\\
\hline
EU II & 113.1 & 8.7 & 3.1\\
\hline
Russia & 41.8 & -4.7 & 4.4\\
\hline
Sub-Saharan Africa & 172.9 & 55.7 & 17.4\\
\hline
North Africa/Middle East & 58.2 & 27.8 & 7.8\\
\hline
ROW & 44.4 & 36.8 & 0.0\\
\hline
\end{tabular}
\end{table}

Focusing on the absolute area of the forested areas, as generated by the GTM model, we estimate that between 2023 and 2033, 144.2 million hectares (3.6\%) of global forests will be converted to agricultural lands with the largest decline expected to happen in Sub-Saharan Africa, which is estimated to lose 39.7 million hectares (7.0\%) of its forestland (Table \ref{tab:baseline-forest-change}). Southeast Asia, Brazil, and the rest of South America also see forestland shrink significantly, with decreases of 22.6 million hectares (9.3\%), 19.0 million hectares (3.8\%), and 23.5 million hectares (6.7\%), respectively. Conversely, the United States is estimated to add 1.3 million hectares (0.5\%) to its forestland.

\begin{table}
\centering
\begin{threeparttable}
\caption{\label{tab:baseline-forest-change}Forestland by region in 2023 and estimated change by 2033}
\centering
\begin{tabular}[t]{|>{}l|r|r|r|>{}r|}
\hline
  & Forestland in 2023 & Forestland in 2033 & $\Delta$ (Mha) & $\Delta$ (\%)\\
\hline
United States & 248.4 & 249.7 & 1.3 & 0.5\\
\hline
China & 164.6 & 164.8 & 0.1 & 0.1\\
\hline
Brazil & 498.2 & 479.2 & -19.0 & -3.8\\
\hline
Canada & 412.8 & 407.9 & -5.0 & -1.2\\
\hline
Russia & 838.1 & 828.6 & -9.5 & -1.1\\
\hline
West/Central Europe & 186.7 & 175.5 & -11.1 & -6.0\\
\hline
EU II & 40.1 & 36.1 & -4.1 & -10.1\\
\hline
South Asia & 47.8 & 42.5 & -5.3 & -11.0\\
\hline
Central America & 94.6 & 91.2 & -3.4 & -3.6\\
\hline
Other Latin America & 348.5 & 325.0 & -23.5 & -6.7\\
\hline
Sub-Saharan Africa & 564.6 & 525.0 & -39.7 & -7.0\\
\hline
SE Asia & 243.9 & 221.3 & -22.6 & -9.3\\
\hline
Oceania & 199.2 & 196.1 & -3.1 & -1.5\\
\hline
Japan & 23.7 & 26.9 & 3.2 & 13.7\\
\hline
North Africa/Middle East & 33.8 & 30.9 & -2.9 & -8.5\\
\hline
East Asia & 14.5 & 14.6 & 0.1 & 0.7\\
\hline
Total & 3959.5 & 3815.3 & -144.2 & -3.6\\
\hline
\end{tabular}
\begin{tablenotes}
\small
\item [a] Southeast Asia includes Brunei Darussalam, Cambodia, Indonesia, Laos, Malaysia, Myanmar, Philippines, Singapore, Thailand, Timor-Leste, and Vietnam.
\item [b] Oceania includes Australia, Cook Islands, Fiji, French Polynesia, Kiribati, Marshall Islands, Micronesia, Nauru, New Caledonia, New Zealand, Palau, Papua New Guinea, Samoa (Western Samoa), Solomon Islands, Tonga, Tuvalu, and Vanuatu.
\item [c] Rest of South America includes Argentina, Bolivia, Czech Republic, Chile, Colombia, Ecuador, the Falkland Islands, French Guiana, Guyana, Paraguay, Peru, South Georgia, the South Sandwich Islands, Suriname, Uruguay, and Venezuela.
\item [d] Western \& Central Europe includes  Austria, Belgium, Bulgaria, Croatia, Denmark, Estonia, Finland, France, Hungary, Iceland, Italy, Ireland, Germany, Greece, Latvia, Liechtenstein, Lithuania, Luxembourg, Malta, Monaco, Montenegro, Netherlands, North Macedonia, Norway, Poland, Portugal, Romania, Serbia, Slovakia, Slovenia, Spain, Sweden, Switzerland, Turkey, and the United Kingdom.
\item [e] Other Europe includes Albania, Armenia, Azerbaijan, Belarus, Bosnia and Herzegovina, Georgia,  Kosovo Kazakhstan, Kyrgyzstan, Moldova, Tajikistan, and Ukraine.
\end{tablenotes}
\end{threeparttable}
\end{table}

\hypertarget{modeling-the-impact-of-possible-measures-to-mitigate-deforestationpromote-afforestation}{%
\section{Modeling the impact of possible measures to mitigate deforestation/promote afforestation}\label{modeling-the-impact-of-possible-measures-to-mitigate-deforestationpromote-afforestation}}

Against the backdrop of the baseline scenario, we turn to the question of what policy could prevent global deforestation in the period of 2023--2033. Because the area of forestland is mostly determined by the relative returns of land in agriculture and forestry, in our policy scenario we modify the return to land in agriculture by applying a uniform global tax on the use of land in the production of agricultural output. Solving the two models together shows that it would require a tax of roughly 70 percent to be assessed on the use of land for the deforestation rate to fall globally to zero. The economic impact of such a policy measure would not be limited to the cost of the tax, it would also impact food production and its prices. For example, with this policy in place, most regions of the world would see significant declines in the amount of agricultural land in 2033 relative to the baseline scenario where land conversion was not taxed. For the United States, there would be 2.1 percent less agricultural land (Table \ref{tab:baseline-scenario1}). Larger decreases would happen in other major agricultural production regions such as China (-2.5 percent), South Asia (-7.9 percent), Brazil (-8.6 percent), and Russia (-9.3 percent).

These reductions in the amount of agricultural land would then impact the quantity of agricultural goods that could be produced. For example, the 15.0 percent decrease in agricultural land in East Asia and the 9.0 percent decrease in agricultural land in Southeast Asia would result in large declines in percentage terms in agricultural production of 5.0 percent and 1.7 percent, respectively. Surprisingly, Brazil and the United States, despite their decreases in available agricultural land compared to the baseline 2033 scenario, are estimated to see increased agricultural output of 0.6 percent and 1.0 percent, respectively. This may be tied to the sharp declines in other regions prompting greater intensity of production to satisfy external demand.

\begin{table}
\centering
\caption{\label{tab:baseline-scenario1}Changes in key variables relative to baseline under a uniform land tax of 70 \%}
\centering
\begin{tabular}[t]{|>{}l|r|r|r|>{}r|}
\hline
  & Real GDP (\%) & Agr. land (\%) & Agr. prices (\%) & Agr. output (\%)\\
\hline
Oceania & -0.1 & -23.9 & 2.8 & -2.5\\
\hline
China & 0.0 & -2.5 & 1.1 & 0.0\\
\hline
Japan & -0.3 & -60.8 & 7.8 & -6.3\\
\hline
East Asia & -0.2 & -15.0 & 5.4 & -5.0\\
\hline
SE Asia & -0.3 & -9.0 & 2.9 & -1.7\\
\hline
South Asia & -0.3 & -7.9 & 3.0 & -1.0\\
\hline
Canada & 0.0 & -7.9 & 0.7 & 1.9\\
\hline
United States & 0.0 & -2.1 & 0.7 & 1.0\\
\hline
Central America & -0.2 & -15.6 & 2.7 & -2.0\\
\hline
Brazil & -0.1 & -8.6 & 1.3 & 0.6\\
\hline
Other Latin America & -0.4 & -23.1 & 3.8 & -3.2\\
\hline
West/Central Europe & 0.0 & -6.1 & 1.0 & 0.7\\
\hline
EU II & -0.1 & -3.2 & 1.3 & 0.3\\
\hline
Russia & -0.1 & -9.3 & 1.1 & 0.3\\
\hline
Sub-Saharan Africa & -0.2 & -4.0 & 1.4 & -0.2\\
\hline
North Africa/Middle East & -0.1 & -10.2 & 2.0 & -0.7\\
\hline
ROW & 0.0 & 0.0 & 0.8 & 0.9\\
\hline
\end{tabular}
\end{table}

\begin{table}
\centering
\caption{\label{tab:unnamed-chunk-4}Percentage change in the amount of forested land in 2033 with deforestation mitigation policy in place (Source: Authors' calculations)}
\centering
\begin{tabular}[t]{|>{}l|r|r|r|>{}r|}
\hline
  & Forestland in 2023 & Forestland in 2033 & $\Delta$ (Mha) & $\Delta$ (\%)\\
\hline
United States & 248.4 & 254.8 & 6.4 & 2.6\\
\hline
China & 164.6 & 174.8 & 10.2 & 6.2\\
\hline
Brazil & 498.2 & 492.2 & -6.0 & -1.2\\
\hline
Canada & 412.8 & 423.2 & 10.4 & 2.5\\
\hline
Russia & 838.1 & 849.7 & 11.6 & 1.4\\
\hline
West/Central Europe & 186.7 & 191.1 & 4.5 & 2.4\\
\hline
EU II & 40.1 & 40.3 & 0.2 & 0.4\\
\hline
South Asia & 47.8 & 46.2 & -1.6 & -3.3\\
\hline
Central America & 94.6 & 94.9 & 0.3 & 0.3\\
\hline
Other Latin America & 348.5 & 348.1 & -0.3 & -0.1\\
\hline
Sub-Saharan Africa & 564.6 & 535.6 & -29.0 & -5.1\\
\hline
SE Asia & 243.9 & 240.5 & -3.5 & -1.4\\
\hline
Oceania & 199.2 & 199.0 & -0.1 & -0.1\\
\hline
Japan & 23.7 & 23.5 & -0.1 & -0.6\\
\hline
North Africa/Middle East & 33.8 & 34.9 & 1.1 & 3.4\\
\hline
East Asia & 14.5 & 15.8 & 1.3 & 9.3\\
\hline
Total & 3959.5 & 3964.9 & 5.3 & 0.1\\
\hline
\end{tabular}
\end{table}

When we examine how this impacts exports of agricultural goods, we find that exports from Japan and the United States grow the most, increasing 8.8 percent and 3.5 percent, respectively (Table 4). The regions where agricultural exports drop most significantly are East Asia, South Asia, Southeast Asia, and Central America who see decreases of 14.5 percent, 5.9 percent, 5.1 percent, and 5.0 percent, respectively.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-5-1.pdf}
\caption{\label{fig:unnamed-chunk-5}Changes in ag trade volumes (red arrows denote reductions, green arrows denote increases)}
\end{figure}

Overall, this tax policy, which results in zero net deforestation globally, has a rather significant negative economic impact on all regions except Japan and Canada who see an increase in consumer well-being of \$977.1 million and \$166.9 million, respectively (Table \ref{tab:EV}). Consumer well-being is a measurement of equivalent variation and here measures the benefit/harm consumers experience in terms of income as a result of the policy changes. The United States, along with 10 other regions, see well-being drop by more than \$1 billion. The most negatively impacted regions are South Asia (-\$28.0 billion), China (-\$14.5 billion), and Southeast Asia (-\$13.6 billion). In total, this tax-driven policy is estimated to decrease global consumer well-being in 2033 by \$92.1 billion relative to the scenario where no tax on forestland conversion was in place.

\begin{table}
\centering
\caption{\label{tab:EV}Change in consumer well-being in 2033, by region, in billions dollars, with deforestation mitigation policy in place}
\centering
\begin{tabular}[t]{|>{}l|r|r|>{}r|}
\hline
  & Baseline (B\$) & Scenario (B\$) & $\Delta$ (B\$)\\
\hline
Oceania & 863.7 & 861.8 & -1.9\\
\hline
China & 23,275.9 & 23,260.0 & -15.9\\
\hline
Japan & 1,052.6 & 1,036.1 & -16.5\\
\hline
East Asia & 1,168.4 & 1,163.0 & -5.4\\
\hline
SE Asia & 3,886.2 & 3,872.4 & -13.9\\
\hline
South Asia & 9,493.4 & 9,465.6 & -27.8\\
\hline
Canada & 713.1 & 713.5 & 0.4\\
\hline
United States & 7,497.2 & 7,496.4 & -0.9\\
\hline
Central America & 1,241.2 & 1,236.4 & -4.8\\
\hline
Brazil & 423.0 & 422.4 & -0.6\\
\hline
Other Latin America & 1,242.6 & 1,233.2 & -9.4\\
\hline
West/Central Europe & 5,126.7 & 5,118.8 & -8.0\\
\hline
EU II & 839.3 & 838.7 & -0.6\\
\hline
Russia & 706.2 & 704.9 & -1.3\\
\hline
Sub-Saharan Africa & 2,766.7 & 2,758.8 & -7.8\\
\hline
North Africa/Middle East & 1,715.9 & 1,710.7 & -5.3\\
\hline
ROW & 13.9 & 13.9 & 0.0\\
\hline
Total & 62,026.3 & 61,906.6 & -119.7\\
\hline
\end{tabular}
\end{table}

Much of this loss in well-being may be the result of higher prices. Japan is the only nation that sees any price impacts that are beneficial for consumers resulting from this policy, with the price of agricultural goods decreasing 0.3 percent (Table \ref{tab:pm}). All other regions, for both agricultural and non-agricultural products, see prices rise between 0.3 and 5.1 percent.

\begin{table}
\centering
\caption{\label{tab:pm}Percent change in market prices for agricultural goods in 2033, with deforestation mitigation policy in place, relative to baseline}
\centering
\begin{tabular}[t]{|>{}l|>{}r|}
\hline
  & Ag prices\\
\hline
Oceania & 2.8\\
\hline
China & 1.1\\
\hline
Japan & 7.8\\
\hline
East Asia & 5.4\\
\hline
SE Asia & 2.9\\
\hline
South Asia & 3.0\\
\hline
Canada & 0.7\\
\hline
United States & 0.7\\
\hline
Central America & 2.7\\
\hline
Brazil & 1.3\\
\hline
Other Latin America & 3.8\\
\hline
West/Central Europe & 1.0\\
\hline
EU II & 1.3\\
\hline
Russia & 1.1\\
\hline
Sub-Saharan Africa & 1.4\\
\hline
North Africa/Middle East & 2.0\\
\hline
ROW & 0.8\\
\hline
\end{tabular}
\end{table}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

This analysis shows that, at an estimated tax rate of 70.3 percent of the value of each hectare of land that is converted to agricultural land, global net deforestation would drop to roughly zero by 2033, preserving 144.2 million hectares of forestland that otherwise would have been converted to agricultural land. In the United States, as USDA-NASS (Land Values 2022 Summary (August 2022)) valued the average hectare of farmland in 2022 at \$1,537.8, the tax would equate to \$1,081.1/hectare. In Brazil, CEIC valued the average price of Brazilian land at roughly 2,500 Brazilian Reals, which would put the tax at roughly \$82.2 per hectare. The economic price of this would likely be transfered to consumers through less agricultural output and higher market prices. Overall, such an action is estimated to reduce consumer well-being by \$92.1 billion or \$10.6 per every person in 2033. Governments might be wary of farmer discontent were they to take such fiscal actions and thus might be reluctant to impose such taxes on them. Thus, policymakers who wish to see deforestation stopped might have to resort to import tariffs on products produced on land converted from forestland for agricultural use.

We also show the utility of joining two major economic models. GTAP has long been valued for its ability to provide robust information on global trade flows and prices, and GTM has proven its utility in understanding the impact of forestry decisions on the carbon sink and land rents. By combining these models through the use of R and allowing the models to pass inputs and outputs back and forth iteratively, the benefits of both models have been maintained and their weaknesses greatly minimized. This advance holds great promise for advancement on a variety of fronts for researchers and policymakers interested in climate change, agricultural trade, and their inter-dependencies.

\hypertarget{conclusions}{%
\section{Conclusions}\label{conclusions}}

According to CEPII's projections, the growth of the global economy in the next ten years is expected to result in tremendous increases in global and regional GDP accompanied with a moderate growth in population. Another set of projections by GTM, however, suggests that this growth would also mean significant reductions in the world's forests that would be converted to agricultural land, resulting in a loss of over 100 million hectares of forest before 2033.

By combining two well known global models---the GTAP model that describes the world markets and trade, and the GTAM model that describes the impact of economic and population growth along with land rents on the amounts of forested areas---we were able to estimate the size of the tax on agricultural land needed to prevent deforestation in the coming ten years through reducing the returns to land when used in agriculture. We find that even though the tax is substantial at about 70 percent, it results in only very small reductions in the projected growth in global GDP and welfare. We also observe that the reduction in available agricultural land relative to the baseline growth results in reduced agricultural output and higher prices, which may require additional policy action to avoid negative impacts in terms of food security.

Finally, we note that a uniform global tax reducing the returns to land to the point of preventing global deforestation would produce highly differentiated impacts around the world's regions. Some regions, such as China, Russia, Canada and the United States would increase their forested areas as a result of the tax, while the regions with the greatest deforestation pressures, such as Sub-Saharan Africa, Brazil and Southeast Asia would only reduce the rate of their deforestation, for a zero change in the global forest area.

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-amjath2019}{}}%
Amjath-Babu, T. S., Pramod K. Aggarwal, Sonja Vermeulen Hans van Meijl, and Paul L. Lucas. 2019. {``Climate Action for Food Security in South Asia? Analyzing the Role of Agriculture in Nationally Determined Contributions to the Paris Agreement.''} \emph{Climate Policy} 19 (3). https://doi.org/\url{https://doi.org/10.1080/14693062.2018.1501329}.

\leavevmode\vadjust pre{\hypertarget{ref-auda2023}{}}%
AUDA-NEPAD. 2023. {``AFR100.''} \url{https://afr100.org}.

\leavevmode\vadjust pre{\hypertarget{ref-baldos2012}{}}%
Baldos, U. L. C., and Thomas W. Hertel. 2012. {``Development of a GTAP 8 Land Use and Land Cover Database for Years 2004 and 2007.''} \emph{GTAP Research Memorandum No. 23. Global Trade Analysis Project}.

\leavevmode\vadjust pre{\hypertarget{ref-veldman2019}{}}%
Bastin, Jean-Francois, Yelena Finegold, Claude Garcia, Danilo Mollicone, Marcelo Rezende, Devin Routh, Constantin M. Zohner, and Thomas W. Crowther. 2019a. {``Comment on "the Global Tree Restoration Potential".''} \emph{Science} 366 (6463).

\leavevmode\vadjust pre{\hypertarget{ref-bastin2019}{}}%
---------. 2019b. {``The Global Tree Restoration Potential.''} \emph{Science} 365 (6448).

\leavevmode\vadjust pre{\hypertarget{ref-byerlee2014}{}}%
Byerlee, Stevenson, D. 2014. {``Does Intensification Slow Crop Land Expansion or Encourage Deforestation?''} \emph{Global Food Security} 3 (2).

\leavevmode\vadjust pre{\hypertarget{ref-davis2022}{}}%
Davis, Eric C., Brent Sohngen, and David J. Lewis. 2022. {``The Effect of Carbon Fertilization on Naturally Regenerated and Planted US Forests.''} \emph{Nature Communications} 13 (1). https://doi.org/\url{https://doi.org/10.1038/s41467-022-33196-x}.

\leavevmode\vadjust pre{\hypertarget{ref-doelman2020}{}}%
Doelman, Jonathan C., Elke Stehfest, Detlef P. van Vuuren, Andrzej Tabeau, Andries F. Hof Maarten, C. Braakhekke, David E. H. J. Gernaat, et al. 2020. {``Afforestation for Climate Change Mitigation: Potentials, Risks and Trade‐offs.''} \emph{Global Change Biology} 26 (3).

\leavevmode\vadjust pre{\hypertarget{ref-fao2021}{}}%
FAO. 2021. {``COP26: Agricultural Expansion Drives Almost 90 Percent of Global Deforestation.''} \url{https://www.fao.org/newsroom/detail/cop26-agricultural-expansion-drives-almost-90-percent-of-global-deforestation/en}.

\leavevmode\vadjust pre{\hypertarget{ref-fao2023}{}}%
FAO, and UNEP. 2020. {``The State of the World's Forests 2020. Forests, Biodiversity and People.''} https://doi.org/\url{https://doi.org/10.4060/ca8642en}.

\leavevmode\vadjust pre{\hypertarget{ref-CEPII2024}{}}%
Fontagné, Perego, L., and G. Santoni. 2022. {``MaGE 3.1: Long-Term Macroeconomic Projections of the World Economy.''} \emph{International Economics} 172.

\leavevmode\vadjust pre{\hypertarget{ref-griscom2017}{}}%
Griscoma, Bronson W., Justin Adamsa, Peter W. Ellis, Richard A. Houghton, Guy Lomax, Daniela A. Miteva, William H. Schlesinger, et al. 2017. {``Natural Climate Solutions.''} \emph{Proceedings of the National Academy of Sciences} 114 (44).

\leavevmode\vadjust pre{\hypertarget{ref-guo2021}{}}%
Guo, Yang, Mohamed Abdalla, Mikk Espenberg, Astley Hastings, Paul Hallett, and Pete Smith. 2021. {``A Systematic Analysis and Review of the Impacts of Afforestation on Soil Quality Indicators as Modified by Climate Zone, Forest Type and Age.''} \emph{Science of the Total Environment} 757.

\leavevmode\vadjust pre{\hypertarget{ref-hertel2017}{}}%
Hertel, Thomas W. 2017. {``Land Use in the 21st Century: Contributing to the Global Public Good.''} \emph{Review of Development Economics} 21 (2).

\leavevmode\vadjust pre{\hypertarget{ref-hertel2008}{}}%
Hertel, Thomas W., L. Huey-Lin, S. Rose, and Brent Sohngen. 2008. {``Modeling Land-Use Related Greenhouse Gas Sources and Sinks and Their Mitigation Potential.''} \emph{GTAP Working Paper No. 44. Global Trade Analysis Project}.

\leavevmode\vadjust pre{\hypertarget{ref-humpenoder2014}{}}%
Humpenöder, Florian, Alexander Popp, Jan Philip Dietrich, David Klein, Hermann Lotze-Campen, Markus Bonsch, Benjamin, et al. 2014. {``Investigating Afforestation and Bioenergy CCS as Climate Change Mitigation Strategies.''} \emph{Environmental Research Letters} 9 (6).

\leavevmode\vadjust pre{\hypertarget{ref-lambin2011}{}}%
Lambin, Eric F., and Patrick Meyfroidt. 2011. {``Global Land Use Change, Economic Globalization, and the Looming Land Scarcity.''} \emph{Proceedings of the National Academy of Sciences} 108 (9).

\leavevmode\vadjust pre{\hypertarget{ref-li2015}{}}%
Li, Yan, Maosheng Zhao, Safa Motesharrei, Qiaozhen Mu, Eugenia Kalnay, and Shuangcheng Li. 2015. {``Local Cooling and Warming Effects of Forests Based on Satellite Observations.''} \emph{Nature Communications} 6 (1). https://doi.org/\url{https://doi.org/10.1038/ncomms76031}.

\leavevmode\vadjust pre{\hypertarget{ref-meyfroidt2010}{}}%
Meyfroidt, Patrick, Thomas K. Rudel, and Eric F. Lambin. 2010. {``Forest Transitions, Trade, and the Global Displacement of Land Use.''} \emph{Proceedings of the National Academy of Sciences} 107 (49).

\leavevmode\vadjust pre{\hypertarget{ref-mykleby2017}{}}%
Mykleby, PM, PK Snyder, and TE Twine. 2017. {``The Global Tree Restoration Potential.''} \emph{Geophysical Research Letters} 44 (5).

\leavevmode\vadjust pre{\hypertarget{ref-un2024}{}}%
Nations, United. 2023. {``World Population Prospects (2023).''} \url{https://population.un.org/wpp/}.

\leavevmode\vadjust pre{\hypertarget{ref-R2021}{}}%
R Core Team. 2021. \emph{R: A Language and Environment for Statistical Computing}. Vienna, Austria: R Foundation for Statistical Computing. \url{https://www.R-project.org/}.

\leavevmode\vadjust pre{\hypertarget{ref-roe2019}{}}%
Roe, Stephanie, Charlotte Streck, Michael Obersteiner, Stefan Frank, Bronson Griscom, Laurent Drouet, Oliver Fricko, et al. 2019. {``Contribution of the Land Sector to a 1.5 c World.''} \emph{Nature Climate Change} 9 (11).

\leavevmode\vadjust pre{\hypertarget{ref-ronneberger2009}{}}%
Ronneberger, Berrittella, K., and R. S. Tol. 2009. {``KLUM@ GTAP: Introducing Biophysical Aspects of Land-Use Decisions into a Computable General Equilibrium Model a Coupling Experiment.''} \emph{Environmental Modeling \& Assessment} 14.

\leavevmode\vadjust pre{\hypertarget{ref-sathaye2006}{}}%
Sathaye, Jayant, Willy Makundi, Larry Dale, Peter Chan, and Kenneth Andrasko. 2006. {``GHG Mitigation Potential, Costs and Benefits in Global Forests: A Dynamic Partial Equilibrium Approach.''} \emph{The Energy Journal}.

\leavevmode\vadjust pre{\hypertarget{ref-sewell2020}{}}%
Sewell, Annelies, Stefan van der Esch, and Hannah Lowenhardt. 2020. {``Goals and Committments for the Restoration Decade.''} \emph{The Hague: PBL Netherlands Environmental Assessment Agency}.

\leavevmode\vadjust pre{\hypertarget{ref-sohngen2001}{}}%
Sohngen, Brent, Robert Mendelsohn, and Roger Sedjo. 2001. {``A Global Model of Climate Change Impacts on Timber Markets.''} \emph{Journal of Agricultural and Resource Economics} 26 (2).

\leavevmode\vadjust pre{\hypertarget{ref-steinbuks2012}{}}%
Steinbuks, Jevgenijs, and Thomas W. Hertel. 2012. {``Forest, Agriculture, and Biofuels in a Land Use Model with Environmental Services (FABLE).''} \emph{GTAP Working Paper No. 71. GTAP}.

\leavevmode\vadjust pre{\hypertarget{ref-stevenson2013}{}}%
Stevenson, Villoria, J. R. 2013. {``Green Revolution Research Saved an Estimated 18 to 27 Million Hectares from Being Brought into Agricultural Production.''} \emph{Proceedings of the National Academy of Sciences} 110 (21).

\end{CSLReferences}

\end{document}

### 2024-03-03 paper.Rmd
Update paper.Rmd

@@
 There are many reasons for this.
-For example, as nations develop and disposable incomes rise, consumers often change to more meat-based diets.
+For example, as nations develop and disposable incomes rise, consumers often increase meat consumpution.
 Population growth is also a driver.
 Currently, the United Nations' medium variant population projections suggest that between 2023 and 2033 the worlds' population will grow by 8.8 percent, a gain of 707.1 million people [@un2024].
-And where population growth is especially high, which is the case at present throughout most of Africa (2023 to 2033 population growth: 25.0%: 360.6 million), this puts extreme pressure on policymakers to allow greater conversion of forestland into farmland.
+And where population growth is especially high, which is the case at present throughout Africa (2023 to 2033 population growth: 25.0%: 360.6 million), there is likely to be greater conversion of forestland into farmland.
 Still, some research has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017].

### 2024-03-03 paper.Rmd
Update paper.Rmd

@@
 
-If such efforts are to be made, it is important to understand where to target them, as afforestation is thought to not provide equal benefits in every region.
-For example, afforestation in boreal zones has been shown to provide the smallest net benefits, as the carbon sequestrations benefits of conifer forests are mitigated by the reduction in albedo in winter [@mykleby2017; @li2015].
+If such efforts are to be made, it is important to understand where to target them, as afforestation does not provide equal carbon benefits in every region.
+For example, afforestation in boreal zones provides the smallest net benefits, as the carbon sequestration benefits of conifer forests are mitigated by the reduction in albedo in winter [@mykleby2017; @li2015].
 Coniferous forests have also been shown to increase soil organic carbon less than broadleaf forests [@guo2021].
@@
 Thus, it is promising that the African Forest Landscape Restoration Initiative (AFR100), which aims to afforest 100 million hectares by 2030, has roughly 70% of its commitments coming from Sub-Saharan nations [@auda2023].
-While many of these nations may be well-suited in biophysical terms for afforestation, there is concern about how long-lasting any investment may be due to proximate zones of political instability.
-There are other nations that have a big potential for gains from afforestation, but they also face challenges.
+There are other nations that have a big potential for gains from afforestation, but they face challenges.
 India, for example, has a great potential to increase its carbon sequestration through afforestation, but its massive population and food security worries limit its ability to consider such actions.
-The United Nations suggests that India's population will grow slightly less fast than the world average between 2023 and 2033 (+8.5%), but due to its sizable base, that will equate to an increase in population of 120.5 million.
-Thus, its NDC does not entertain the idea of agricultural land being re-purposed and instead primarily considers targeting wastelands that may not be well suited for afforestation [@amjath2019].
+The United Nations suggests that India's population will grow less rapidly than the world average between 2023 and 2033 (+8.5%), but due to its sizable base, the increase will be substantial nonetheless: 120.5 million.
+India's NDC does not entertain the idea of agricultural land being re-purposed and instead primarily considers targeting land that may not be well suited for afforestation [@amjath2019].
 Moreover, as land-use decisions in one country also tend to have impacts that spillover into many others [@meyfroidt2010; @lambin2011], analysis needs to be conducted at a global scale.
@@
 Similar steps were taken to update and recalibrate the GTM model.
+As most forest inventory data are not updated annually, the GTM model updates data for regions with inventory information approximately every five years, and it updates data for other regions using the Global Forest Resource Assessment [@FAO2020].
 

### 2024-03-03 paper.Rmd
Update paper.Rmd

@@
 Both the GTAP and GTM models represent widely used models for policy analysis.
-In the case of GTAP, the model is used to describe the implications of various policy instruments for national and globla markets, including prices, output and trade.
-The GTM model is, on the other hand, capable to translating the impacts of population and global growth, and the changes in land returns, to changes in global forested areas and calculate the impacts on total carbon contained in forests.
+In the case of GTAP, the model is used to describe the implications of various policy instruments for national and global markets, including prices, output and trade.
+The GTM model is, on the other hand, capable to translating the impacts of population and global growth, and the changes in land returns, to changes in global forest stocks and calculate the impacts on total carbon contained in forests.
 
 Even though the GTAP and GTM models capture important aspects of the global economy and the biophysical world, they do not contain certain behaviors.
-In the case of GTAP, the model assumes that the amount of agricultural land is fixed, while the GTM considers the returns to land fixed.
+In the case of GTAP, the model assumes that the amount of agricultural land is fixed, while the GTM considers the agricultural returns to land fixed.
 By combining the two models, we can fill each model missing behavior for a more complete description of the global impacts.
 
-The combination of the two models means that we use the GTAP model to create the link between the amounts of available land and its returns for the benefit of the GTM model, while at the same time we use the GTM model to create the response of  agricultural land quantity to returns in land returns for the benefit of the GTAP model.
-By solving both models together, we can greatly expand the descriptive power of both models.
+The combination of the two models means that we use the GTAP model to create a link between the amount of available land and its returns for the benefit of the GTM model, while at the same time we use the GTM model to create the response of  agricultural land quantity to returns in land returns for the benefit of the GTAP model.
+By solving the models together, we can expand the descriptive power of both.
 
-We achieve the combined solution of GTAP and GTM by passing outputs back and forth between the models until a dual-state equilibrium is achieved.
+We achieve a combined solution of GTAP and GTM by passing outputs back and forth between the models until a dual-state equilibrium is achieved.
 Establishing a connection between the GTAP model and the GTM model requires that we  model the return on land using the GTAP model and feed those results into the GTM model as exogenous inputs.
@@
 This is required to operate the GTAP model, which is written in GEMPACK, and the GTM model, which is written in GAMS.
-As GEMPACK and GAMS were developed long before software integration became important or even possible, we use much more modern technologies made available through the R software  [@R2021] to create the user functions that execut both models, collect their outputs, and automatically create any required input files.
+As GEMPACK and GAMS were developed long before software integration became important or even possible, we use much more modern technologies made available through the R software [@R2021] to create the user functions that execute both models, collect their outputs, and automatically create any required input files.
 

### 2024-03-03 paper.Rmd
Update paper.Rmd

@@
 The CEPII projections do not contain any information on land use.
-To build a more complete baseline, we also include the projections on forested land change included in the Global Timber Model (GTM).
+To build a more complete baseline, we also include the projections on forested land change included in the GTM.
 Using the GTAP model in connection with the GTM model (to assure that the changes in land are consistent between the two models), we calculate the overall change in factor quantity/productivity growth required to achieve the targeted values in 2033.
@@
 Japan, Russia, and East Asia are projected to see their populations decline by 10.4%, 4.7%, and 2.6%, respectively.
-As for the projected change in agricultural land, our baseline prjoections show significant growth in Oceania (+27.6%), Rest of Latin America (+25.8%), Sub-Saharan Africa (+17.4%), Central America (+13.9%), and Brazil (+13.9%).
-Japan is the only nation that sees a significant decline (-34.9%) in agricultural land, which is likely tied in part to the population decline just described.
+As for the projected change in agricultural land, our baseline projections show a significant increase in Oceania (+27.6%), Rest of Latin America (+25.8%), Sub-Saharan Africa (+17.4%), Central America (+13.9%), and Brazil (+13.9%).
+Japan is the only nation that sees a significant decline (-34.9%) in agricultural land, which is likely tied in part to population decline.
 
@@
 Focusing on the absolute area of the forested areas, as generated by the GTM model, we estimate that between 2023 and 2033, 144.2 million hectares (3.6%) of global forests will be converted to agricultural lands with the largest decline expected to happen in Sub-Saharan Africa, which is estimated to lose 39.7 million hectares (7.0%) of its forestland (Table \@ref(tab:baseline-forest-change)).
-Southeast Asia, Brazil, and the rest of South America also see forestland shrink significantly, with decreases of 22.6 million hectares (9.3%), 19.0 million hectares (3.8%), and 23.5 million hectares (6.7%), respectively.
+Southeast Asia, Brazil, and the rest of South America also see forestland shrink, with decreases of 22.6 million hectares (9.3%), 19.0 million hectares (3.8%), and 23.5 million hectares (6.7%), respectively.
 Conversely, the United States is estimated to add 1.3 million hectares (0.5%) to its forestland.
@@
           label = 'baseline-forest-change',
-          caption = 'Forestland by region in 2023 and estimated change by 2033'
+          caption = 'Forestland by region in 2023 and estimated change in 2033 relative to 2033 baseline scenario'
         ),
@@
 
-Against the backdrop of the baseline scenario, we turn to the question of what policy could prevent global deforestation in the period of 2023--2033.
+Against the backdrop of the baseline scenario, we turn to the question of what policy could prevent global deforestation in the period of 2023–2033.
 Because the area of forestland is mostly determined by the relative returns of land in agriculture and forestry, in our policy scenario we modify the return to land in agriculture by applying a uniform global tax on the use of land in the production of agricultural output.
@@
 For example, with this policy in place, most regions of the world would see significant declines in the amount of agricultural land in 2033 relative to the baseline scenario where land conversion was not taxed.
-For the United States, there would be 2.1 percent less agricultural land (Table \@ref(tab:baseline-scenario1)).
+For the United States, there would be 2.1 percent less agricultural land (Table 3).
 Larger decreases would happen in other major agricultural production regions such as China (-2.5 percent), South Asia (-7.9 percent), Brazil (-8.6 percent), and Russia (-9.3 percent).
@@
 These reductions in the amount of agricultural land would then impact the quantity of agricultural goods that could be produced.
-For example, the 15.0 percent decrease in agricultural land in East Asia and the 9.0 percent decrease in agricultural land in Southeast Asia would result in large declines in percentage terms in agricultural production of 5.0 percent and 1.7 percent, respectively.
+For example, the 15.0 percent decrease in agricultural land in East Asia and the 9.0 percent decrease in agricultural land in Southeast Asia would result in reductions in agricultural production of 5.0 percent and 1.7 percent, respectively.
 Surprisingly, Brazil and the United States, despite their decreases in available agricultural land compared to the baseline 2033 scenario, are estimated to see increased agricultural output of 0.6 percent and 1.0 percent, respectively.
-This may be tied to the sharp declines in other regions prompting greater intensity of production to satisfy external demand.
+This happens when the sharp output decline in other regions prompts increased intensity of production in the United States and Brazil.
 
@@
 
-When we examine how this impacts exports of agricultural goods, we find that exports from Japan and the United States grow the most, increasing 8.8 percent and 3.5 percent, respectively (Table 4).
-The regions where agricultural exports drop most significantly are East Asia, South Asia, Southeast Asia, and Central America who see decreases of 14.5 percent, 5.9 percent, 5.1 percent, and 5.0 percent, respectively.
+When we examine how the policy impacts exports of agricultural goods, we find that exports from Japan and the United States grow the most, increasing 8.8 percent and 3.5 percent, respectively (Table 4).
+The regions where agricultural exports drop most are East Asia, South Asia, Southeast Asia, and Central America, which see decreases of 14.5 percent, 5.9 percent, 5.1 percent, and 5.0 percent, respectively.
 
@@
 
-Overall, this tax policy, which results in zero net deforestation globally, has a rather significant negative economic impact on all regions except Japan and Canada who see an increase in consumer well-being of $977.1 million and $166.9 million, respectively (Table \@ref(tab:EV)).
+Overall, the tax policy, which results in zero net deforestation globally, has a rather substantial negative economic impact on all regions except Japan and Canada, which see an increase in consumer well-being of
+$977.1 million and $166.9 million, respectively (Table 5).
 Consumer well-being is a measurement of equivalent variation and here measures the benefit/harm consumers experience in terms of income as a result of the policy changes.

### 2024-03-03 paper.Rmd
Update paper.Rmd

@@
   kableExtra::column_spec(kableExtra::kable_styling(
-    knitr::kable(out, caption = 'Changes in key variables relative to baseline under a uniform land tax of 70 \\%', digits = 1, escape = FALSE, label = 'baseline-scenario1')
+    knitr::kable(out, caption = 'Changes in key variables in 2033 under a uniform land tax of 70 and relative to 2033 baseline scenario \\%', digits = 1, escape = FALSE, label = 'baseline-scenario1')
   ), 1, border_left = TRUE),
@@
           digits = 1,
-          caption = 'Percentage change in the amount of forested land in 2033 with deforestation mitigation policy in place (Source: Authors\' calculations)'
+          caption = 'Percentage change in the amount of forested land in 2033 with deforestation mitigation policy in place relative to 2033 baseline scenario (Source: Authors\' calculations)'
         )
@@
   kableExtra::column_spec(kableExtra::kable_styling(
-    knitr::kable(out, caption = 'Change in consumer well-being in 2033, by region, in billions dollars, with deforestation mitigation policy in place', digits = 1, escape = FALSE, label = 'EV',format.args = list(big.mark = ","))
+    knitr::kable(out, caption = 'Change in consumer well-being in 2033, by region, in billions of U.S.
+dollars, with deforestation mitigation policy in place relative to 2033 baseline scenario', digits = 1, escape = FALSE, label = 'EV',format.args = list(big.mark = ","))
   ), 1, border_left = TRUE),
@@
   kableExtra::column_spec(kableExtra::kable_styling(
-    knitr::kable(out, caption = 'Percent change in market prices for agricultural goods in 2033, with deforestation mitigation policy in place, relative to baseline', digits = 1, escape = FALSE, label = 'pm')
+    knitr::kable(out, caption = 'Percent change in market prices for agricultural goods in 2033, with deforestation mitigation policy in place, relative to 2033 baseline scenario', digits = 1, escape = FALSE, label = 'pm')
   ), 1, border_left = TRUE),
@@
 In Brazil, CEIC valued the average price of Brazilian land at roughly 2,500 Brazilian Reals, which would put the tax at roughly $82.2 per hectare.
-The economic price of this would likely be transfered to consumers through less agricultural output and higher market prices.
-Overall, such an action is estimated to reduce consumer well-being by $92.1 billion or $10.6 per every person in 2033.
+The economic price of this would be transfered to consumers through lower agricultural output and higher market prices.
+Overall, this land tax would reduce consumer well-being by $92.1 billion or $10.60 per person in 2033.
 Governments might be wary of farmer discontent were they to take such fiscal actions and thus might be reluctant to impose such taxes on them.
-Thus, policymakers who wish to see deforestation stopped might have to resort to import tariffs on products produced on land converted from forestland for agricultural use.
+Thus, if policymakers wish to see deforestation stopped, they could instead use import tariffs on products produced on land converted from forestland for agricultural use.
 
@@
 
-According to CEPII's projections, the growth of the global economy in the next ten years is expected to result in tremendous increases in global and regional GDP accompanied with a moderate growth in population.
-Another set of projections by GTM, however, suggests that this growth would also mean significant reductions in the world's forests that would be converted to agricultural land, resulting in a loss of over 100 million hectares of forest before 2033.
+According to CEPII's projections, the growth of the global economy in the next ten years is expected to result in tremendous increases in global and regional GDP accompanied by a moderate growth in population.
+Another set of projections by GTM suggests that this growth would also mean significant reductions in the world's forests because they would be converted to agricultural land.
+The model predicts a loss of over 144 million hectares of forest before 2033.
 
-By combining two well known global models---the GTAP model that describes the world markets and trade, and the GTAM model that describes the impact of economic and population growth along with land rents on the amounts of forested areas---we were able to estimate the size of the tax on agricultural land needed to prevent deforestation in the coming ten years through reducing the returns to land when used in agriculture.
+By combining two well-known global models---the GTAP model that describes the world markets and trade, and the GTAM model that describes the impact of economic and population growth along with land rents on the amounts of forested areas---we were able to estimate the size of the tax on agricultural land needed to prevent deforestation in the coming ten years through reducing the returns to land when used in agriculture.
 We find that even though the tax is substantial at about 70 percent, it results in only very small reductions in the projected growth in global GDP and welfare.

### 2024-03-08 paper.Rmd
Update paper.Rmd

@@
 
-Even though the GTAP and GTM models capture important aspects of the global economy and the biophysical world, they do not contain certain behaviors.
+Even though the GTAP and GTM models capture important aspects of the global economy and the biophysical world, they do not model certain economic responses, and, instead, consider several important variables exogenous.
 In the case of GTAP, the model assumes that the amount of agricultural land is fixed, while the GTM considers the agricultural returns to land fixed.
-By combining the two models, we can fill each model missing behavior for a more complete description of the global impacts.
+By combining the two models, we can fill each model's missing behavior for a more complete description of the global impacts.
 
-The combination of the two models means that we use the GTAP model to create a link between the amount of available land and its returns for the benefit of the GTM model, while at the same time we use the GTM model to create the response of  agricultural land quantity to returns in land returns for the benefit of the GTAP model.
-By solving the models together, we can expand the descriptive power of both.
+The combination of the two models means that we use the GTAP model to create a link between the amount of available land and its returns for the benefit of the GTM model, which uses returns as an exogenous input, while at the same time we use the GTM model to create the response of  agricultural land quantity to returns in land returns for the benefit of the GTAP model, which accepts change in available agricultural land as an exogenous input.
+Specifically, we  solve both models concurrently for variable `qo` for land in the GTAP model and the parameter `RENTA` in the GTM model, which are internally consistent, i.e., the resulting rent change from agricultural land changes in the GTAP model are exactly the same as the changes in rents and the associated changes in agricultural land (calculated as the non-forested area from the total available land).
+Solving both models together in this way therefore allows us to expand their descriptive power.
 
@@
 Thus, if policymakers wish to see deforestation stopped, they could instead use import tariffs on products produced on land converted from forestland for agricultural use.
+Those scenarios, however, will be covered in a subsequent study.
 

### 2024-03-09 paper.Rmd
Update paper.Rmd

@@
     - \usepackage{setspace}\doublespacing
-abstract: "The projected growth in population and incomes is expected to create pressure to convert forestland into farmland.
-At the same time, the increasingly negative climate impacts are expected to generate further pressure to enhance the terrestrial carbon sink.
-Even though these goals are  incompatible as reversing the deforestation trend by afforesting cropland would result in negative market impacts such as higher food prices, using the GTAP and GTM models, we find that these impacts would be relatively small if the goal of  preserving 144.2 million hectares of forestland that otherwise would be converted to agricultural land by 2033 is achieved through a tax on using land in agricultural production.
-As to the economic price for doing so, the avoided deforestation would in most regions of the world result in less agricultural output and higher market prices.
-This is estimated to impact the well-being of global consumers by $92.1 billion, which translates to a global average cost of $10.6 per person in 2033."
+abstract: "The projected global growth in population and incomes is expected to create pressure to convert forestland into farmland, with an estimated 144.2 million hectares of forestland estimated to be converted to agricultural land by 2033.
+At the same time, increasingly negative climate impacts are expected to generate further pressure to enhance the terrestrial carbon sink, but taking such actions would likely result in less agricultural output and higher food prices.
+Using the GTAP and GTM models, we model the impact of applying a tax on converting forestland for use in agricultural production and find that the net deforestation rate could be brought to zero for a relatively small cost.
+Results suggest that the well-being of global consumers would be reduced by $92.1 billion in 2033, which translates to a global average annual cost of $10.6 per person."
 ---
@@
 Both the GTAP and GTM models represent widely used models for policy analysis.
-In the case of GTAP, the model is used to describe the implications of various policy instruments for national and global markets, including prices, output and trade.
+In the case of GTAP, the model is used to describe the implications of various policy instruments for national and global markets, including prices, output, and trade.
 The GTM model is, on the other hand, capable to translating the impacts of population and global growth, and the changes in land returns, to changes in global forest stocks and calculate the impacts on total carbon contained in forests.

### 2024-03-09 paper.Rmd
Update paper.Rmd

@@
 There are many reasons for this.
-For example, as nations develop and disposable incomes rise, consumers often increase meat consumpution.
+For example, as nations develop and disposable incomes rise, consumers often increase meat consumption.
 Population growth is also a driver.
@@
 
-# Modeling the impact of possible measures to mitigate deforestation/promote afforestation
+# Modeling the impact of possible measures to mitigate deforestation
 

### 2024-03-09 paper.Rmd
Update paper.Rmd

@@
 This is because of forests proven ability to sequester carbon not only in the stems of trees [@davis2022] but in their soils as well [@guo2021].
-With increasingly negative climate impacts, the pressure to enhance the terrestrial sink is growing, and one of the tools to achieve this is afforestation.
-As such, there is growing attention being placed globally to afforestation efforts, which is highlighted by the UN's declaration of the years from 2021 through 2030 as the UN Decade of Ecosystem Restoration [@sewell2020].
-Many of the Nationally Determined Contributions (NDCs) countries have submitted, which detail their plans to keep warming under the 2C threshold, also rely to a significant degree on afforestation projects [@roe2019].
-
-The global gains that can be realized from afforestation, however, are uncertain, with broad estimates of cumulative sequestration ranging from 42 gigatonnes to 700 gigatonnes by the end of the century [@veldman2019; @bastin2019; @humpenoder2014; @sathaye2006; @doelman2020].
-In addition, the amount of land conversion that would be required also varies greatly, with estimates ranging from 0.3 billion hectares to 2.8 billion hectares [@veldman2019; @bastin2019; @humpenoder2014; @doelman2020].
-In many cases, these results suggest massive losses in the amount of pasture and cropland that would be available.
-This puts into question the feasibility of attaining these results, as almost 90 percent of deforestation worldwide is undertaken to expand the amount of agricultural land [@fao2021].
+With increasingly negative climate impacts, there is increased attention being placed on maintaining these carbon sinks, but this requires, at minimum, avoiding further deforestation.
+However, from 2015 to 2020, roughly 10 million hectares of forestland was converted annually to other land uses [@].
 There are many reasons for this.
@@
 Population growth is also a driver.
-Currently, the United Nations' medium variant population projections suggest that between 2023 and 2033 the worlds' population will grow by 8.8 percent, a gain of 707.1 million people [@un2024].
-And where population growth is especially high, which is the case at present throughout Africa (2023 to 2033 population growth: 25.0%: 360.6 million), there is likely to be greater conversion of forestland into farmland.
-Still, some research has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017].
+Currently, the United Nations' medium variant population projections suggest that between 2023 and 2033 the world's population will grow by 8.8 percent, a gain of 707.1 million people [@un2024], and where population growth is especially high, which is the case at present throughout Africa (2023 to 2033 population growth: 25.0%: 360.6 million), there is likely to be greater conversion of forestland into farmland.
+This puts into question whether avoiding further deforestation is attainable.
 
-If such efforts are to be made, it is important to understand where to target them, as afforestation does not provide equal carbon benefits in every region.
-For example, afforestation in boreal zones provides the smallest net benefits, as the carbon sequestration benefits of conifer forests are mitigated by the reduction in albedo in winter [@mykleby2017; @li2015].
+Moreover, if efforts to end deforestation are to be made, it is important to understand where to target them, as forests do not provide equal carbon benefits in every region.
+For example, conifer forestland in boreal zones provides the smallest net benefits, as gains in carbon sequestration are mitigated by reductions in albedo in winter [@mykleby2017; @li2015].
 Coniferous forests have also been shown to increase soil organic carbon less than broadleaf forests [@guo2021].
-Conversely, tropical forests have been linked with a strong positive result from afforestation, as the carbon sequestration benefit is complemented by a cooling impact due to changes in both albedo and evapotranspiration, and with temperate forests, the impact of afforestation is also shown to be positive, albeit less strongly so due to a winter warming effect [@li2015].
-
-South America and Sub-Saharan Africa are thus two regions that show much potential for afforestation.
-It is estimated that these two regions hold at least 50% of the potential for global gains while regions such as Northern Africa and the Middle East show little promise as forest growth rates are quite low [@doelman2020].
-Thus, it is promising that the African Forest Landscape Restoration Initiative (AFR100), which aims to afforest 100 million hectares by 2030, has roughly 70% of its commitments coming from Sub-Saharan nations [@auda2023].
-There are other nations that have a big potential for gains from afforestation, but they face challenges.
-India, for example, has a great potential to increase its carbon sequestration through afforestation, but its massive population and food security worries limit its ability to consider such actions.
-The United Nations suggests that India's population will grow less rapidly than the world average between 2023 and 2033 (+8.5%), but due to its sizable base, the increase will be substantial nonetheless: 120.5 million.
-India's NDC does not entertain the idea of agricultural land being re-purposed and instead primarily considers targeting land that may not be well suited for afforestation [@amjath2019].
-Moreover, as land-use decisions in one country also tend to have impacts that spillover into many others [@meyfroidt2010; @lambin2011], analysis needs to be conducted at a global scale.
+Conversely, tropical forests have been linked with a strong positive results, as the carbon sequestration benefit is complemented by a cooling impact due to changes in both albedo and evapotranspiration, and with temperate forests, the impact is also shown to be positive, albeit less strongly so due to a winter warming effect [@li2015].
+As land-use decisions in one country also tend to have impacts that spillover into many others [@meyfroidt2010; @lambin2011], analysis needs to be conducted at a global scale.
 
@@
 While the GTAP-AEZ model benefits from the inclusion of competition among different crops, grazing, and forest-based uses, it has limitations, primarily due to its infrequent updates.
-The KLUM@GTAP framework integrated the Kleines Land Use Model (KLUM) with an extended yet static version of GTAP called GTAP-EFL to evaluate the impact of climate change on cropland allocation.
-KLUM is a global agricultural land-use model that connects the economy to global crop allocation to maximize producer returns under specific risk assumptions, and GTAP-EFL separates energy factors from intermediate inputs and incorporates them into capital, while also considering CO2 emissions.
-KLUM@GTAP substitutes the land allocation mechanism within GTAP-EFL by utilizing regionally aggregated area changes in cropland determined by KLUM to update cropland shares in GTAP-EFL [@ronneberger2009].
 
@@
 As the model applies nested CES production functions to all sectors and differentiates them by parameters only, to model one sector, in this case forestry, differently from the rest would require a major redesign of the model.
-While not impossible, this project will attempt to get around these concerns and assess the impact of various likely land-use scenarios by melding the dynamic GTAP model with the Global Timber Model (GTM), which is a dynamic optimization model that has been specifically designed to analyze the relationship among land rents, forested land cover, and carbon sequestration through forests and thus is better able to examine the impact of afforestation or avoided deforestation on the carbon sink [@sohngen2001].
+
+This project gets around these concerns and assesses the impact of various likely land-use scenarios by melding the dynamic GTAP model with the Global Timber Model (GTM), which is a dynamic optimization model that has been specifically designed to analyze the relationship among land rents, forested land cover, and carbon sequestration through forests and thus is better able to examine the impact of afforestation or avoided deforestation on the carbon sink [@sohngen2001].
 The latest versions of both models were used.
@@
 Surprisingly, Brazil and the United States, despite their decreases in available agricultural land compared to the baseline 2033 scenario, are estimated to see increased agricultural output of 0.6 percent and 1.0 percent, respectively.
-This happens when the sharp output decline in other regions prompts increased intensity of production in the United States and Brazil.
+This is the result of the sharp output declines in other regions prompting increased intensity of production in the United States and Brazil.
 

### 2024-03-09 paper.Rmd
Update paper.Rmd

@@
 With increasingly negative climate impacts, there is increased attention being placed on maintaining these carbon sinks, but this requires, at minimum, avoiding further deforestation.
-However, from 2015 to 2020, roughly 10 million hectares of forestland was converted annually to other land uses [@].
+However, from 2015 to 2020, roughly 10 million hectares of forestland was converted annually to other land uses [@fao2023].
 There are many reasons for this.
-For example, as nations develop and disposable incomes rise, consumers often increase meat consumption.
+For example, as nations develop and disposable incomes rise, consumers often increase meat consumption, which requires a greater quantity of agricultural land than is needed for vegetable-based diets [@elferink2007].
 Population growth is also a driver.

### 2024-03-09 paper.Rmd
Update paper.Rmd

@@
 There are many reasons for this.
-For example, as nations develop and disposable incomes rise, consumers often increase meat consumption, which requires a greater quantity of agricultural land than is needed for vegetable-based diets [@elferink2007].
+As nations develop and disposable incomes rise, for example, consumers often increase meat consumption, which requires a greater quantity of agricultural land than is needed for vegetable-based diets [@elferink2007].
 Population growth is also a driver.
@@
 Moreover, if efforts to end deforestation are to be made, it is important to understand where to target them, as forests do not provide equal carbon benefits in every region.
-For example, conifer forestland in boreal zones provides the smallest net benefits, as gains in carbon sequestration are mitigated by reductions in albedo in winter [@mykleby2017; @li2015].
-Coniferous forests have also been shown to increase soil organic carbon less than broadleaf forests [@guo2021].
+Coniferous forests in boreal zones are known to provide the smallest net benefits, as gains in carbon sequestration are mitigated by reductions in albedo in winter [@mykleby2017; @li2015].
+These forests have also been shown to increase soil organic carbon less than broadleaf forests [@guo2021].
 Conversely, tropical forests have been linked with a strong positive results, as the carbon sequestration benefit is complemented by a cooling impact due to changes in both albedo and evapotranspiration, and with temperate forests, the impact is also shown to be positive, albeit less strongly so due to a winter warming effect [@li2015].
+
 As land-use decisions in one country also tend to have impacts that spillover into many others [@meyfroidt2010; @lambin2011], analysis needs to be conducted at a global scale.
-
 Several notable modeling attempts have been made in this direction.
@@
 
-More recent modeling attempts of land-use change have begun to move away from comparative static analysis though.
+More recent modeling attempts of land-use change have begun to move away from comparative static analysis.
 For example, the dynamic GTAP model, supported by the GTAP-AEZ database, considers land market effects, which were identified as significant in driving results by @stevenson2013.

### 2024-03-09 paper.Rmd
Update paper.Rmd

@@
 Similar steps were taken to update and recalibrate the GTM model.
-As most forest inventory data are not updated annually, the GTM model updates data for regions with inventory information approximately every five years, and it updates data for other regions using the Global Forest Resource Assessment [@FAO2020].
+As most forest inventory data are not updated annually, the GTM model updates data for regions with inventory information approximately every five years, and it updates data for other regions using the Global Forest Resource Assessment [@fao2020].
 

### 2024-03-09 paper.Rmd
Update paper.Rmd

@@
 
-Both the GTAP and GTM models represent widely used models for policy analysis.
-In the case of GTAP, the model is used to describe the implications of various policy instruments for national and global markets, including prices, output, and trade.
-The GTM model is, on the other hand, capable to translating the impacts of population and global growth, and the changes in land returns, to changes in global forest stocks and calculate the impacts on total carbon contained in forests.
+The GTAP and GTM models are both widely used for policy analysis.
+The GTAP model is frequently used to describe the implications of various policy instruments on national and global markets and their prices, output, and trade.
+The GTM model is often used to estimate the impacts of changes in population, GDP, and land rents on global forest stocks and carbon sequestration.
+While both models capture important aspects of the global economy and the biophysical world, they do not model certain economic responses, as both consider several important variables exogenous.
+By combining the two models, we expand each model's descriptive power and thus produce a better depiction of the global economy, which should allow for a more precise assessment of the impact of policy actions.
 
-Even though the GTAP and GTM models capture important aspects of the global economy and the biophysical world, they do not model certain economic responses, and, instead, consider several important variables exogenous.
-In the case of GTAP, the model assumes that the amount of agricultural land is fixed, while the GTM considers the agricultural returns to land fixed.
-By combining the two models, we can fill each model's missing behavior for a more complete description of the global impacts.
+Specifically, we use the GTAP model to detail the impact on land rents based on the amount of available agricultural land.
+This benefits the GTM model, which uses returns as an exogenous input.
+We then use the GTM model to determine the change in agricultural land quantity based on the returns to land.
+This benefits the GTAP model, which takes the change in the amount of land available for crop and livestock production as an exogenous input.
+Using these links, we are able to solve both models concurrently, and outputs are passed back and forth between the models until we achieve a dual-state equilibrium.
+In the GTAP model, we solve for the variable `qo` for land, and in the GTM model, we solve for the parameter `RENTA`.
+These are internally consistent, i.e., the resulting rent change from agricultural land changes in the GTAP model mirrors the associated change in agricultural land (calculated as the non-forested area from the total available land) from the change in rents in GTM.
 
-The combination of the two models means that we use the GTAP model to create a link between the amount of available land and its returns for the benefit of the GTM model, which uses returns as an exogenous input, while at the same time we use the GTM model to create the response of  agricultural land quantity to returns in land returns for the benefit of the GTAP model, which accepts change in available agricultural land as an exogenous input.
-Specifically, we  solve both models concurrently for variable `qo` for land in the GTAP model and the parameter `RENTA` in the GTM model, which are internally consistent, i.e., the resulting rent change from agricultural land changes in the GTAP model are exactly the same as the changes in rents and the associated changes in agricultural land (calculated as the non-forested area from the total available land).
-Solving both models together in this way therefore allows us to expand their descriptive power.
-
-We achieve a combined solution of GTAP and GTM by passing outputs back and forth between the models until a dual-state equilibrium is achieved.
-Establishing a connection between the GTAP model and the GTM model requires that we  model the return on land using the GTAP model and feed those results into the GTM model as exogenous inputs.
-The GTM model then endogenously determines the amount of land available for crop and livestock production, and we feed that back into the GTAP model as an exogenous input.
-
-On the software level, we establish the connection between the models using an R script, which is capable of handling other software, including generating required inputs, executing it and reading in any outputs.
-This is required to operate the GTAP model, which is written in GEMPACK, and the GTM model, which is written in GAMS.
-As GEMPACK and GAMS were developed long before software integration became important or even possible, we use much more modern technologies made available through the R software [@R2021] to create the user functions that execute both models, collect their outputs, and automatically create any required input files.
+On the software level, we establish the connection between the models using R software [@R2021].
+This is required as the two models were written using using different languages.
+Moreover, GTAP, which was written in GEMPACK, and GTM, which was written in GAMS, were developed long before software integration became important or even possible, and so we use an R script, which is capable of handling other software, including generating required inputs, executing them, and reading in any outputs, to create the user functions that execute both models, collect their outputs, and automatically create any required input files.
 

### 2024-03-09 paper.Rmd
Update paper.Rmd

@@
 At the same time, increasingly negative climate impacts are expected to generate further pressure to enhance the terrestrial carbon sink, but taking such actions would likely result in less agricultural output and higher food prices.
-Using the GTAP and GTM models, we model the impact of applying a tax on converting forestland for use in agricultural production and find that the net deforestation rate could be brought to zero for a relatively small cost.
+Using the GTAP and GTM models, we model the impact of applying a tax on the value of output from land used in agricultural production and find that the net deforestation rate could be brought to zero for a relatively small cost.
 Results suggest that the well-being of global consumers would be reduced by $92.1 billion in 2033, which translates to a global average annual cost of $10.6 per person."
@@
 
-To understand the impact of potential regional and global afforestation actions, an important first step was creating a realistic baseline scenario that depicts the likely changes over the decade under study, 2023--2033.
+To understand the potential for policy actions to achieve net zero global deforestation, an important first step was creating a realistic baseline scenario that depicts the likely changes over the decade under study, 2023--2033.
 Data for this were drawn from CEPII [@CEPII2024].
@@
 
-Focusing on the absolute area of the forested areas, as generated by the GTM model, we estimate that between 2023 and 2033, 144.2 million hectares (3.6%) of global forests will be converted to agricultural lands with the largest decline expected to happen in Sub-Saharan Africa, which is estimated to lose 39.7 million hectares (7.0%) of its forestland (Table \@ref(tab:baseline-forest-change)).
+Focusing on the absolute area of the forestland, as generated by the GTM model, we estimate that between 2023 and 2033, 144.2 million hectares (3.6%) of global forests will be converted to agricultural lands with the largest decline expected to happen in Sub-Saharan Africa, which is estimated to lose 39.7 million hectares (7.0%) of its forestland (Table \@ref(tab:baseline-forest-change)).
 Southeast Asia, Brazil, and the rest of South America also see forestland shrink, with decreases of 22.6 million hectares (9.3%), 19.0 million hectares (3.8%), and 23.5 million hectares (6.7%), respectively.
@@
 This is the result of the sharp output declines in other regions prompting increased intensity of production in the United States and Brazil.
+When we examine how the policy impacts exports of agricultural goods, we find that exports from Japan and the United States grow the most, increasing 8.8 percent and 3.5 percent, respectively (Table 4).
 
@@
 
-When we examine how the policy impacts exports of agricultural goods, we find that exports from Japan and the United States grow the most, increasing 8.8 percent and 3.5 percent, respectively (Table 4).
-The regions where agricultural exports drop most are East Asia, South Asia, Southeast Asia, and Central America, which see decreases of 14.5 percent, 5.9 percent, 5.1 percent, and 5.0 percent, respectively.
+A visual depiction of how this policy would impact global trade is presented in Figure 1.
+The United States would be expected to see its agricultural exports increase significantly.
+Conversely, the regions where agricultural exports would be expected to drop the most are East Asia, South Asia, Southeast Asia, and Central America, which would see decreases of 14.5 percent, 5.9 percent, 5.1 percent, and 5.0 percent, respectively.
 
@@
 
-Overall, the tax policy, which results in zero net deforestation globally, has a rather substantial negative economic impact on all regions except Japan and Canada, which see an increase in consumer well-being of
+Overall, the tax policy, which results in zero net deforestation globally, has a negative economic impact on all regions except Japan and Canada, which see an increase in consumer well-being of
 $977.1 million and $166.9 million, respectively (Table 5).
@@
 The most negatively impacted regions are South Asia (-$28.0 billion), China (-$14.5 billion), and Southeast Asia (-$13.6 billion).
-In total, this tax-driven policy is estimated to decrease global consumer well-being in 2033 by $92.1 billion relative to the scenario where no tax on forestland conversion was in place.
+In total, this tax-driven policy is estimated to decrease global consumer well-being in 2033 by $92.1 billion relative to the scenario where no tax on agricultural land was in place.
 
@@
 
-This analysis shows that, at an estimated tax rate of 70.3 percent of the value of each hectare of land that is converted to agricultural land, global net deforestation would drop to roughly zero by 2033, preserving 144.2 million hectares of forestland that otherwise would have been converted to agricultural land.
-In the United States, as USDA-NASS (Land Values 2022 Summary (August 2022)) valued the average hectare of farmland in 2022 at $1,537.8, the tax would equate to $1,081.1/hectare.
-In Brazil, CEIC valued the average price of Brazilian land at roughly 2,500 Brazilian Reals, which would put the tax at roughly $82.2 per hectare.
-The economic price of this would be transfered to consumers through lower agricultural output and higher market prices.
-Overall, this land tax would reduce consumer well-being by $92.1 billion or $10.60 per person in 2033.
+This analysis shows that, at an estimated global tax rate of 70.3 percent of the value of output from land used in agricultural production, global net deforestation would drop to roughly zero by 2033, preserving 144.2 million hectares of forestland that otherwise would have been converted to agricultural land.
+The economic price of this would be transferred to consumers through lower agricultural output and higher market prices.
+Overall, this land tax would reduce consumer well-being by $92.1 billion in 2033, which equates to $10.60 per person annually.
 Governments might be wary of farmer discontent were they to take such fiscal actions and thus might be reluctant to impose such taxes on them.
@@
 
-We also show the utility of joining two major economic models.
+This work also shows the utility of joining two major economic models.
 GTAP has long been valued for its ability to provide robust information on global trade flows and prices, and GTM has proven its utility in understanding the impact of forestry decisions on the carbon sink and land rents.
-By combining these models through the use of R and allowing the models to pass inputs and outputs back and forth iteratively, the benefits of both models have been maintained and their weaknesses greatly minimized.
+By combining these models through the use of R software and allowing the models to pass inputs and outputs back and forth iteratively, the benefits of both models have been maintained and their weaknesses greatly minimized.
 This advance holds great promise for advancement on a variety of fronts for researchers and policymakers interested in climate change, agricultural trade, and their inter-dependencies.
@@
 
-According to CEPII's projections, the growth of the global economy in the next ten years is expected to result in tremendous increases in global and regional GDP accompanied by a moderate growth in population.
-Another set of projections by GTM suggests that this growth would also mean significant reductions in the world's forests because they would be converted to agricultural land.
-The model predicts a loss of over 144 million hectares of forest before 2033.
+According to CEPII’s projections, the growth of the global economy in the next ten years is expected to result in tremendous increases in global and regional GDP accompanied by a moderate growth in population.
+Another set of projections by GTM suggests that this growth would also mean significant reductions in the world’s forests because they would be converted to agricultural land, with the model predicting a loss of over 144 million hectares of forest by 2033.
 
-By combining two well-known global models---the GTAP model that describes the world markets and trade, and the GTAM model that describes the impact of economic and population growth along with land rents on the amounts of forested areas---we were able to estimate the size of the tax on agricultural land needed to prevent deforestation in the coming ten years through reducing the returns to land when used in agriculture.
+By combining two well-known global models—the GTAP model that describes the world markets and trade, and the GTM model that describes the impact of economic and population growth along with land rents on the amounts of forested areas—we were able to estimate the size of the tax on agricultural land needed to
+
+prevent deforestation in the coming ten years through reducing the returns to land when used in agriculture.
 We find that even though the tax is substantial at about 70 percent, it results in only very small reductions in the projected growth in global GDP and welfare.
@@
 
-Finally, we note that a uniform global tax reducing the returns to land to the point of preventing global deforestation would produce highly differentiated impacts around the world's regions.
-Some regions, such as China, Russia, Canada and the United States would increase their forested areas as a result of the tax, while the regions with the greatest deforestation pressures, such as Sub-Saharan Africa, Brazil and Southeast Asia would only reduce the rate of their deforestation, for a zero change in the global forest area.
+Finally, we note that a uniform global tax reducing the returns to land to the point of preventing global deforestation would produce highly differentiated impacts around the world’s regions.
+Some regions, such as China, Russia, Canada, and the United States would increase their forested areas as a result of the tax, while the regions with the greatest deforestation pressures, such as Sub-Saharan Africa, Brazil, and Southeast Asia would only reduce the rate of their deforestation, for a zero change in the global forest area.
+Therefore, future studies will need to consider how best to target policies to counteract deforestation in zones facing the greatest land use conversion pressures.
 

### 2024-03-14 paper.Rmd
Update paper.Rmd

@@
 At the same time, increasingly negative climate impacts are expected to generate further pressure to enhance the terrestrial carbon sink, but taking such actions would likely result in less agricultural output and higher food prices.
-Using the GTAP and GTM models, we model the impact of applying a tax on the value of output from land used in agricultural production and find that the net deforestation rate could be brought to zero for a relatively small cost.
-Results suggest that the well-being of global consumers would be reduced by $92.1 billion in 2033, which translates to a global average annual cost of $10.6 per person."
+Using the GTAP and GTM models, we model how different tax rates on the value of output from land used in agricultural production would impact deforestation.
+We find that, with a tax rate of zero percent on agricultural production, 144.2 million hectares (3.6%) of global forests would be converted to agricultural lands by 2033.
+With a 70 percent tax, however, the net deforestation rate would be brought to zero while the well-being of global consumers would be reduced by $92.1 billion in 2033, which translates to a global average annual cost of $10.6 per person."
 ---

### 2024-03-14 paper.Rmd
Update paper.Rmd

@@
 To solve the system simultaneously, we require that $V_1$ and $V_2$ be equal.
+We focus on the relative returns of land in agriculture and forestry as they greatly impact land use changes.
+Thus, in our scenarios, we examine the impact of different tax rates on the use of land in the production of agricultural output.
 With several iterations, the system converges to the change in land and land rent that is consistent across the models, i.e., the land change results in the land rent changes that, when entered into the GTM model, result in the same land changes as assumed by the GTAP model.
@@
 
-# Changes under the baseline projections
+# Changes under the zero tax scenario
 
-To understand the potential for policy actions to achieve net zero global deforestation, an important first step was creating a realistic baseline scenario that depicts the likely changes over the decade under study, 2023--2033.
-Data for this were drawn from CEPII [@CEPII2024].
-To build our scenario, we use two of their variables that we determined to be key: population and GDP.
-The CEPII projections do not contain any information on land use.
-To build a more complete baseline, we also include the projections on forested land change included in the GTM.
-Using the GTAP model in connection with the GTM model (to assure that the changes in land are consistent between the two models), we calculate the overall change in factor quantity/productivity growth required to achieve the targeted values in 2033.
+Our first scenario examines the likely changes over the decade under study, 2023--2033.
+To build our scenario, we use data from CEPII [@CEPII2024] and focus on two of their variables that we determined to be key: population and GDP.
 In Table \@ref(tab:baseline-growth), we show the projected population and GDP growth per region.
@@
 Japan, Russia, and East Asia are projected to see their populations decline by 10.4%, 4.7%, and 2.6%, respectively.
-As for the projected change in agricultural land, our baseline projections show a significant increase in Oceania (+27.6%), Rest of Latin America (+25.8%), Sub-Saharan Africa (+17.4%), Central America (+13.9%), and Brazil (+13.9%).
+As the CEPII projections do not contain any information on land use, we also include the projections on forested land change included in the GTM.
+Using the GTAP model in connection with the GTM model (to assure that the changes in land are consistent between the two models), we then calculate the overall change in factor quantity/productivity growth required to achieve the targeted values in 2033 while assuming a uniform zero percent tax on the use of land in the production of agricultural output.
+Results suggest that there will be a significant increase in agricultural land in Oceania (+27.6%), Rest of Latin America (+25.8%), Sub-Saharan Africa (+17.4%), Central America (+13.9%), and Brazil (+13.9%).
 Japan is the only nation that sees a significant decline (-34.9%) in agricultural land, which is likely tied in part to population decline.
@@
 
-Focusing on the absolute area of the forestland, as generated by the GTM model, we estimate that between 2023 and 2033, 144.2 million hectares (3.6%) of global forests will be converted to agricultural lands with the largest decline expected to happen in Sub-Saharan Africa, which is estimated to lose 39.7 million hectares (7.0%) of its forestland (Table \@ref(tab:baseline-forest-change)).
+Focusing on the absolute area of the forestland, as generated by the GTM model, we estimate that under this scenario between 2023 and 2033, 144.2 million hectares (3.6%) of global forests will be converted to agricultural lands with the largest decline expected to happen in Sub-Saharan Africa, which is estimated to lose 39.7 million hectares (7.0%) of its forestland (Table \@ref(tab:baseline-forest-change)).
 Southeast Asia, Brazil, and the rest of South America also see forestland shrink, with decreases of 22.6 million hectares (9.3%), 19.0 million hectares (3.8%), and 23.5 million hectares (6.7%), respectively.
@@
 
-Against the backdrop of the baseline scenario, we turn to the question of what policy could prevent global deforestation in the period of 2023–2033.
-Because the area of forestland is mostly determined by the relative returns of land in agriculture and forestry, in our policy scenario we modify the return to land in agriculture by applying a uniform global tax on the use of land in the production of agricultural output.
-Solving the two models together shows that it would require a tax of roughly 70 percent to be assessed on the use of land for the deforestation rate to fall globally to zero.
-The economic impact of such a policy measure would not be limited to the cost of the tax, it would also impact food production and its prices.
-For example, with this policy in place, most regions of the world would see significant declines in the amount of agricultural land in 2033 relative to the baseline scenario where land conversion was not taxed.
+Our second scenario examines the impact of a tax of roughly 70 percent on the use of agricultural land.
+Solving the two models together shows that the economic impact would not be limited to the cost of the tax, it would also impact food production and its prices.
+For example, most regions of the world would see significant declines in the amount of agricultural land in 2033 relative to the baseline scenario where land conversion was not taxed.
 For the United States, there would be 2.1 percent less agricultural land (Table 3).
 Larger decreases would happen in other major agricultural production regions such as China (-2.5 percent), South Asia (-7.9 percent), Brazil (-8.6 percent), and Russia (-9.3 percent).
+This would be because the global net deforestation rate would be brought to zero and 144.2 million hectares that would otherwise have been converted to agricultural land by 2033 would be preserved as forestland.
 
@@
 
-A visual depiction of how this policy would impact global trade is presented in Figure 1.
+A visual depiction of how this tax would impact global trade is presented in Figure 1.
 The United States would be expected to see its agricultural exports increase significantly.
@@
 
-Overall, the tax policy, which results in zero net deforestation globally, has a negative economic impact on all regions except Japan and Canada, which see an increase in consumer well-being of
+Overall, the 70 percent tax, which results in zero net deforestation globally, is liekly to have a negative economic impact on all regions except Japan and Canada, which see an increase in consumer well-being of
 $977.1 million and $166.9 million, respectively (Table 5).
 Consumer well-being is a measurement of equivalent variation and here measures the benefit/harm consumers experience in terms of income as a result of the policy changes.
-The United States, along with 10 other regions, see well-being drop by more than $1 billion.
-The most negatively impacted regions are South Asia (-$28.0 billion), China (-$14.5 billion), and Southeast Asia (-$13.6 billion).
+The United States, along with 10 other regions, are estimated to see well-being drop by more than $1 billion.
+The most negatively impacted regions would be South Asia (-$28.0 billion), China (-$14.5 billion), and Southeast Asia (-$13.6 billion).
 In total, this tax-driven policy is estimated to decrease global consumer well-being in 2033 by $92.1 billion relative to the scenario where no tax on agricultural land was in place.
@@
 Much of this loss in well-being may be the result of higher prices.
-Japan is the only nation that sees any price impacts that are beneficial for consumers resulting from this policy, with the price of agricultural goods decreasing 0.3 percent (Table \@ref(tab:pm)).
+Japan is the only nation that is estimated to see any price impacts that are beneficial for consumers, with the price of agricultural goods decreasing 0.3 percent (Table \@ref(tab:pm)).
 All other regions, for both agricultural and non-agricultural products, see prices rise between 0.3 and 5.1 percent.

### 2024-03-14 paper.Rmd
Update paper.Rmd

@@
 ---
-title: "Avoiding global deforestation by taxing land in agricultural production: The implications for global markets"
+title: "The global implications of taxing land in agricultural production: Land use, agricultural production, and consumer well-being"
 date: "`r format(Sys.time(), '%B %d, %Y')`"
@@
 Larger decreases would happen in other major agricultural production regions such as China (-2.5 percent), South Asia (-7.9 percent), Brazil (-8.6 percent), and Russia (-9.3 percent).
-This would be because the global net deforestation rate would be brought to zero and 144.2 million hectares that would otherwise have been converted to agricultural land by 2033 would be preserved as forestland.
+This would be because the global net deforestation rate would be brought to roughly zero, meaning that the 144.2 million hectares that would have been converted to agricultural land by 2033 in the zero tax scenario would be preserved as forestland.
 
-These reductions in the amount of agricultural land would then impact the quantity of agricultural goods that could be produced.
+This would have impacts on the quantity of agricultural goods that could be produced.
 For example, the 15.0 percent decrease in agricultural land in East Asia and the 9.0 percent decrease in agricultural land in Southeast Asia would result in reductions in agricultural production of 5.0 percent and 1.7 percent, respectively.
@@
 
-Overall, the 70 percent tax, which results in zero net deforestation globally, is liekly to have a negative economic impact on all regions except Japan and Canada, which see an increase in consumer well-being of
+Overall, the 70 percent tax, which results in zero net deforestation globally, is likely to have a negative economic impact on all regions except Japan and Canada, which see an increase in consumer well-being of
 $977.1 million and $166.9 million, respectively (Table 5).
@@
 
-This analysis shows that, at an estimated global tax rate of 70.3 percent of the value of output from land used in agricultural production, global net deforestation would drop to roughly zero by 2033, preserving 144.2 million hectares of forestland that otherwise would have been converted to agricultural land.
-The economic price of this would be transferred to consumers through lower agricultural output and higher market prices.
-Overall, this land tax would reduce consumer well-being by $92.1 billion in 2033, which equates to $10.60 per person annually.
-Governments might be wary of farmer discontent were they to take such fiscal actions and thus might be reluctant to impose such taxes on them.
-Thus, if policymakers wish to see deforestation stopped, they could instead use import tariffs on products produced on land converted from forestland for agricultural use.
-Those scenarios, however, will be covered in a subsequent study.
+This analysis shows that tax rates on agricultural land can have a significant impact on land use.
+At a global average tax rate of zero percent of the value of land used in agricultural production, an estimated 144.2 million hectares of forestland would be converted to agricultural land by 2033.
+With an estimated global tax rate of 70.3 percent, global net deforestation would drop to roughly zero by 2033, preserving 144.2 million hectares of forestland.
+The economic price of this would likely be transferred to consumers through lower agricultural output and higher market prices.
+Overall, this land tax is estimated to reduce consumer well-being by $92.1 billion in 2033, which equates to $10.60 per person annually.
 
@@
 Another set of projections by GTM suggests that this growth would also mean significant reductions in the world’s forests because they would be converted to agricultural land, with the model predicting a loss of over 144 million hectares of forest by 2033.
-
-By combining two well-known global models—the GTAP model that describes the world markets and trade, and the GTM model that describes the impact of economic and population growth along with land rents on the amounts of forested areas—we were able to estimate the size of the tax on agricultural land needed to
-
-prevent deforestation in the coming ten years through reducing the returns to land when used in agriculture.
-We find that even though the tax is substantial at about 70 percent, it results in only very small reductions in the projected growth in global GDP and welfare.
-We also observe that the reduction in available agricultural land relative to the baseline growth results in reduced agricultural output and higher prices, which may require additional policy action to avoid negative impacts in terms of food security.
-
-Finally, we note that a uniform global tax reducing the returns to land to the point of preventing global deforestation would produce highly differentiated impacts around the world’s regions.
-Some regions, such as China, Russia, Canada, and the United States would increase their forested areas as a result of the tax, while the regions with the greatest deforestation pressures, such as Sub-Saharan Africa, Brazil, and Southeast Asia would only reduce the rate of their deforestation, for a zero change in the global forest area.
-Therefore, future studies will need to consider how best to target policies to counteract deforestation in zones facing the greatest land use conversion pressures.
+By combining two well-known global models—the GTAP model that describes the world markets and trade, and the GTM model that describes the impact of economic and population growth along with land rents on the amounts of forested areas—we were able to estimate the impact of a tax on agricultural land on land use change.
+We find that a tax of about 70 percent would likely prevent net global deforestation in the coming ten years through reducing the returns to land when used in agriculture.
+We also observe that the reduction in available agricultural land relative to the baseline growth would likely result in reduced agricultural output and higher prices, which might have food security implications.
+Finally, we note that such a uniform global tax that would reduce the returns to land to the point of preventing global deforestation appears to produce highly differentiated impacts around the world’s regions.
+Some regions, such as China, Russia, Canada, and the United States are estimated to see their forested areas increase as a result of the tax, while the regions with the greatest deforestation pressures, such as Sub-Saharan Africa, Brazil, and Southeast Asia would only reduce the rate of their deforestation, for a zero change in the global forest area.
 

### 2024-03-14 paper.Rmd
Update paper.Rmd

@@
 
-This project gets around these concerns and assesses the impact of various likely land-use scenarios by melding the dynamic GTAP model with the Global Timber Model (GTM), which is a dynamic optimization model that has been specifically designed to analyze the relationship among land rents, forested land cover, and carbon sequestration through forests and thus is better able to examine the impact of afforestation or avoided deforestation on the carbon sink [@sohngen2001].
+This project gets around these concerns and assesses the impact of various tax scenarios on land-use by melding the dynamic GTAP model with the Global Timber Model (GTM), which is a dynamic optimization model that has been specifically designed to analyze the relationship among land rents, forested land cover, and carbon sequestration through forests and thus is better able to examine the impact of afforestation or avoided deforestation on the carbon sink [@sohngen2001].
 The latest versions of both models were used.
@@
 
-# Modeling the impact of possible measures to mitigate deforestation
+# Modeling the impact of higher tax scenarios
 
